---
title: "Network characteristics of ensemble networks"
author: "Hanne Puype"
date: '2022-03-25'
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

# How many genes per module
```{r}
library(tidyverse)
modulesAD <- read.table("modules_AD.txt", header = F, sep = "\t")
head(modulesAD)
modulesAD %>% group_by(V2) %>% summarise(n())
modulesMDD <- read.table("modules_MDD.txt", header = F, sep = "\t")
head(modulesMDD)
modulesMDD %>% group_by(V2) %>% summarise(n())
```

# Visualise modules in PCA plot
```{r}
countsAD <- read.table("Processed_counts_AD_GS.txt", sep = "\t", header = T)
countsMDD <- read.table("Processed_counts_MDD_GS.txt", sep = "\t", header = T)
nodesAD <- unique(modulesAD$V1)
length(nodesAD)
nodesMDD <- unique(modulesMDD$V1)
countsAD <- countsAD %>% filter(GeneSymbol %in% nodesAD)
dim(countsAD)
tail(countsAD)
countsMDD <- countsMDD %>% filter(GeneSymbol %in% nodesMDD)
dim(countsMDD)
```
the genes are in the same order in the modules and in the counts  

```{r}
countsAD <- column_to_rownames(countsAD, "GeneSymbol")
countsAD_t <- t(countsAD)
pca <- prcomp(countsAD_t)
modulesAD %>%
  mutate(PC1 = pca$rotation[, 1], PC2 = pca$rotation[, 2]) %>%
  ggplot(aes(x = PC1, y = PC2, colour = V2)) +
        geom_point()
```

```{r}
countsMDD <- column_to_rownames(countsMDD, "GeneSymbol")
countsMDD_t <- t(countsMDD)
pca <- prcomp(countsMDD_t)
modulesMDD %>%
  mutate(PC1 = pca$rotation[, 1], PC2 = pca$rotation[, 2]) %>%
  ggplot(aes(x = PC1, y = PC2, colour = V2)) +
        geom_point()
```

# Modules with jaccard index 
k = 155 and 156  
```{r}
modulesAD <- read.table("kmed_modules_AD.txt", header = F, sep = "\t")
head(modulesAD)
modulesAD %>% group_by(V2) %>% summarise(n())
modulesMDD <- read.table("kmed_modules_MDD.txt", header = F, sep = "\t")
head(modulesMDD)
modulesMDD %>% group_by(V2) %>% summarise(n())
```
module 28 of AD has 1162 genes  
module 5 of MDD has 1086 genes


# Hubs in network
```{r}
network_AD <- read.table("ensemble_netw_AD.txt", header = T, sep = "\t")
head(network_AD)
regAD <- network_AD$TF 
regAD <- as.data.frame(table(regAD)) # frequency TF --> ~ out-degree
regAD <- regAD %>% arrange(desc(Freq))
head(regAD)
regAD$regAD <- as.character(regAD$regAD) # are factors
```

```{r}
network_MDD <- read.table("ensemble_netw_MDD.txt", header = T, sep = "\t")
head(network_MDD)
regMDD <- network_MDD$TF 
regMDD <- as.data.frame(table(regMDD))
regMDD <- regMDD %>% arrange(desc(Freq))
head(regMDD)
regMDD$regMDD <- as.character(regMDD$regMDD)
```


```{r}
regAD <- regAD %>% slice(c(1:100)) %>% select(regAD)
regAD <- regAD$regAD
print(regAD, quote = F)
write.table(regAD, file = "Hubs_AD_bulk.txt", sep = "\t")
regMDD <- regMDD %>% slice(c(1:100)) %>% select(regMDD)
regMDD <- regMDD$regMDD
print(regMDD, quote = F)
write.table(regMDD, file = "Hubs_MDD_bulk.txt", sep = "\t")
common_reg <- Reduce(intersect, list(regAD, regMDD)) 
length(common_reg)
print(common_reg, quote = F)
```


# Visualisation 
```{r fig.width=10}
library(igraph)
igraph_options(annotate.plot = T)
graph_AD <- graph_from_data_frame(network_AD, directed = T)
graph_MDD <- graph_from_data_frame(network_MDD, directed = T)
pdf("graph_AD_ens.pdf", 
    width = 9, height = 8, 
    family = "ArialMT")
plot(graph_AD, vertex.label = NA, main = "Graph of ensemble network of AD", vertex.color = "#88CCEE")
dev.off()
pdf("graph_MDD_ens.pdf", 
    width = 9, height = 8, 
    family = "ArialMT")
plot(graph_MDD, vertex.label = NA, main = "Graph of ensemble network of MDD", vertex.color = "#88CCEE")
dev.off()
```

# Compare networks with Jaccard index
Jaccard Similarity = (number of observations in both sets) / (number in either set)  
https://www.statology.org/jaccard-similarity-in-r/  

## Jaccard similarity index
```{r}
jaccard <- function(a, b) {
    intersection = length(intersect(a, b))
    union = length(a) + length(b) - intersection
    return (intersection/union)
}
```

Edge list 
```{r}
head(network_AD)
network_AD$concat <- paste(network_AD$TF, network_AD$Target_gene, sep = ":")
network_MDD$concat <- paste(network_MDD$TF, network_MDD$Target_gene, sep = ":")
edgesAD <- network_AD$concat
edgesMDD <- network_MDD$concat
```

```{r}
jaccard(edgesAD, edgesMDD)
```
the higher the number, the more similar  

## Jaccard distance
```{r}
1- jaccard(edgesAD, edgesMDD)
```

```{r}
length(union(edgesAD, edgesMDD)) # all edges of both networks
length(intersect(edgesAD, edgesMDD)) # common edges
```

## Venn diagram of overlap
```{r}
library("ggvenn")
x <- list(edgesAD, edgesMDD)
names(x) <- c("AD","MDD")
ggvenn(
  x, show_percentage = F, 
  fill_color = c("#0073C2FF", "#CD534CFF"),
  stroke_size = 0.5, set_name_size = 7, text_size = 6) + 
  labs(title = "Venn diagram for overlap ensemble networks")
```

## Ranks of overlapping edges
```{r}
intersec_AD_MDD <- intersect(edgesAD, edgesMDD)
network_AD_shared <- network_AD %>% filter(concat %in% intersec_AD_MDD)
dim(network_AD_shared)
head(network_AD_shared)
mean(network_AD_shared$Rank)
median(network_AD_shared$Rank)
hist(network_AD_shared$Rank, breaks = 35, main = "Rank of shared edges of AD network", xlab = "Rank")

network_MDD_shared <- network_MDD %>% filter(concat %in% intersec_AD_MDD)
dim(network_MDD_shared)
head(network_MDD_shared)
mean(network_MDD_shared$Rank)
median(network_MDD_shared$Rank)
hist(network_MDD_shared$Rank, breaks = 35, main = "Rank of shared edges of MDD network", xlab = "Rank")
```

# Annotation for ModuleViewer  
## AD 
```{r}
annotation_AD <- read.delim("C:/UGent/Master 2/Master's dissertation/R/Lemon-Tree/annotation_AD.txt")
head(annotation_AD)
annotation_AD$Diagnosis <- as.factor(annotation_AD$Diagnosis)
annotation_AD$Sex <- as.factor(annotation_AD$Sex)
annotation_AD$Tangle.Stage <- as.factor(annotation_AD$Tangle.Stage)
annotation_AD$Plaque.Stage <- as.factor(annotation_AD$Plaque.Stage)
annotation_AD$Neuropath.Dx.1 <- as.factor(annotation_AD$Neuropath.Dx.1)
levels(annotation_AD$Diagnosis) #2
levels(annotation_AD$Neuropath.Dx.1) #4
levels(annotation_AD$Sex) #2
levels(annotation_AD$Tangle.Stage) #6
levels(annotation_AD$Plaque.Stage) #4
```

Diagnosis
```{r}
annotation_AD$col <- factor(annotation_AD$Diagnosis, levels=c("AD","Control"), 
                         labels=c("#eb764f","#44BB99"))
annotation_AD$cond <- paste(annotation_AD$Sample, annotation_AD$col, sep=":")
head(annotation_AD)
annot_AD <- t(annotation_AD[,8])
write.table(annot_AD, file = "ModuleViewer/Ens_Anno_Diagn_AD", quote = F, sep = '|', na = "", row.names = F,
            col.names = F)
```

Sex
```{r}
annotation_AD$col <- factor(annotation_AD$Sex, levels=c("male","female"), 
                         labels=c("#99DDFF","#FFAABB"))
annotation_AD$cond <- paste(annotation_AD$Sample, annotation_AD$col, sep=":")
head(annotation_AD)
annot_AD <- t(annotation_AD[,8])
write.table(annot_AD, file = "ModuleViewer/Ens_Anno_Sex_AD", quote = F, sep = '|', na = "", row.names = F,
            col.names = F)
```

Neuropath.Dx.1
```{r}
annotation_AD$col <- factor(annotation_AD$Neuropath.Dx.1, levels=c("Alzheimer's disease",
                                                                   "Normal (Mild Braak Changes)", 
                                                                   "Normal - No Pathology Detected", 
                                                                   "Normal (Mild Vascular Changes)"), 
                         labels=c("#004C99", "#0080FF", "#FFFFFF", "#99CCFF"))
annotation_AD$cond <- paste(annotation_AD$Sample, annotation_AD$col, sep=":")
head(annotation_AD)
annot_AD <- t(annotation_AD[,8])
write.table(annot_AD, file = "ModuleViewer/Ens_Anno_Neuropath_AD", quote = F, sep = '|', na = "", row.names = F,
            col.names = F)
```

Tangle Stage
```{r}
annotation_AD$col <- factor(annotation_AD$Tangle.Stage, levels=c("Stage 1","Stage 2", "Stage 3", "Stage 4",
                                                                 "Stage 5", "Stage 6"), 
                         labels=c("#FFFFFF", "#CCFFCC", "#66FF66", "#00CC00", "#009900", "#006600"))
annotation_AD$cond <- paste(annotation_AD$Sample, annotation_AD$col, sep=":")
head(annotation_AD)
annot_AD <- t(annotation_AD[,8])
write.table(annot_AD, file = "ModuleViewer/Ens_Anno_Tangle_AD", quote = F, sep = '|', na = "", row.names = F,
            col.names = F)
```

Plaque stage
```{r}
annotation_AD$col <- factor(annotation_AD$Plaque.Stage, levels=c("Stage A","Stage B", "Stage C", "None"), 
                         labels=c("#FFCC99", "#FF9933", "#994C00", "#FFFFFF"))
annotation_AD$cond <- paste(annotation_AD$Sample, annotation_AD$col, sep=":")
head(annotation_AD)
annot_AD <- t(annotation_AD[,8])
write.table(annot_AD, file = "ModuleViewer/Ens_Anno_Plaque_AD", quote = F, sep = '|', na = "", row.names = F,
            col.names = F)
```

## MDD 

```{r}
annotation_MDD <- read.delim("C:/UGent/Master 2/Master's dissertation/R/Lemon-Tree/annotation_MDD.txt")
head(annotation_MDD)
annotation_MDD$Diagnosis <- as.factor(annotation_MDD$Diagnosis)
annotation_MDD$Sex <- as.factor(annotation_MDD$Sex)
annotation_MDD <- annotation_MDD %>% dplyr::select(-BioProject)
levels(annotation_MDD$Diagnosis) #4
levels(annotation_MDD$Sex) #2
```

Diagnosis
```{r}
annotation_MDD$col <- factor(annotation_MDD$Diagnosis, levels=c("Major Depression", "Control",  
                                                                "Major depression non-suicide", 
                                                                "Major depression suicide"), 
                         labels=c("#eb764f","#44BB99", "#eb764f", "#eb764f"))
annotation_MDD$cond <- paste(annotation_MDD$Sample, annotation_MDD$col, sep=":")
head(annotation_MDD)
annot_MDD <- t(annotation_MDD[,5])
write.table(annot_MDD, file = "ModuleViewer/Ens_Anno_Diagn_MDD", quote = F, sep = '|', na = "", row.names = F,
            col.names = F)
```
Sex
```{r}
annotation_MDD$col <- factor(annotation_MDD$Sex, levels=c("male", "female"), 
                         labels=c("#99DDFF","#FFAABB"))
annotation_MDD$cond <- paste(annotation_MDD$Sample, annotation_MDD$col, sep=":")
head(annotation_MDD)
annot_MDD <- t(annotation_MDD[,5])
write.table(annot_MDD, file = "ModuleViewer/Ens_Anno_Sex_MDD", quote = F, sep = '|', na = "", row.names = F,
            col.names = F)
```


# Regulators of modules
```{r}
km_reg_AD <- read.delim("kmclust_reg_AD")
head(km_reg_AD)
colnames(km_reg_AD) <- c("Cluster", "genes_clust", "Regulators")
km_reg_AD <- km_reg_AD %>% arrange(Cluster)
dim(km_reg_AD)
km_reg_AD[(km_reg_AD$Regulators == ""), ] #only large cluster 28
km_reg_AD <- km_reg_AD %>% slice(-28)
dim(km_reg_AD)
```

```{r}
km_reg_MDD <- read.delim("kmclust_reg_MDD")
head(km_reg_MDD)
colnames(km_reg_MDD) <- c("Cluster", "genes_clust", "Regulators")
km_reg_MDD <- km_reg_MDD %>% arrange(Cluster)
dim(km_reg_MDD)
km_reg_MDD[(km_reg_MDD$Regulators == ""), ] #only large cluster 5
km_reg_MDD <- km_reg_MDD %>% slice(-5)
dim(km_reg_MDD)
```

Split regulators into different columns 
```{r}
km_reg_AD[paste0("Regulator", c(1:11))] <- str_split_fixed(km_reg_AD$Regulators, '_', 11)
head(km_reg_AD)
km_reg_AD <- km_reg_AD %>% select(-Regulator11) # 10 regulators but last _ gone
```

Make new data frame with one regulator and cluster 
```{r}
new_reg_AD1 <- data.frame("Cluster" = km_reg_AD$Cluster, "Regulator" = km_reg_AD$Regulator1)
head(new_reg_AD1)
new_reg_AD2 <- data.frame("Cluster" = km_reg_AD$Cluster, "Regulator" = km_reg_AD$Regulator2)
new_reg_AD3 <- data.frame("Cluster" = km_reg_AD$Cluster, "Regulator" = km_reg_AD$Regulator3)
new_reg_AD4 <- data.frame("Cluster" = km_reg_AD$Cluster, "Regulator" = km_reg_AD$Regulator4)
new_reg_AD5 <- data.frame("Cluster" = km_reg_AD$Cluster, "Regulator" = km_reg_AD$Regulator5)
new_reg_AD6 <- data.frame("Cluster" = km_reg_AD$Cluster, "Regulator" = km_reg_AD$Regulator6)
new_reg_AD7 <- data.frame("Cluster" = km_reg_AD$Cluster, "Regulator" = km_reg_AD$Regulator7)
new_reg_AD8 <- data.frame("Cluster" = km_reg_AD$Cluster, "Regulator" = km_reg_AD$Regulator8)
new_reg_AD9 <- data.frame("Cluster" = km_reg_AD$Cluster, "Regulator" = km_reg_AD$Regulator9)
new_reg_AD10 <- data.frame("Cluster" = km_reg_AD$Cluster, "Regulator" = km_reg_AD$Regulator10)
new_reg_AD <- rbind(new_reg_AD1, new_reg_AD2, new_reg_AD3, new_reg_AD4, new_reg_AD5,
                    new_reg_AD6, new_reg_AD7, new_reg_AD8, new_reg_AD9, new_reg_AD10)
head(new_reg_AD)
dim(new_reg_AD)
new_reg_AD <- new_reg_AD[!(new_reg_AD$Regulator == ""), ]
dim(new_reg_AD)
new_reg_AD <- new_reg_AD %>% arrange(Cluster)
head(new_reg_AD)
new_reg_AD <- new_reg_AD %>% select(Regulator, Cluster)
write.table(new_reg_AD, file = "reg_clust_ens_AD.txt", sep = "\t", quote = F, row.names = F, col.names = F)
```
MDD 
```{r}
km_reg_MDD[paste0("Regulator", c(1:11))] <- str_split_fixed(km_reg_MDD$Regulators, '_', 11)
head(km_reg_MDD)
km_reg_MDD <- km_reg_MDD %>% select(-Regulator11)
```
 
```{r}
new_reg_MDD1 <- data.frame("Cluster" = km_reg_MDD$Cluster, "Regulator" = km_reg_MDD$Regulator1)
head(new_reg_MDD1)
new_reg_MDD2 <- data.frame("Cluster" = km_reg_MDD$Cluster, "Regulator" = km_reg_MDD$Regulator2)
new_reg_MDD3 <- data.frame("Cluster" = km_reg_MDD$Cluster, "Regulator" = km_reg_MDD$Regulator3)
new_reg_MDD4 <- data.frame("Cluster" = km_reg_MDD$Cluster, "Regulator" = km_reg_MDD$Regulator4)
new_reg_MDD5 <- data.frame("Cluster" = km_reg_MDD$Cluster, "Regulator" = km_reg_MDD$Regulator5)
new_reg_MDD6 <- data.frame("Cluster" = km_reg_MDD$Cluster, "Regulator" = km_reg_MDD$Regulator6)
new_reg_MDD7 <- data.frame("Cluster" = km_reg_MDD$Cluster, "Regulator" = km_reg_MDD$Regulator7)
new_reg_MDD8 <- data.frame("Cluster" = km_reg_MDD$Cluster, "Regulator" = km_reg_MDD$Regulator8)
new_reg_MDD9 <- data.frame("Cluster" = km_reg_MDD$Cluster, "Regulator" = km_reg_MDD$Regulator9)
new_reg_MDD10 <- data.frame("Cluster" = km_reg_MDD$Cluster, "Regulator" = km_reg_MDD$Regulator10)
new_reg_MDD <- rbind(new_reg_MDD1, new_reg_MDD2, new_reg_MDD3, new_reg_MDD4, new_reg_MDD5,
                    new_reg_MDD6, new_reg_MDD7, new_reg_MDD8, new_reg_MDD9, new_reg_MDD10)
head(new_reg_MDD)
dim(new_reg_MDD)
new_reg_MDD <- new_reg_MDD[!(new_reg_MDD$Regulator == ""), ]
dim(new_reg_MDD)
new_reg_MDD <- new_reg_MDD %>% arrange(Cluster)
head(new_reg_MDD, 10)
new_reg_MDD <- new_reg_MDD %>% select(Regulator, Cluster)
write.table(new_reg_MDD, file = "reg_clust_ens_MDD.txt", sep = "\t", quote = F, row.names = F, col.names = F)
```

