---
title: "Rank aggregation"
author: "Hanne Puype"
date: '2022-03-23'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

# Rank aggregation with Borda
https://cran.r-project.org/web/packages/TopKLists/vignettes/TopKLists.pdf  
https://cran.r-project.org/web/packages/TopKLists/TopKLists.pdf  

## Packages
```{r}
#install.packages("TopKLists")
library(TopKLists)
library(tidyverse)
```

## Prepare networks
```{r}
edgesAD1 <- read.table("edgesAD1.txt", sep = ",", header = F)
edgesAD2 <- read.table("edgesAD2.txt", sep = ",", header = F)
edgesAD3 <- read.table("edgesAD3.txt", sep = ",", header = F)
edgesMDD1 <- read.table("edgesMDD1.txt", sep = ",", header = F)
edgesMDD2 <- read.table("edgesMDD2.txt", sep = ",", header = F)
edgesMDD3 <- read.table("edgesMDD3.txt", sep = ",", header = F)
head(edgesAD1)
```

```{r}
edgesAD1 <- edgesAD1$V1
edgesAD2 <- edgesAD2$V1
edgesAD3 <- edgesAD3$V1
edgesMDD1 <- edgesMDD1$V1
edgesMDD2 <- edgesMDD2$V1
edgesMDD3 <- edgesMDD3$V1
```

```{r}
list_edges_AD <- list(as.list(edgesAD1), as.list(edgesAD2), as.list(edgesAD3))
summary(list_edges_AD)
list_edges_MDD <- list(as.list(edgesMDD1), as.list(edgesMDD2), as.list(edgesMDD3))
```

## Borda
### AD
```{r}
rankaggr_AD <- Borda(list_edges_AD, k = 100000)
head(rankaggr_AD$TopK)
dim(rankaggr_AD$TopK)
head(rankaggr_AD$Scores)
dim(rankaggr_AD$Scores)
tail(rankaggr_AD$Scores)
```

#### Back to edge list
```{r}
rank_aggr_AD_mean <- rankaggr_AD$TopK[,"mean"]
head(rank_aggr_AD_mean)
rank_aggr_AD_mean <- as.data.frame(rank_aggr_AD_mean)
rank_aggr_AD_mean[c('TF', 'Target_gene')] <- str_split_fixed(rank_aggr_AD_mean$rank_aggr_AD_mean, ':', 2)
rank_aggr_AD_mean <- rownames_to_column(rank_aggr_AD_mean) #add rank
names(rank_aggr_AD_mean)[colnames(rank_aggr_AD_mean) == "rowname"] <- "Rank"
rank_aggr_AD_mean <- rank_aggr_AD_mean %>% select(TF, Target_gene, Rank)
head(rank_aggr_AD_mean)
rank_aggr_AD_mean <- rank_aggr_AD_mean %>% add_column("Score" = rankaggr_AD$Scores[c(1:100000), "mean"])
```

### MDD
```{r}
rankaggr_MDD <- Borda(list_edges_MDD, k = 100000)
head(rankaggr_MDD$TopK)
dim(rankaggr_MDD$TopK)
head(rankaggr_MDD$Scores)
tail(rankaggr_MDD$Scores)
```

#### Back to edge list
```{r}
rank_aggr_MDD_mean <- rankaggr_MDD$TopK[,"mean"]
head(rank_aggr_MDD_mean)
rank_aggr_MDD_mean <- as.data.frame(rank_aggr_MDD_mean)
rank_aggr_MDD_mean[c('TF', 'Target_gene')] <- str_split_fixed(rank_aggr_MDD_mean$rank_aggr_MDD_mean, ':', 2)
rank_aggr_MDD_mean <- rownames_to_column(rank_aggr_MDD_mean)
names(rank_aggr_MDD_mean)[colnames(rank_aggr_MDD_mean) == "rowname"] <- "Rank"
rank_aggr_MDD_mean <- rank_aggr_MDD_mean %>% select(TF, Target_gene, Rank)
rank_aggr_MDD_mean <- rank_aggr_MDD_mean %>% add_column("Score" = rankaggr_MDD$Scores[c(1:100000), "mean"])
head(rank_aggr_MDD_mean)
```

## Save files
```{r}
write.table(rank_aggr_AD_mean, file = "ensemble_netw_AD.txt", sep = "\t", col.names = T, row.names = F, 
            quote = F)
write.table(rank_aggr_MDD_mean, file = "ensemble_netw_MDD.txt", sep = "\t", col.names = T, row.names = F, 
            quote = F)
```

```{r}
Borda.plot(rankaggr_AD, k = 100)
Borda.plot(rankaggr_MDD, k = 100)
```

# Overlap between ensemble network and initial networks
Will also see then if one method has bigger influence than the others or not
```{r}
edges_ens_AD <- rankaggr_AD$TopK[,"mean"]
head(edges_ens_AD)
edges_ens_MDD <- rankaggr_MDD$TopK[, "mean"]
```

```{r}
overlap_G3AD <- Reduce(intersect,list(edgesAD1,edges_ens_AD)) # GENIE3
length(overlap_G3AD) #53029
overlap_CLRAD <- Reduce(intersect,list(edgesAD2,edges_ens_AD)) # CLR
length(overlap_CLRAD) #47513
overlap_LTAD <- Reduce(intersect,list(edgesAD3,edges_ens_AD)) # Lemon-Tree
length(overlap_LTAD) #48105
```

```{r}
overlap_G3MDD <- Reduce(intersect,list(edgesMDD1,edges_ens_MDD)) # GENIE3
length(overlap_G3MDD) #58810
overlap_CLRMDD <- Reduce(intersect,list(edgesMDD2,edges_ens_MDD)) # CLR
length(overlap_CLRMDD) #54358
overlap_LTMDD <- Reduce(intersect,list(edgesMDD3,edges_ens_MDD)) # Lemon-Tree
length(overlap_LTMDD) #49779
```

# Number of nodes and TFs left in the ensemble network
```{r}
RegulatorAD <- rank_aggr_AD_mean$TF
RegulatorAD <- unique(RegulatorAD)
length(RegulatorAD) #1041 regulators
Target_gene <- rank_aggr_AD_mean$Target_gene
Target_gene <- unique(Target_gene)
length(Target_gene) #9675 target genes 
nodesAD <- c(RegulatorAD, Target_gene)
nodesAD <- unique(nodesAD)
length(nodesAD) #9685 nodes
```

```{r}
RegulatorMDD <- rank_aggr_MDD_mean$TF
RegulatorMDD <- unique(RegulatorMDD)
length(RegulatorMDD) #1101 regulators
Target_gene <- rank_aggr_MDD_mean$Target_gene
Target_gene <- unique(Target_gene)
length(Target_gene) #9006 target genes 
nodesMDD <- c(RegulatorMDD, Target_gene)
nodesMDD <- unique(nodesMDD)
length(nodesMDD) #9021 nodes
```

overlap of nodes between AD and MDD
```{r}
length(Reduce(intersect, list(nodesAD, nodesMDD)))
```

for edges: see netw_char_ensemble
