---
title: "HumanNet PPI"
author: "Hanne Puype"
date: "2023-03-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Read in files 
CX co-expression
DB pathway database
DP protein domain
GI genetic interaction
GN gene neighborhood
```{r}
HN_PI <- read.table("HumanNet/HS-PI.tsv")
HN_CX <- read.table("HumanNet/HS-CX.tsv")
HN_DB <- read.table("HumanNet/HS-DB.tsv")
HN_DP <- read.table("HumanNet/HS-DP.tsv")
HN_GI <- read.table("HumanNet/HS-GI.tsv")
HN_GN <- read.table("HumanNet/HS-GN.tsv")
modulesAD <- read.table("kmed_modules_AD.txt", header = F, sep = "\t")
modulesMDD <- read.table("kmed_modules_MDD.txt", header = F, sep = "\t")
colnames(modulesAD) <- c("Gene", "Module")
colnames(modulesMDD) <- c("Gene", "Module")
```

```{r}
HN_funct <- rbind(HN_DB, HN_DP, HN_GI, HN_GN)
```

The genes are in Entrez gene ids

# Convert genes to gene symbols
```{r}
library(tidyverse)
library(biomaRt)
library(annotate)
library(org.Hs.eg.db)
library(EnsDb.Hsapiens.v86)
first_col <- HN_PI$V1
first_col <- as.character(first_col)
geneIDs1 <- ensembldb::select(EnsDb.Hsapiens.v86, keys = first_col, keytype = "ENTREZID", 
                              columns = c("ENTREZID","SYMBOL"))

sec_col <- HN_PI$V2
sec_col <- as.character(sec_col)
geneIDs2 <- ensembldb::select(EnsDb.Hsapiens.v86, keys = sec_col, keytype = "ENTREZID", 
                              columns = c("ENTREZID","SYMBOL"))

HN_PI_temp <- merge(HN_PI, geneIDs1, by.x = "V1", by.y = "ENTREZID")
HN_PI_temp <- unique(HN_PI_temp)
colnames(HN_PI_temp)[names(HN_PI_temp) == "SYMBOL"] <- "Gene1"
HN_PI_temp %>% dplyr::filter(Gene1 == "")

HN_PI_temp <- merge(HN_PI_temp, geneIDs2, by.x = "V2", by.y = "ENTREZID")
HN_PI_temp <- unique(HN_PI_temp)
colnames(HN_PI_temp)[names(HN_PI_temp) == "SYMBOL"] <- "Gene2"
HN_PI_temp %>% dplyr::filter(Gene2 == "")
HN_PI <- HN_PI_temp %>% dplyr::select(Gene1, Gene2)
```

## Add co-functional interactions (co-expression)
```{r}
first_col <- HN_CX$V1
first_col <- as.character(first_col)
geneIDs1 <- ensembldb::select(EnsDb.Hsapiens.v86, keys = first_col, keytype = "ENTREZID", 
                              columns = c("ENTREZID","SYMBOL"))
sec_col <- HN_CX$V2
sec_col <- as.character(sec_col)
geneIDs2 <- ensembldb::select(EnsDb.Hsapiens.v86, keys = sec_col, keytype = "ENTREZID", 
                              columns = c("ENTREZID","SYMBOL"))
HN_CX_temp <- merge(HN_CX, geneIDs1, by.x = "V1", by.y = "ENTREZID")
HN_CX_temp <- unique(HN_CX_temp)
colnames(HN_CX_temp)[names(HN_CX_temp) == "SYMBOL"] <- "Gene1"
HN_CX_temp %>% dplyr::filter(Gene1 == "")

HN_CX_temp <- merge(HN_CX_temp, geneIDs2, by.x = "V2", by.y = "ENTREZID")
HN_CX_temp <- unique(HN_CX_temp)
colnames(HN_CX_temp)[names(HN_CX_temp) == "SYMBOL"] <- "Gene2"
HN_CX_temp %>% dplyr::filter(Gene2 == "")
HN_CX <- HN_CX_temp %>% dplyr::select(Gene1, Gene2)
```


Other functional interactions
```{r}
first_col <- HN_funct$V1
first_col <- as.character(first_col)
geneIDs1 <- ensembldb::select(EnsDb.Hsapiens.v86, keys = first_col, keytype = "ENTREZID", 
                              columns = c("ENTREZID","SYMBOL"))

sec_col <- HN_funct$V2
sec_col <- as.character(sec_col)
geneIDs2 <- ensembldb::select(EnsDb.Hsapiens.v86, keys = sec_col, keytype = "ENTREZID", 
                              columns = c("ENTREZID","SYMBOL"))

HN_funct_temp <- merge(HN_funct, geneIDs1, by.x = "V1", by.y = "ENTREZID")
HN_funct_temp <- unique(HN_funct_temp)
colnames(HN_funct_temp)[names(HN_funct_temp) == "SYMBOL"] <- "Gene1"
HN_funct_temp %>% dplyr::filter(Gene1 == "")

HN_funct_temp <- merge(HN_funct_temp, geneIDs2, by.x = "V2", by.y = "ENTREZID")
HN_funct_temp <- unique(HN_funct_temp)
colnames(HN_funct_temp)[names(HN_funct_temp) == "SYMBOL"] <- "Gene2"
HN_funct_temp %>% dplyr::filter(Gene2 == "")
HN_funct <- HN_funct_temp %>% dplyr::select(Gene1, Gene2)
```


## Automated way for all modules 
dataframe with all modules and their respective genes
```{r}
B <- modulesAD %>% group_by(Module) %>% mutate(rid = row_number()) %>% spread(Module, Gene)
for (i in 2:length(names(B))){
  names(B)[i] <- paste0('module', names(B)[i])
}
B_MDD <- modulesMDD %>% group_by(Module) %>% mutate(rid = row_number()) %>% spread(Module, Gene)
for (i in 2:length(names(B_MDD))){
  names(B_MDD)[i] <- paste0('module', names(B_MDD)[i])
}
colnames(B)
colnames(B_MDD)
```

### AD 
PPI
```{r}
B <- B %>% dplyr::select(-rid)
PI_AD <- data.frame(Gene1 = "", Gene2 = "", Module = "")
for (i in (1:length(B))){
  module_of_interest <- B[, i]
  genes_module <-  module_of_interest[,1] %>% unlist() # remains tibble otherwise
  PI_module <- HN_PI %>% filter(Gene1 %in% genes_module) %>% 
                                  filter(Gene2 %in% genes_module) %>%
                                  add_column("Module" = i)
  PI_AD <- rbind(PI_AD, PI_module)
}
PI_AD <- PI_AD %>% dplyr::select(Module, Gene1, Gene2)
PI_AD <- PI_AD %>% dplyr::slice(-1)
write.table(PI_AD, file = "Ens_HumanNet_PI_AD.mvf", row.names = F, col.names = F, sep = "\t", quote = F)
```
ready for ModuleViewer, only need to add header

CX
```{r}
CX_AD <- data.frame(Gene1 = "", Gene2 = "", Module = "")
for (i in (1:length(B))){
  module_of_interest <- B[, i]
  genes_module <-  module_of_interest[,1] %>% unlist() # remains tibble otherwise
  CX_module <- HN_CX %>% filter(Gene1 %in% genes_module) %>% 
                                  filter(Gene2 %in% genes_module) %>%
                                  add_column("Module" = i)
  CX_AD <- rbind(CX_AD, CX_module)
}
CX_AD <- CX_AD %>% dplyr::select(Module, Gene1, Gene2)
CX_AD <- CX_AD %>% dplyr::slice(-1)
write.table(CX_AD, file = "Ens_HumanNet_CX_AD.mvf", row.names = F, col.names = F, sep = "\t", quote = F)
```


Other functional interactions
```{r}
funct <- data.frame(Gene1 = "", Gene2 = "", Module = "")
for (i in (1:length(B))){
  module_of_interest <- B[, i]
  genes_module <-  module_of_interest[,1] %>% unlist() # remains tibble otherwise
  func_module <- HN_funct %>% dplyr::filter(Gene1 %in% genes_module) %>% 
                                  dplyr::filter(Gene2 %in% genes_module) %>%
                                  add_column("Module" = i)
  funct <- rbind(funct, func_module)
}
funct <- funct %>% dplyr::select(Module, Gene1, Gene2)
funct <- funct %>% dplyr::slice(-1)
write.table(funct, file = "Ens_HumanNet_Funct_AD.mvf", row.names = F, col.names = F, sep = "\t", quote = F)
```

### MDD
PPI
```{r}
B_MDD <- B_MDD %>% dplyr::select(-rid)
PI_MDD <- data.frame(Gene1 = "", Gene2 = "", Module = "")
for (i in (1:length(B_MDD))){
  module_of_interest <- B_MDD[, i]
  genes_module <-  module_of_interest[,1] %>% unlist() # remains tibble otherwise
  PI_module <- HN_PI %>% filter(Gene1 %in% genes_module) %>% 
                                  filter(Gene2 %in% genes_module) %>%
                                  add_column("Module" = i)
  PI_MDD <- rbind(PI_MDD, PI_module)
}
PI_MDD <- PI_MDD %>% dplyr::select(Module, Gene1, Gene2)
PI_MDD <- PI_MDD %>% dplyr::slice(-1)
write.table(PI_MDD, file = "Ens_HumanNet_PI_MDD.mvf", row.names = F, col.names = F, sep = "\t", quote = F)
```

CX
```{r}
CX_MDD <- data.frame(Gene1 = "", Gene2 = "", Module = "")
for (i in (1:length(B_MDD))){
  module_of_interest <- B_MDD[, i]
  genes_module <-  module_of_interest[,1] %>% unlist() # remains tibble otherwise
  CX_module <- HN_CX %>% filter(Gene1 %in% genes_module) %>% 
                                  filter(Gene2 %in% genes_module) %>%
                                  add_column("Module" = i)
  CX_MDD <- rbind(CX_MDD, CX_module)
}
CX_MDD <- CX_MDD %>% dplyr::select(Module, Gene1, Gene2)
CX_MDD <- CX_MDD %>% dplyr::slice(-1)
write.table(CX_MDD, file = "Ens_HumanNet_CX_MDD.mvf", row.names = F, col.names = F, sep = "\t", quote = F)
```

Other functional interactions
```{r}
funct_MDD <- data.frame(Gene1 = "", Gene2 = "", Module = "")
for (i in (1:length(B_MDD))){
  module_of_interest <- B_MDD[, i]
  genes_module <-  module_of_interest[,1] %>% unlist() # remains tibble otherwise
  func_module <- HN_funct %>% dplyr::filter(Gene1 %in% genes_module) %>% 
                                  dplyr::filter(Gene2 %in% genes_module) %>%
                                  add_column("Module" = i)
  funct_MDD <- rbind(funct_MDD, func_module)
}
funct_MDD <- funct_MDD %>% dplyr::select(Module, Gene1, Gene2)
funct_MDD <- funct_MDD %>% dplyr::slice(-1)
write.table(funct_MDD, file = "Ens_HumanNet_Funct_MDD.mvf", row.names = F, col.names = F, sep = "\t", quote = F)
```

# Add PPI and co-expression to TFs
```{r}
regAD <- read.table("reg_clust_ens_AD.txt", header = F, sep = "\t")
regMDD <- read.table("reg_clust_ens_MDD.txt", header = F, sep = "\t")
colnames(regAD) <- c("TF", "Module")
colnames(regMDD) <- c("TF", "Module")
```

```{r}
B_reg_AD <- regAD %>% group_by(Module) %>% mutate(rid = row_number()) %>% spread(Module, TF)
for (i in 2:length(names(B_reg_AD))){
  names(B_reg_AD)[i] <- paste0('module', names(B_reg_AD)[i])
}
B_reg_MDD <- regMDD %>% group_by(Module) %>% mutate(rid = row_number()) %>% spread(Module, TF)
for (i in 2:length(names(B_reg_MDD))){
  names(B_reg_MDD)[i] <- paste0('module', names(B_reg_MDD)[i])
}
colnames(B_reg_AD) # 28 is out
B_reg_AD <- B_reg_AD %>% add_column("module28" = "", .after = "module27")
colnames(B_reg_AD)
colnames(B_reg_MDD) # module 5 is out
B_reg_MDD <- B_reg_MDD %>% add_column("module5" = "", .after = "module4")
```

## PPI
```{r}
B_reg_AD <- B_reg_AD %>% dplyr::select(-rid)
PI <- data.frame(Gene1 = "", Gene2 = "", Module = "")
for (i in (1:length(B_reg_AD))){
  module_of_interest <- B_reg_AD[, i]
  genes_module <-  module_of_interest[,1] %>% unlist() # remains tibble otherwise
  PI_module <- HN_PI %>% dplyr::filter(Gene1 %in% genes_module) %>% 
                                  dplyr::filter(Gene2 %in% genes_module) %>%
                                  add_column("Module" = i)
  PI <- rbind(PI, PI_module)
}
PI <- PI %>% dplyr::select(Module, Gene1, Gene2)
PI <- PI %>% dplyr::slice(-1)
PI <- unique(PI)
write.table(PI, file = "Ens_HumanNet_PI_reg_AD.mvf", row.names = F, col.names = F, sep = "\t", quote = F)
```

```{r}
B_reg_MDD <- B_reg_MDD %>% dplyr::select(-rid)
PI <- data.frame(Gene1 = "", Gene2 = "", Module = "")
for (i in (1:length(B_reg_MDD))){
  module_of_interest <- B_reg_MDD[, i]
  genes_module <-  module_of_interest[,1] %>% unlist() # remains tibble otherwise
  PI_module <- HN_PI %>% dplyr::filter(Gene1 %in% genes_module) %>% 
                                  dplyr::filter(Gene2 %in% genes_module) %>%
                                  add_column("Module" = i)
  PI <- rbind(PI, PI_module)
}
PI <- PI %>% dplyr::select(Module, Gene1, Gene2)
PI <- PI %>% dplyr::slice(-1)
PI <- unique(PI)
write.table(PI, file = "Ens_HumanNet_PI_reg_MDD.mvf", row.names = F, col.names = F, sep = "\t", quote = F)
```

## Co-expression 

```{r}
CX_AD <- data.frame(Gene1 = "", Gene2 = "", Module = "")
for (i in (1:length(B_reg_AD))){
  module_of_interest <- B_reg_AD[, i]
  genes_module <-  module_of_interest[,1] %>% unlist() # remains tibble otherwise
  CX_module <- HN_CX %>% dplyr::filter(Gene1 %in% genes_module) %>% 
                                  dplyr::filter(Gene2 %in% genes_module) %>%
                                  add_column("Module" = i)
  CX_AD <- rbind(CX_AD, CX_module)
}
CX_AD <- CX_AD %>% dplyr::select(Module, Gene1, Gene2)
CX_AD <- CX_AD %>% dplyr::slice(-1)
write.table(CX_AD, file = "Ens_HumanNet_CX_reg_AD.mvf", row.names = F, col.names = F, sep = "\t", quote = F)
```

```{r}
CX_MDD <- data.frame(Gene1 = "", Gene2 = "", Module = "")
for (i in (1:length(B_reg_MDD))){
  module_of_interest <- B_reg_MDD[, i]
  genes_module <-  module_of_interest[,1] %>% unlist() # remains tibble otherwise
  CX_module <- HN_CX %>% dplyr::filter(Gene1 %in% genes_module) %>% 
                                  dplyr::filter(Gene2 %in% genes_module) %>%
                                  add_column("Module" = i)
  CX_MDD <- rbind(CX_MDD, CX_module)
}
CX_MDD <- CX_MDD %>% dplyr::select(Module, Gene1, Gene2)
CX_MDD <- CX_MDD %>% dplyr::slice(-1)
write.table(CX_MDD, file = "Ens_HumanNet_CX_reg_MDD.mvf", row.names = F, col.names = F, sep = "\t", quote = F)
```


## Other functional interactions
```{r}
funct_AD <- data.frame(Gene1 = "", Gene2 = "", Module = "")
for (i in (1:length(B_reg_AD))){
  module_of_interest <- B_reg_AD[, i]
  genes_module <-  module_of_interest[,1] %>% unlist() # remains tibble otherwise
  func_module <- HN_funct %>% dplyr::filter(Gene1 %in% genes_module) %>% 
                                  dplyr::filter(Gene2 %in% genes_module) %>%
                                  add_column("Module" = i)
  funct_AD <- rbind(funct_AD, func_module)
}
funct_AD <- funct_AD %>% dplyr::select(Module, Gene1, Gene2)
funct_AD <- funct_AD %>% dplyr::slice(-1)
write.table(funct_AD, file = "Ens_HumanNet_funct_reg_AD.mvf", row.names = F, col.names = F, sep = "\t", quote = F)
```


```{r}
funct_MDD <- data.frame(Gene1 = "", Gene2 = "", Module = "")
for (i in (1:length(B_reg_MDD))){
  module_of_interest <- B_reg_MDD[, i]
  genes_module <-  module_of_interest[,1] %>% unlist() # remains tibble otherwise
  func_module <- HN_funct %>% dplyr::filter(Gene1 %in% genes_module) %>% 
                                  dplyr::filter(Gene2 %in% genes_module) %>%
                                  add_column("Module" = i)
  funct_MDD <- rbind(funct_MDD, func_module)
}
funct_MDD <- funct_MDD %>% dplyr::select(Module, Gene1, Gene2)
funct_MDD <- funct_MDD %>% dplyr::slice(-1)
write.table(funct_MDD, file = "Ens_HumanNet_funct_reg_MDD.mvf", row.names = F, col.names = F, sep = "\t", quote = F)
```

