---
title: "DisGeNET"
author: "Hanne Puype"
date: "2023-06-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Packages
```{r}
library(tidyverse)
library(enrichR)
```

# Load datasets
```{r}
modulesAD <- read.table("kmed_modules_AD.txt", header = F, sep = "\t")
modulesMDD <- read.table("kmed_modules_MDD.txt", header = F, sep = "\t")
colnames(modulesAD) <- c("Gene", "Module")
colnames(modulesMDD) <- c("Gene", "Module")
modulesAD <- as_tibble(modulesAD)
modulesMDD <- as_tibble(modulesMDD)

B <- modulesAD %>% group_by(Module) %>% mutate(rid = row_number()) %>% spread(Module, Gene)
for (i in 2:length(names(B))){
  names(B)[i] <- paste0('module', names(B)[i])
}
B_MDD <- modulesMDD %>% group_by(Module) %>% mutate(rid = row_number()) %>% spread(Module, Gene)
for (i in 2:length(names(B_MDD))){
  names(B_MDD)[i] <- paste0('module', names(B_MDD)[i])
}
```

# Enrichment analysis of modules in DisGeNET
```{r}
dbs <- c("DisGeNET")
```

AD 
```{r eval=FALSE}
modules <- paste0("module", c(1:155))
modules_OI_AD <- c()
for (module in modules){
  modIds <- B[module]
  modIds <- as_tibble(modIds)
  names(modIds) <- "GeneSymbol"
  symbols <- modIds$GeneSymbol
  enriched <- enrichr(symbols, dbs)
  enriched <- lapply(enriched, function(x) x <- x %>% filter(Adjusted.P.value <= 0.05))
  if (dim(as.data.frame(enriched[[1]]) %>% filter(grepl("Alz", Term, ignore.case = T)))[1] != 0) {
      modules_OI_AD <- c(modules_OI_AD, module)
}
  lapply(enriched, function(x) write.table(data.frame(x),
                                      paste0('DisGeNET/Enrichment_',module,'.txt'),
                                           append = T, sep='\t', quote = F))
}
```

```{r}
print(modules_OI_AD, quote = F)
```

```{r}
modules <- paste0("module", c(1:156))
modules_OI_MDD <- c()
for (module in modules){
  modIds <- B_MDD[module]
  modIds <- as_tibble(modIds)
  names(modIds) <- "GeneSymbol"
  symbols <- modIds$GeneSymbol
  enriched <- enrichr(symbols, dbs)
  enriched <- lapply(enriched, function(x) x <- x %>% filter(Adjusted.P.value <= 0.05))
  if (dim(as.data.frame(enriched[[1]]) %>% filter(grepl("depr", Term, ignore.case = T)))[1] != 0) {
      modules_OI_MDD <- c(modules_OI_MDD, module)
  }
  lapply(enriched, function(x) write.table(data.frame(x),
                                    paste0('DisGeNET/EnrichmentMDD_',module,'.txt'),
                                           append = T, sep='\t', quote = F, col.names = T))
}
```

```{r}
print(modules_OI_MDD, quote = F)
```

# DisGeNET enrichment in target genes of TFs
```{r}
network_AD <- read.delim("./ensemble_netw_AD.txt")
network_MDD <- read.delim("./ensemble_netw_MDD.txt") 
network_AD <- network_AD %>% select(TF, Target_gene)
network_MDD <- network_MDD %>% select(TF, Target_gene)
B_AD <- network_AD %>% group_by(TF) %>% mutate(rid = row_number()) %>% spread(TF, Target_gene)
B_AD <- B_AD %>% select(-rid)
B_MDD <- network_MDD %>% group_by(TF) %>% mutate(rid = row_number()) %>% spread(TF, Target_gene)
B_MDD <- B_MDD %>% select(-rid)
```

```{r}
TFs_OI_AD_network <- c()
for (regulator in colnames(B_AD)){
  modIds <- B_AD[regulator]
  modIds <- as_tibble(modIds)
  names(modIds) <- "GeneSymbol"
  symbols <- modIds$GeneSymbol
  enriched <- enrichr(symbols, dbs)
  enriched <- lapply(enriched, function(x) x <- x %>% filter(Adjusted.P.value <= 0.05))
  if (dim(as.data.frame(enriched[[1]]) %>% filter(grepl("Alz", Term, ignore.case = T)))[1] != 0) {
      TFs_OI_AD_network <- c(TFs_OI_AD_network, regulator)
}
  lapply(enriched, function(x) write.table(data.frame(x),
                                      paste0('DisGeNET_network/Enrichment_',regulator,'.txt'),
                                           append = T, sep='\t', quote = F))
}
```

```{r}
print(TFs_OI_AD_network, quote = F) # CEBPD, FOS, FOSB, IRF8, MYC, NFATC2, TFEC
```

```{r}
TFs_OI_MDD_network <- c()
for (regulator in colnames(B_MDD)){
  modIds <- B_MDD[regulator]
  modIds <- as_tibble(modIds)
  names(modIds) <- "GeneSymbol"
  symbols <- modIds$GeneSymbol
  enriched <- enrichr(symbols, dbs)
  enriched <- lapply(enriched, function(x) x <- x %>% filter(Adjusted.P.value <= 0.05))
  if (dim(as.data.frame(enriched[[1]]) %>% filter(grepl("depr", Term, ignore.case = T)))[1] != 0) {
      TFs_OI_MDD_network <- c(TFs_OI_MDD_network, regulator)
}
  lapply(enriched, function(x) write.table(data.frame(x),
                                      paste0('DisGeNET_network/EnrichmentMDD_',regulator,'.txt'),
                                           append = T, sep='\t', quote = F))
}
```

```{r}
print(TFs_OI_MDD_network, quote = F) # FOS, FOSB, JUNB, MAFF, RUNX1, STAT3
```

CREB1, ESR2, ESRRA are bipolar depression, so will not include this  
KLF12, KLF13 have Involutional Depression, will also not include  
KLF15, TP53 Depressed nasal ridge  
FLI1, FOS only have depressive symptoms  
MSX1 only has depressed mood  
STAT3 only has Mixed anxiety and depressive disorder  
ZIC1 only Depressive episode, unspecified  

# DisGeNET enrichment in regulons 
```{r}
network_AD_sc <- read.table("single-cell/edgelist_sc_AD.txt", sep = "\t", header = T)
network_MDD_sc <- read.table("single-cell/edgelist_sc_MDD.txt", sep = "\t", header = T)
```

```{r}
B_sc <- network_AD_sc %>% group_by(regulator) %>% mutate(rid = row_number()) %>% spread(regulator, gene)
B_sc <- B_sc %>% select(-rid, -value)
B_sc_MDD <- network_MDD_sc %>% group_by(regulator) %>% mutate(rid = row_number()) %>% spread(regulator, gene)
B_sc_MDD <- B_sc_MDD %>% select(-rid, -value)
```

```{r}
regulons_OI_AD_sc <- c()
for (regulator in colnames(B_sc)){
  modIds <- B_sc[regulator]
  modIds <- as_tibble(modIds)
  names(modIds) <- "GeneSymbol"
  symbols <- modIds$GeneSymbol
  enriched <- enrichr(symbols, dbs)
  enriched <- lapply(enriched, function(x) x <- x %>% filter(Adjusted.P.value <= 0.05))
  if (dim(as.data.frame(enriched[[1]]) %>% filter(grepl("Alz", Term, ignore.case = T)))[1] != 0) {
      regulons_OI_AD_sc <- c(regulons_OI_AD_sc, regulator)
}
  lapply(enriched, function(x) write.table(data.frame(x),
                                      paste0('single-cell/DisGeNET/Enrichment_',regulator,'.txt'),
                                           append = T, sep='\t', quote = F))
}
```

```{r}
print(regulons_OI_AD_sc, quote = F) # FOS, IKZF1, IRF8, JUN, MAF, MAFF, NFATC2, RUNX1, SPI1, STAT3, TAL1
```
ESR1 only has Dementia due to Alzheimer's disease (disorder)	 
MYC, ZBTB7B only ALZHEIMER DISEASE 5  

```{r}
regulons_OI_MDD_sc <- c()
for (regulator in colnames(B_sc_MDD)){
  modIds <- B_sc_MDD[regulator]
  modIds <- as_tibble(modIds)
  names(modIds) <- "GeneSymbol"
  symbols <- modIds$GeneSymbol
  enriched <- enrichr(symbols, dbs)
  enriched <- lapply(enriched, function(x) x <- x %>% filter(Adjusted.P.value <= 0.05))
  if (dim(as.data.frame(enriched[[1]]) %>% filter(grepl("depr", Term, ignore.case = T)))[1] != 0) {
      regulons_OI_MDD_sc <- c(regulons_OI_MDD_sc, regulator)
}
  lapply(enriched, function(x) write.table(data.frame(x),
                                      paste0('single-cell/DisGeNET/EnrichmentMDD_',regulator,'.txt'),
                                           append = T, sep='\t', quote = F))
}
```

```{r}
print(regulons_OI_MDD_sc, quote = F) # CEBPD, FOS, FOSB, IRF8, JUN, JUNB, PAX6, RUNX1
```

IRF1 only Mixed anxiety and depressive disorder  
OVOL2, ZNF43 only Recurrent major depressive episodes  
SOX4 only bipolar disorder
USF2 endogenous depression, depressive syndrome
ZSCAN5A only Recurrent depression

