---
title: "Analysis modules"
author: "Hanne Puype"
date: "2022-10-25"
output:  
  html_document:
    toc: TRUE
    toc_float: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

# Load in modules 
```{r}
modulesAD <- read.table("kmed_modules_AD.txt", header = F, sep = "\t")
modulesMDD <- read.table("kmed_modules_MDD.txt", header = F, sep = "\t")
colnames(modulesAD) <- c("Gene", "Module")
colnames(modulesMDD) <- c("Gene", "Module")
library(tidyverse)
modulesAD <- as_tibble(modulesAD)
modulesMDD <- as_tibble(modulesMDD)
```

# Number of genes and names

```{r}
df = data.frame()

for (i in 1:155) {
    Module <- modulesAD %>% filter(Module == i)
    output = length(Module$Gene)
    df = rbind(df, output)
    comma_vec <- paste(Module$Gene, collapse = ", ")
    df[i, 2] <- comma_vec
}
head(df)
write.table(df, file = "length_Mod_AD.tsv", quote = F, row.names = F, sep = "\t")
```

```{r}
df = data.frame()

for (i in 1:156) {
    Module <- modulesMDD %>% filter(Module == i)
    output = length(Module$Gene)
    df = rbind(df, output)
    comma_vec <- paste(Module$Gene, collapse = ", ")
    df[i, 2] <- comma_vec
}
head(df)
write.table(df, file = "length_Mod_MDD.tsv", quote = F, row.names = F, sep = "\t")
```

# Regulators
```{r}
reg_AD <- read.table("reg_clust_ens_AD.txt")
head(reg_AD)
reg_MDD <- read.table("reg_clust_ens_MDD.txt")
head(reg_MDD)
```

```{r}
df = data.frame()
for (i in 1:155) {
    Reg <- reg_AD %>% filter(V2 == i)
    output = length(Reg$V1)
    df = rbind(df, output)
    comma_vec <- paste(Reg$V1, collapse = ", ")
    df[i, 2] <- comma_vec
}
head(df)
write.table(df, file = "reg_Mod_AD.tsv", quote = F, row.names = F, sep = "\t")
```

```{r}
df = data.frame()
for (i in 1:156) {
    Reg <- reg_MDD %>% filter(V2 == i)
    output = length(Reg$V1)
    df = rbind(df, output)
    comma_vec <- paste(Reg$V1, collapse = ", ")
    df[i, 2] <- comma_vec
}
tail(df)
write.table(df, file = "reg_Mod_MDD.tsv", quote = F, row.names = F, sep = "\t")
```

# Known regulatory interactions 
load networks 
```{r}
gri <- read.delim("Omnipath.txt")
dbTF <- read.csv("dbTF.csv", sep="")
gri <- gri %>% filter(source_genesymbol %in% dbTF$Gene_symbol)
dim(gri)
frontal_lobe_adult <- read.delim("networks/FANTOM5_individual_networks/394_individual_networks/frontal_lobe_adult.txt.gz", header=FALSE)
frontal_lobe_adult <- frontal_lobe_adult %>% filter(V1 %in% dbTF$Gene_symbol)
frontal_lobe_adult <- frontal_lobe_adult %>% arrange(desc(V3))
```

Load ensemble networks 
```{r}
network_AD <- read.table("ensemble_netw_AD.txt", header = T, sep = "\t")
head(network_AD)
network_MDD <- read.table("ensemble_netw_MDD.txt", header = T, sep = "\t")
```

Format for intersection
```{r}
network_AD$concat <- paste(network_AD$TF, network_AD$Target_gene, sep = ":")
network_MDD$concat <- paste(network_MDD$TF, network_MDD$Target_gene, sep = ":")
edgesAD <- network_AD$concat
edgesMDD <- network_MDD$concat
edges_OmniPath <- gri %>% dplyr::select(source_genesymbol, target_genesymbol)
edges_OmniPath$concat <- paste(edges_OmniPath$source_genesymbol, edges_OmniPath$target_genesymbol, sep = ":")
edges_OP <- edges_OmniPath$concat
frontal_lobe_adult$concat <- paste(frontal_lobe_adult$V1, frontal_lobe_adult$V2, sep = ":")
edges_Marbach <- frontal_lobe_adult$concat
```

```{r}
frontal_lobe_adult_2 <- frontal_lobe_adult %>% slice(1:100000)
dim(frontal_lobe_adult_2)
tail(frontal_lobe_adult_2)
edges_Marbach <- frontal_lobe_adult_2$concat
```

Intersection
```{r}
length(intersect(edgesAD, edges_OP))
length(intersect(edgesMDD, edges_OP))
length(intersect(edgesAD, edges_Marbach))
length(intersect(edgesMDD, edges_Marbach))
```

```{r}
length(intersect(edges_OP, edges_Marbach))
```


OmniPath + Marbach 
```{r}
length(c(edges_OP, edges_Marbach)[duplicated(c(edges_OP, edges_Marbach))]) #4290 in common
edges_GT <- unique(c(edges_OP, edges_Marbach))
length(edges_GT)
```

# Overlap ensemble, single-cell and ground truth 
```{r}
network_sc_AD <- read.table("Single-cell/edgelist_sc_AD.txt", sep = "\t", header = T)
network_sc_MDD <- read.table("Single-cell/edgelist_sc_MDD.txt", sep = "\t", header = T)
network_sc_AD$concat <- paste(network_sc_AD$regulator, network_sc_AD$gene, sep = ":")
network_sc_MDD$concat <- paste(network_sc_MDD$regulator, network_sc_MDD$gene, sep = ":")
edges_sc_AD <- network_sc_AD$concat
edges_sc_MDD <- network_sc_MDD$concat
```

Put intersections into dataframe and split again into TF and target gene
```{r}
intersect_AD_OP <- intersect(edgesAD, edges_OP)
intersect_MDD_OP <- intersect(edgesMDD, edges_OP)
intersect_AD_OP2 <- data.frame(intersect_AD_OP)
intersect_MDD_OP2 <- data.frame(intersect_MDD_OP)
head(intersect_AD_OP2)
intersect_AD_OP2[c('TF', 'TG')] <- str_split_fixed(intersect_AD_OP2$intersect_AD_OP, ':', 2)
intersect_AD_OP2 <- intersect_AD_OP2 %>% select(-intersect_AD_OP)
intersect_MDD_OP2[c('TF', 'TG')] <- str_split_fixed(intersect_MDD_OP2$intersect_MDD_OP, ':', 2)
intersect_MDD_OP2 <- intersect_MDD_OP2 %>% select(-intersect_MDD_OP)
head(intersect_MDD_OP2)
```

```{r}
intersect_AD_Ma <- intersect(edgesAD, edges_Marbach)
intersect_MDD_Ma <- intersect(edgesMDD, edges_Marbach)
intersect_AD_Ma2 <- data.frame(intersect_AD_Ma)
intersect_MDD_Ma2 <- data.frame(intersect_MDD_Ma)
intersect_AD_Ma2[c('TF', 'TG')] <- str_split_fixed(intersect_AD_Ma2$intersect_AD_Ma, ':', 2)
intersect_AD_Ma2 <- intersect_AD_Ma2 %>% select(-intersect_AD_Ma)
intersect_MDD_Ma2[c('TF', 'TG')] <- str_split_fixed(intersect_MDD_Ma2$intersect_MDD_Ma, ':', 2)
intersect_MDD_Ma2 <- intersect_MDD_Ma2 %>% select(-intersect_MDD_Ma)
```


## Known interactions in modules 
TG in which module? --> also regulator inferred?  
Add number of module to df, order according to module and look 
```{r}
head(modulesAD)
head(intersect_AD_OP2)
intersect_AD <- left_join(intersect_AD_OP2, modulesAD, by = c("TG" = "Gene"))
head(intersect_AD)
intersect_AD <- intersect_AD %>% arrange(Module)

head(intersect_MDD_OP2)
intersect_MDD <- left_join(intersect_MDD_OP2, modulesMDD, by = c("TG" = "Gene"))
head(intersect_MDD)
intersect_MDD <- intersect_MDD %>% arrange(Module)
```

```{r}
write.table(intersect_AD, file = "intersect_AD_OP.txt", sep = "\t", row.names = F, quote = F, col.names = T)
write.table(intersect_MDD, file = "intersect_MDD_OP.txt", sep = "\t", row.names = F, quote = F, col.names = T)
```

See excel file Overview modules  
Module 22 from AD: also TAL1-CSF3R  
Module 39 from AD: TAL1-CXCL16; SPI1-ITGAL; SPI1-PTPN6  

```{r}
intersect_AD <- left_join(intersect_AD_Ma2, modulesAD, by = c("TG" = "Gene"))
intersect_AD <- intersect_AD %>% arrange(Module)
head(intersect_AD)
intersect_MDD <- left_join(intersect_MDD_Ma2, modulesMDD, by = c("TG" = "Gene"))
head(intersect_MDD)
intersect_MDD <- intersect_MDD %>% arrange(Module)
```

In format to paste into excel file  
Regulators must also be inferred for modules 
```{r}
reg_clust_AD <- read.delim("C:/UGent/Master 2/Master's dissertation/R/reg_clust_ens_AD.txt", header=FALSE)
reg_clust_MDD <- read.delim("C:/UGent/Master 2/Master's dissertation/R/reg_clust_ens_MDD.txt", header=FALSE)
head(reg_clust_AD)
intersect_AD1 <- merge(intersect_AD, reg_clust_AD, by.x = "TF", by.y = "V1")
head(intersect_AD1)
intersect_AD1 <- intersect_AD1 %>% arrange(Module)
intersect_AD1$new_col <- ifelse(intersect_AD1$Module == intersect_AD1$V2, 'A',
                ifelse(intersect_AD1$Module != intersect_AD1$V2, 'B', "C"))
intersect_AD1 <- intersect_AD1 %>% filter(new_col == "A") %>% select(TF, TG, Module)
intersect_AD1$concat <- paste(intersect_AD1$TF, intersect_AD1$TG, sep = "-")
dim(intersect_AD1)
intersect_AD1 <- intersect_AD1 %>% select(concat, Module)
write.table(intersect_AD1, file = "intersect_AD_Ma_mod.txt", sep = "\t", row.names = F, quote = F, col.names = T)
```


```{r}
intersect_MDD1 <- merge(intersect_MDD, reg_clust_MDD, by.x = "TF", by.y = "V1")
head(intersect_MDD1)
intersect_MDD1 <- intersect_MDD1 %>% arrange(Module)
intersect_MDD1$new_col <- ifelse(intersect_MDD1$Module == intersect_MDD1$V2, 'A',
                ifelse(intersect_MDD1$Module != intersect_MDD1$V2, 'B', "C"))
intersect_MDD1 <- intersect_MDD1 %>% filter(new_col == "A") %>% select(TF, TG, Module)
intersect_MDD1$concat <- paste(intersect_MDD1$TF, intersect_MDD1$TG, sep = "-")
head(intersect_MDD1)
dim(intersect_MDD1)
intersect_MDD1 <- intersect_MDD1 %>% select(concat, Module)
write.table(intersect_MDD1, file = "intersect_MDD_Ma_mod.txt", sep = "\t", row.names = F, quote = F, col.names = T)
```


## Network figure with significant modules
And with all immune modules 
```{r}
head(reg_AD)
reg_AD_sign <- reg_AD %>% filter(V2 == 21 | V2 == 35 | V2 == 26 | V2 == 40 | V2 == 58 | V2 == 60 | V2 == 66 | 
                                   V2 == 85 | V2 == 87 | V2 == 105 | V2 == 123 | V2 == 129 | V2 == 130 | V2 == 150 |
                                   V2 == 153| V2 == 22 | V2 == 4 | V2 == 9 | V2 == 36 | V2 == 39 | V2 == 51 | 
                                   V2 == 86 | 
                                   V2 == 93 | V2 == 139 | V2 == 141)
dim(reg_AD_sign)
reg_MDD_sign <- reg_MDD %>% filter(V2 == 3 | V2 == 12 | V2 == 24 | V2 == 36 | V2 == 44 | V2 == 48 | V2 == 50 | 
                                     V2 == 49 | V2 == 93 | V2 == 110 | V2 == 115 | V2 == 4 | V2 == 37 | V2 == 56 | 
                                     V2 == 59 | V2 == 61 | V2 == 63 | V2 == 81 | V2 == 147)
reg_AD_sign %>% group_by(V2) %>% summarise(n())
reg_MDD_sign %>% group_by(V2) %>% summarise(n())
write.csv(reg_AD_sign, "reg_AD_sign.csv", quote = F, row.names = F)
write.csv(reg_MDD_sign, "reg_MDD_sign.csv", quote = F, row.names = F)
```

# Overlap of prioritized modules 
```{r}
modulesAD_prior <- modulesAD %>% filter(Module == 21 | Module == 35 | Module == 26 | Module == 40 | Module == 58 |
                                    Module == 60 | Module == 66 | Module == 85 | Module == 87 | Module == 105 |
                                    Module == 123 | Module == 129 | Module == 130 | Module == 150 |
                                    Module == 153| Module == 22 | Module == 4 | Module == 9 | Module == 36 | 
                                    Module == 39 | Module == 51 | Module == 86 | 
                                   Module == 93 | Module == 139 | Module == 141)
modulesMDD_prior <- modulesMDD %>% filter(Module == 3 | Module == 12 | Module == 24 | Module == 36 | Module == 44 |
                                            Module == 48 |Module == 50 |  Module == 49 | Module == 93 | 
                                            Module == 110 | Module == 115 | Module == 4 | Module == 37 | 
                                            Module == 56 | Module == 59 | Module == 61 | Module == 63 | 
                                            Module == 81 | Module == 147)
```

```{r}
intersection_modules <- c()
name_intersec <- c()
for (module in unique(modulesAD_prior$Module)) {
  modulesAD_module <- modulesAD_prior %>% filter(Module == module)
  for (module2 in unique(modulesMDD_prior$Module)) {
    modulesMDD_module <- modulesMDD_prior %>% filter(Module == module2)
    intersection <- length(intersect(modulesAD_module$Gene, modulesMDD_module$Gene))
    intersection_modules <- append(intersection_modules, intersection)
    name_intersec <- append(name_intersec, paste0("Mod_AD", module, "-", "Mod_MDD", module2))
  }
}
df <- data.frame(intersection_modules, name_intersec)
# for every module in AD and for every module in MDD 
# save the overlap of the number of genes
# save also the name of the overlap
```


```{r}
df <- df %>% arrange(desc(intersection_modules)) %>% select(name_intersec, intersection_modules)
write.csv(df, file = "overlap_prioritized_modules.csv", col.names = T, row.names = F, quote = F)
```

## TFs
```{r}
intersection_modules <- c()
name_intersec <- c()
for (module in unique(reg_AD_sign$V2)) {
  modulesAD_regs <- reg_AD_sign %>% filter(V2 == module)
  for (module2 in unique(reg_MDD_sign$V2)) {
    modulesMDD_regs <- reg_MDD_sign %>% filter(V2 == module2)
    intersection <- length(intersect(modulesAD_regs$V1, modulesMDD_regs$V1))
    intersection_modules <- append(intersection_modules, intersection)
    name_intersec <- append(name_intersec, paste0("Mod_AD", module, "-", "Mod_MDD", module2))
  }
}
df <- data.frame(intersection_modules, name_intersec)
```

```{r}
df <- df %>% arrange(desc(intersection_modules)) %>% select(name_intersec, intersection_modules)
write.csv(df, file = "overlap_prioritized_modules_regs.csv", col.names = T, row.names = F, quote = F)
```


