---
title: "SCENIC CSI"
author: "Hanne Puype"
date: "2023-03-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

# Install/load packages and load files 
```{r}
#library(devtools)
#install_github("FloWuenne/scFunctions")
#library(scFunctions)
library(tidyverse)
library(pheatmap)
#install.packages("viridis")
library(viridis)
```

```{r}
regulonAUC_AD <- read.csv2("regulonAUC_AD.csv", header = T)
regulonAUC_AD <- as.data.frame(regulonAUC_AD)
regulonAUC_AD <- column_to_rownames(regulonAUC_AD, "X")
regulonAUC_MDD <- read.csv2("regulonAUC_MDD.csv")
regulonAUC_MDD <- column_to_rownames(regulonAUC_MDD, "X")
```

# Connection specificity index 
https://github.com/FloWuenne/scFunctions/blob/master/Tutorials/process_SCENIC.md 
Functions from above, was not able to install package
```{r}
calculate_csi <- function(regulonAUC,
                          calc_extended = FALSE,
                          verbose = FALSE){

  compare_pcc <- function(vector_of_pcc,pcc){
    pcc_larger <- length(vector_of_pcc[vector_of_pcc > pcc])
    if(pcc_larger == length(vector_of_pcc)){
      return(0)
    }else{
      return(length(vector_of_pcc))
    }
  }

  calc_csi <- function(reg,reg2,pearson_cor){
    test_cor <- pearson_cor[reg,reg2]
    total_n <- ncol(pearson_cor)
    pearson_cor_sub <- subset(pearson_cor,rownames(pearson_cor) == reg | rownames(pearson_cor) == reg2)

    sums <- apply(pearson_cor_sub,MARGIN = 2, FUN = compare_pcc, pcc = test_cor)
    fraction_lower <- length(sums[sums == nrow(pearson_cor_sub)]) / total_n
    return(fraction_lower)
  }

  regulonAUC_sub <- regulonAUC

  if(calc_extended == TRUE){
    regulonAUC_sub <- subset(regulonAUC_sub,grepl("extended",rownames(regulonAUC_sub)))
  } else if (calc_extended == FALSE){
    regulonAUC_sub <- subset(regulonAUC_sub,!grepl("extended",rownames(regulonAUC_sub)))
}

  regulonAUC_sub <- t(regulonAUC_sub)

  pearson_cor <- cor(regulonAUC_sub)
  pearson_cor_df <- as.data.frame(pearson_cor)
  pearson_cor_df$regulon_1 <- rownames(pearson_cor_df)
  pearson_cor_long <- pearson_cor_df %>%
    gather(regulon_2,pcc,-regulon_1) %>%
    mutate("regulon_pair" = paste(regulon_1,regulon_2,sep="_"))


  regulon_names <- unique(colnames(pearson_cor))
  num_of_calculations <- length(regulon_names)*length(regulon_names)

  csi_regulons <- data.frame(matrix(nrow=num_of_calculations,ncol = 3))

  colnames(csi_regulons) <- c("regulon_1",
                              "regulon_2",
                              "CSI")

  num_regulons <- length(regulon_names)

  f <- 0
  for(reg in regulon_names){
    ## Check if user wants to print info
    if(verbose == TRUE){
      print(reg)
      }
    for(reg2 in regulon_names){
      f <- f + 1

      fraction_lower <- calc_csi(reg,reg2,pearson_cor)

      csi_regulons[f,] <- c(reg,reg2,fraction_lower)

    }
  }
  csi_regulons$CSI <- as.numeric(csi_regulons$CSI)
  return(csi_regulons)
}
```

```{r}
plot_csi_modules <- function(csi_df,
                             nclust = 10,
                             font_size_regulons = 6){

  ## subset csi data frame based on threshold
  csi_test_mat <- csi_df %>%
    spread(regulon_2,CSI)

  future_rownames <- csi_test_mat$regulon_1
  csi_test_mat <- as.matrix(csi_test_mat[,2:ncol(csi_test_mat)])
  rownames(csi_test_mat) <- future_rownames
  pdf("./csi_heatmap15_AD_w.pdf",
       width = 15,
       height = 20)
  pheatmap(csi_test_mat,
           show_colnames = FALSE,
           color = viridis(n = 20),
           cutree_cols = 15,
           cutree_rows = 15,
           fontsize_row = font_size_regulons,
           cluster_cols = TRUE,
           cluster_rows = TRUE,
           treeheight_row = nclust,
           treeheight_col = nclust,
           clustering_distance_rows = "euclidean",
           clustering_distance_cols = "euclidean",
           width = 2000,
           height = 6200)
  dev.off()

}
```

The final statistics that we want to calculate is the connection specificty index. The CSI is a measure of connectedness between the different regulons. Regulons that share high CSI likely are co-regulating downstream genes and are together responsible for cell function.  
We can calculate the CSI scores for all regulon pairs based on the AUCs matrix for all regulons. Again, this function has a switch to either select high confidence regulons or run the CSI calculation only on _extended regulons (calc_extended = TRUE). Here we choose to calculate CSI only for high confidence regulons.
```{r}
regulons_csi <- calculate_csi(regulonAUC_AD,
                              calc_extended = FALSE)
```

Once we have calculated the CSI values for all regulon pairs, we can visualize the regulon modules using the function plot_csi_modules, which will create a heatmap of the CSI values. The function has an argument that lets you change the number of clusters the heatmap is divided into via nclust.
```{r fig.height=15}
plot_csi_modules(regulons_csi,
                 nclust = 15,
                 font_size_regulons = 4)
```

## MDD
```{r}
regulons_csi_MDD <- calculate_csi(regulonAUC_MDD,
                              calc_extended = FALSE)
```

```{r}
plot_csi_modules(regulons_csi_MDD,
                 nclust = 20,
                 font_size_regulons = 4)
```

# Save clusters in file for UMAPs
## AD
```{r}
csi_csi_wide <- regulons_csi %>%
    spread(regulon_2,CSI)

future_rownames <- csi_csi_wide$regulon_1
csi_csi_wide <- as.matrix(csi_csi_wide[,2:ncol(csi_csi_wide)])
rownames(csi_csi_wide) <- future_rownames
  
regulons_hclust <- hclust(dist(csi_csi_wide, method = "euclidean"))

clusters <- cutree(regulons_hclust,k= 15)
clusters_df <- data.frame("regulon" = names(clusters),
                          "csi_cluster" = clusters)
write.table(clusters_df, file = "regulonclusters15_AD.txt", col.names = T, row.names = F, quote = F, sep = "\t")
```

## MDD

```{r}
csi_csi_wide <- regulons_csi_MDD %>%
    spread(regulon_2,CSI)

future_rownames <- csi_csi_wide$regulon_1
csi_csi_wide <- as.matrix(csi_csi_wide[,2:ncol(csi_csi_wide)])
rownames(csi_csi_wide) <- future_rownames
  
regulons_hclust <- hclust(dist(csi_csi_wide, method = "euclidean"))

clusters <- cutree(regulons_hclust, k= 15)
clusters_df <- data.frame("regulon" = names(clusters),
                          "csi_cluster" = clusters)
write.table(clusters_df, file = "regulonclusters15_MDD.txt", col.names = T, row.names = F, quote = F, sep = "\t")
```

