---
title: "Overlap edges"
author: "Hanne Puype"
date: "11 december 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

# Format of edges to compare 

## GENIE3
```{r}
library(tidyverse)
linklistAD <- read.table("LinkList_AD_top.csv", header = T, sep = ",")
head(linklistAD)
#linklistAD <- linklistAD %>% dplyr::slice(c(1:100000))
linklistMDD <- read.table("LinkList_MDD_top.csv", header = T, sep = ",")
head(linklistMDD)
#linklistMDD <- linklistMDD %>% slice(c(1:100000))
linklistAD$concat <- paste(linklistAD$regulatoryGene, linklistAD$targetGene, sep = ":")
edgesAD1 <- linklistAD$concat
linklistMDD$concat <- paste(linklistMDD$regulatoryGene, linklistMDD$targetGene, sep = ":")
edgesMDD1 <- linklistMDD$concat
```

## CLR
```{r}
networkCLR_AD <- read.table("network_AD_reg.csv", header = T, sep = ",")
head(networkCLR_AD)
#networkCLR_AD <- networkCLR_AD %>% slice(c(1:100000))
networkCLR_MDD <- read.table("network_MDD_reg.csv", header = T, sep = ",")
head(networkCLR_MDD)
#networkCLR_MDD <- networkCLR_MDD %>% slice(c(1:100000))
networkCLR_AD$concat <- paste(networkCLR_AD$gene1, networkCLR_AD$gene2, sep = ":")
edgesAD2 <- networkCLR_AD$concat
networkCLR_MDD$concat <- paste(networkCLR_MDD$gene1, networkCLR_MDD$gene2, sep = ":")
edgesMDD2 <- networkCLR_MDD$concat
```

## Lemon-Tree
```{r}
edgelist_AD <- read.table("Lemon-Tree/EdgeListLT3_AD.csv", sep = ",", header = T)
head(edgelist_AD)
edgelist_MDD <- read.table("Lemon-Tree/EdgeListLT3_MDD.csv", sep = ",", header = T)
head(edgelist_MDD)
dim(edgelist_AD)
edgelist_AD$concat <- paste(edgelist_AD$Regulator, edgelist_AD$Target_gene, sep = ":")
edgesAD3 <- edgelist_AD$concat
edgelist_MDD$concat <- paste(edgelist_MDD$Regulator, edgelist_MDD$Target_gene, sep = ":")
edgesMDD3 <- edgelist_MDD$concat
```

## Save files
```{r}
write.table(edgesAD1, file = "edgesAD1.txt", sep = ",", col.names = F, row.names = F, quote = F)
write.table(edgesAD2, file = "edgesAD2.txt", sep = ",", col.names = F, row.names = F, quote = F)
write.table(edgesAD3, file = "edgesAD3.txt", sep = ",", col.names = F, row.names = F, quote = F)
write.table(edgesMDD1, file = "edgesMDD1.txt", sep = ",", col.names = F, row.names = F, quote = F)
write.table(edgesMDD2, file = "edgesMDD2.txt", sep = ",", col.names = F, row.names = F, quote = F)
write.table(edgesMDD3, file = "edgesMDD3.txt", sep = ",", col.names = F, row.names = F, quote = F)
```

# Venn diagram  
AD
```{r}
#install.packages("BioVenn")
library(BioVenn)
venn_diag <- draw.venn(edgesAD1, edgesAD2, edgesAD3, title = "Venn diagram of edges", nrtype = "abs", xtitle =
                         "GENIE3", ytitle = "CLR", ztitle = "  Lemon-Tree", subtitle = "AD", output = "png", 
                       filename = "VennDiagram100_AD.png", t_s = 5, st_s = 4, xt_s = 4, yt_s = 4, zt_s = 4, 
                       nr_s = 3, x_c = "salmon", y_c = "seagreen3", z_c = "royalblue1")
```
MDD
```{r}
venn_diag <- draw.venn(edgesMDD1, edgesMDD2, edgesMDD3, title = "Venn diagram of edges", nrtype = "abs", xtitle =
                         "  GENIE3", ytitle = "\nCLR", ztitle = "    Lemon-Tree", subtitle = "MDD", output = "png", 
                       filename = "VennDiagram100_MDD.png", t_s = 5, st_s = 4, xt_s = 4, yt_s = 4, zt_s = 4, 
                       nr_s = 3, x_c = "salmon", y_c = "seagreen3", z_c = "royalblue1")
```

# Number of nodes
Make barplots per disease of number of target genes, number of regulators and total number of nodes

## GENIE3
put genes into list, take unique elements and number of elements  
Have selected top 100 000 as there are only so much edges in Lemon-Tree network
```{r}
linklistAD <- read.table("LinkList_AD_top.csv", header = T)
head(linklistAD)
linklistMDD <- read.table("LinkList_MDD_top.csv", header = T)
head(linklistMDD)
#AD
dim(linklistAD)
linklistAD <- linklistAD %>% dplyr::slice(c(1:100000))
regulatoryGeneAD <- linklistAD$regulatoryGene
head(regulatoryGeneAD)
regulatoryGeneAD <- unique(regulatoryGeneAD)
length(regulatoryGeneAD) #994 regulators
targetGeneAD <- linklistAD$targetGene
targetGeneAD <- unique(targetGeneAD)
length(targetGeneAD) #9340 genes 
nodesAD1 <- c(regulatoryGeneAD, targetGeneAD)
length(nodesAD1)
nodesAD1 <- unique(nodesAD1)
length(nodesAD1) #9384 nodes in total
#MDD 
linklistMDD <- linklistMDD %>% slice(c(1:100000))
regulatoryGeneMDD <- linklistMDD$regulatoryGene
head(regulatoryGeneMDD)
regulatoryGeneMDD <- unique(regulatoryGeneMDD)
length(regulatoryGeneMDD) #1069 regulators
targetGeneMDD <- linklistMDD$targetGene
targetGeneMDD <- unique(targetGeneMDD)
length(targetGeneMDD) #8859 genes 
nodesMDD1 <- c(regulatoryGeneMDD, targetGeneMDD)
length(nodesMDD1)
nodesMDD1 <- unique(nodesMDD1)
length(nodesMDD1) #8871 nodes in total
```

## CLR
```{r}
networkCLR_AD <- read.table("network_AD_reg.csv", header = T)
head(networkCLR_AD)
dim(networkCLR_AD)
networkCLR_MDD <- read.table("network_MDD_reg.csv", header = T)
head(networkCLR_MDD)
#AD
networkCLR_AD <- networkCLR_AD %>% slice(c(1:100000))
gene1AD <- networkCLR_AD$gene1
head(gene1AD)
gene1AD <- unique(gene1AD)
length(gene1AD) #1041 regulators
gene2AD <- networkCLR_AD$gene2
gene2AD <- unique(gene2AD)
length(gene2AD) #9714 target genes
nodesAD2 <- c(gene1AD, gene2AD)
length(nodesAD2)
nodesAD2 <- unique(nodesAD2)
length(nodesAD2) #9715 nodes 
#MDD
networkCLR_MDD <- networkCLR_MDD %>% slice(c(1:100000))
gene1MDD <- networkCLR_MDD$gene1
head(gene1MDD)
gene1MDD <- unique(gene1MDD)
length(gene1MDD) #1101 regulators
gene2MDD <- networkCLR_MDD$gene2
gene2MDD <- unique(gene2MDD)
length(gene2MDD) #9182 target genes
nodesMDD2 <- c(gene1MDD, gene2MDD)
length(nodesMDD2)
nodesMDD2 <- unique(nodesMDD2)
length(nodesMDD2) #9183 nodes 
```

## Lemon-Tree
```{r}
edgelist_AD <- read.table("Lemon-Tree/EdgeListLT3_AD.csv", sep = ",", header = T)
head(edgelist_AD)
dim(edgelist_AD)
edgelist_MDD <- read.table("Lemon-Tree/EdgeListLT3_MDD.csv", sep = ",", header = T)
head(edgelist_MDD)
dim(edgelist_MDD)
#AD
RegulatorAD <- edgelist_AD$Regulator
RegulatorAD <- unique(RegulatorAD)
length(RegulatorAD) #618 regulators
Target_gene <- edgelist_AD$Target_gene
Target_gene <- unique(Target_gene)
length(Target_gene) #8674 target genes 
nodesAD3 <- c(RegulatorAD, Target_gene)
nodesAD3 <- unique(nodesAD3)
length(nodesAD3) #8736 nodes
#MDD
RegulatorMDD <- edgelist_MDD$Regulator
RegulatorMDD <- unique(RegulatorMDD)
length(RegulatorMDD) #763 regulators
Target_geneMDD <- edgelist_MDD$Target_gene
Target_geneMDD <- unique(Target_geneMDD)
length(Target_geneMDD) #7934 target genes 
nodesMDD3 <- c(RegulatorMDD, Target_geneMDD)
nodesMDD3 <- unique(nodesMDD3)
length(nodesMDD3) #8012 nodes
```

## Bar plots
Number of nodes
```{r}
# Nodes in AD
nodesAD <- data.frame(method = c("GENIE3", "CLR", "Lemon-Tree", "MERLIN-P"),
                      nodes = c(9384, 9715, 8736, 1))
head(nodesAD)
library(tidyverse)
ggplot(data = nodesAD, aes(x = method, y = nodes)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  geom_text(aes(label = nodes), vjust = 1.6, color = "white", size = 3.5) +
  labs(title = "Number of nodes for AD networks") +
  theme_minimal()

#nodes in MDD
nodesMDD <- data.frame(method = c("GENIE3", "CLR", "Lemon-Tree", "MERLIN-P"),
                       nodes = c(8871, 9183, 8011, 1))
ggplot(data = nodesMDD, aes(x = method, y = nodes)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  geom_text(aes(label = nodes), vjust = 1.6, color = "white", size = 3.5) +
  labs(title = "Number of nodes for MDD networks") +
  theme_minimal()

nodes <- data.frame(disease = rep(c("AD", "MDD"), each = 3), 
                    method = rep(c("GENIE3", "CLR", "Lemon-Tree"), 2),
                    nodes = c(9384, 9715, 8736, 8871, 9183, 8011))
head(nodes)
ggplot(data = nodes, aes(x = method, y = nodes, fill = disease)) +
  geom_bar(stat = "identity", position=position_dodge()) +
  geom_text(aes(label = nodes), vjust = 1.6, color = "white", size = 3.5, position = position_dodge(0.9)) +
  labs(title = "Number of nodes for each method per disease") +
  scale_fill_manual(values = c("lightcoral", "royalblue")) + #indianred1
  theme_minimal()
```

```{r}
png(filename = "nodes.png")
ggplot(data = nodes, aes(x = method, y = nodes, fill = disease)) +
  geom_bar(stat = "identity", position=position_dodge()) +
  geom_text(aes(label = nodes), vjust = 1.6, color = "white", size = 3.5, position = position_dodge(0.9)) +
  labs(title = "Number of nodes for each method per disease") +
  scale_fill_manual(values = c("lightcoral", "royalblue")) +
  theme_minimal()
dev.off()
```

# Hubs in networks
## GENIE3
```{r}
library(tidyverse)
linklistAD <- read.table("LinkList_AD_top.csv", header = T, sep = ",")
head(linklistAD)
tail(linklistAD)
#linklistAD <- linklistAD %>% dplyr::slice(c(1:100000))
linklistMDD <- read.table("LinkList_MDD_top.csv", header = T, sep = ",")
head(linklistMDD)
#linklistMDD <- linklistMDD %>% slice(c(1:100000))
```

frequency of nodes
```{r}
nodesall_AD1 <- c(linklistAD$regulatoryGene, linklistAD$targetGene)
head(nodesall_AD1)
nodesnr_AD <- as.data.frame(table(nodesall_AD1))
nodesnr_AD <- nodesnr_AD %>% arrange(desc(Freq))
head(nodesnr_AD)
```
Frequency is number of times these nodes are either a regulator or target gene, so number of times it is in an edge

frequency of regulators
```{r}
regAD1 <- linklistAD$regulatoryGene 
regAD1 <- as.data.frame(table(regAD1))
regAD1 <- regAD1 %>% arrange(desc(Freq))
head(regAD1)
regAD1$regAD1 <- as.character(regAD1$regAD1)
```
All top nodes are also regulators, so I will only look at the regulators

```{r}
regMDD1 <- linklistMDD$regulatoryGene 
regMDD1 <- as.data.frame(table(regMDD1))
regMDD1 <- regMDD1 %>% arrange(desc(Freq))
regMDD1$regMDD1 <- as.character(regMDD1$regMDD1)
head(regMDD1)
```

## CLR
```{r}
networkCLR_AD <- read.table("network_AD_reg.csv", header = T, sep = ",")
head(networkCLR_AD)
dim(networkCLR_AD)
#networkCLR_AD <- networkCLR_AD %>% slice(c(1:100000))
networkCLR_MDD <- read.table("network_MDD_reg.csv", header = T, sep = ",")
head(networkCLR_MDD)
#networkCLR_MDD <- networkCLR_MDD %>% slice(c(1:100000))
```

Frequency of regulators
```{r}
regAD2 <- networkCLR_AD$gene1
regAD2 <- as.data.frame(table(regAD2))
regAD2 <- regAD2 %>% arrange(desc(Freq))
regAD2$regAD2 <- as.character(regAD2$regAD2)
head(regAD2)
```
MDD
```{r}
regMDD2 <- networkCLR_MDD$gene1
regMDD2 <- as.data.frame(table(regMDD2))
regMDD2 <- regMDD2 %>% arrange(desc(Freq))
regMDD2$regMDD2 <- as.character(regMDD2$regMDD2)
head(regMDD2)
```

## Lemon-Tree
```{r}
edgelist_AD <- read.table("Lemon-Tree/EdgeListLT3_AD.csv", sep = ",", header = T)
head(edgelist_AD)
edgelist_MDD <- read.table("Lemon-Tree/EdgeListLT3_MDD.csv", sep = ",", header = T)
head(edgelist_MDD)
```

```{r}
regAD3 <- edgelist_AD$Regulator
regAD3 <- as.data.frame(table(regAD3))
regAD3 <- regAD3 %>% arrange(desc(Freq))
regAD3$regAD3 <- as.character(regAD3$regAD3)
head(regAD3)
```

```{r}
regMDD3 <- edgelist_MDD$Regulator
regMDD3 <- as.data.frame(table(regMDD3))
regMDD3 <- regMDD3 %>% arrange(desc(Freq))
regMDD3$regMDD3 <- as.character(regMDD3$regMDD3)
head(regMDD3)
```
MEF2C always returns 

## Venn diagram of top 100 regulators
```{r}
regAD1 <- regAD1 %>% slice(c(1:100)) %>% select(regAD1)
regAD2 <- regAD2 %>% slice(c(1:100)) %>% select(regAD2)
regAD3 <- regAD3 %>% slice(c(1:100)) %>% select(regAD3)
regMDD1 <- regMDD1 %>% slice(c(1:100)) %>% select(regMDD1)
regMDD2 <- regMDD2 %>% slice(c(1:100)) %>% select(regMDD2)
regMDD3 <- regMDD3 %>% slice(c(1:100)) %>% select(regMDD3)
head(regAD3)
regAD1 <- regAD1$regAD1
regAD2 <- regAD2$regAD2
regAD3 <- regAD3$regAD3
regMDD1 <- regMDD1$regMDD1
regMDD2 <- regMDD2$regMDD2
regMDD3 <- regMDD3$regMDD3
```

```{r}
library(BioVenn)
venn_diag <- draw.venn(regAD1, regAD2, regAD3, title = "Venn diagram of top 100 regulators", nrtype = "abs", xtitle =
                         "GENIE3", ytitle = "\n\nCLR", ztitle = "\n\n        Lemon-Tree", subtitle = "AD", output = "png", 
                       filename = "VennReg_AD.png", t_s = 5, st_s = 4, xt_s = 4, yt_s = 4, zt_s = 4, 
                       nr_s = 3.5, x_c = "salmon", y_c = "seagreen3", z_c = "royalblue1")
```

```{r}
venn_diag <- draw.venn(regMDD1, regMDD2, regMDD3, title = "Venn diagram of top 100 regulators", nrtype = "abs", 
                       xtitle = "GENIE3", ytitle = "\n\nCLR", ztitle = "\n\n        Lemon-Tree", subtitle = "MDD", 
                       output = "png", filename = "VennReg_MDD.png", t_s = 5, st_s = 4, xt_s = 4, yt_s = 4, zt_s = 4,
                       nr_s = 3.5, x_c = "salmon", y_c = "seagreen3", z_c = "royalblue1")
```

## Shared regulators between AD and MDD
```{r}
regAD <- c(regAD1, regAD2, regAD3)
regAD <- unique(regAD)
length(regAD)
regMDD <- c(regMDD1, regMDD2, regMDD3)
regMDD <- unique(regMDD)
length(regMDD)
reg <- NULL
venn_diag <- draw.venn(reg, regAD, regMDD, title = "Venn diagram of top regulators", nrtype = "abs", 
                       ytitle = "\n\nAD", ztitle = "\n\n   MDD", subtitle = "", output = "png", 
                       filename = "VennReg_AD_MDD.png", t_s = 5, st_s = 4, zt_s = 4, yt_s = 4, nr_s = 3.5, 
                       y_c = "salmon", z_c = "royalblue1")
```
62 top regulators in common

```{r}
common_reg <- Reduce(intersect, list(regAD, regMDD)) 
common_reg <- sort(common_reg)
common_reg
saveRDS(common_reg, "commonreg.rds")
```

