---
title: "GSEA original modules"
author: "Hanne Puype"
date: "2025-04-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

```{r warning=FALSE, message=FALSE}}
library(tidyverse)
library(clusterProfiler)
library(fgsea)
library(WGCNA)
library(org.Hs.eg.db)
```

# GSEA analysis per module: rank genes based on coexpression with module eigengene 
## AD
### Load data
```{r}
modulesAD <- read.table("kmed_modules_AD.txt", header = F)
names(modulesAD) <- c("gene", "module")
```

```{r}
counts_AD <- read.table("Processed_counts_AD_GS.txt", header = T)
counts_AD <- column_to_rownames(counts_AD, var = "GeneSymbol")
```

Prepare input data in right formats
```{r}
counts_AD <- counts_AD %>% filter(rownames(counts_AD) %in% modulesAD$gene)
```

```{r}
modulesAD_vec <- modulesAD$module
names(modulesAD_vec) <- modulesAD$gene
head(modulesAD_vec)
```

Calculate eigengenes of modules 
```{r}
MEs0_AD <- moduleEigengenes(t(counts_AD), modulesAD_vec)$eigengenes 
```


Select eigengenes of interest
```{r}
eigengene_AD_22 <- MEs0_AD[,"ME22"]
eigengene_AD_39 <- MEs0_AD[,"ME39"]
eigengene_AD_51 <- MEs0_AD[,"ME51"]
eigengene_AD_153 <- MEs0_AD[,"ME153"]

names(eigengene_AD_22) <- rownames(MEs0_AD)
names(eigengene_AD_39) <- rownames(MEs0_AD)
names(eigengene_AD_51) <- rownames(MEs0_AD)
names(eigengene_AD_153) <- rownames(MEs0_AD)
```

```{r}
set.seed(123)
organism <- 'org.Hs.eg.db' 
```

Calculate correlation of each gene with the eigengene  
```{r}
ranked_genes_by_cluster_AD <- list()
for (eigengene in c(22, 39, 51, 153)){
   # Calculate correlation of each gene with the eigengene
  correlations <- cor(t(counts_AD), get(paste0("eigengene_AD_", eigengene)), use = "pairwise.complete.obs")
  ranked_genes <- as.vector(correlations)
  names(ranked_genes) <- rownames(correlations)
  # Rank genes by correlation
  ranked_genes <- sort(ranked_genes, decreasing = TRUE)
  # Store ranked genes in the list
  ranked_genes_by_cluster_AD[[eigengene]] <- ranked_genes
}
head(ranked_genes_by_cluster_AD[[22]])
```

Module 22
KEGG
```{r fig.height=8}
# Convert gene symbols to Entrez IDs
entrez_ids <- mapIds(org.Hs.eg.db, keys = names(ranked_genes_by_cluster_AD[[22]]), column = "ENTREZID", 
                     keytype = "SYMBOL")
names(ranked_genes_by_cluster_AD[[22]]) <- entrez_ids
ranked_genes_by_cluster_AD[[22]] <- ranked_genes_by_cluster_AD[[22]][!(is.na(names(ranked_genes_by_cluster_AD[[22]])))]
```

```{r}
kk2 <- gseKEGG(geneList     = ranked_genes_by_cluster_AD[[22]],
               organism     = 'hsa',
               minGSSize    = 10, 
               pvalueCutoff = 0.05,
               verbose      = FALSE)
dotplot(kk2, showCategory=10, split=".sign") + facet_grid(.~.sign) + theme(axis.text.y = element_text(size=10))
ggsave("GSEA_AD_22_KEGG.pdf", device = "pdf", width = 8, height = 10)
```

```{r}
head(kk2, 20)
gseaplot(kk2, geneSetID = "hsa05010", by = "runningScore", title = kk2$Description[8])
ggsave("GSEA_AD_22_KEGG_scoreplot.pdf", device = "pdf", width = 10, height = 6)
gseaplot(kk2, geneSetID = "hsa05010", by = "preranked", title = kk2$Description[8])
gseaplot(kk2, geneSetID = "hsa05010", title = kk2$Description[8])
```

Disease Ontology
```{r fig.height=10}
library(DOSE)
gse_DO <- gseDO(geneList      = ranked_genes_by_cluster_AD[[22]],
                minGSSize     = 10,
                pvalueCutoff  = 0.05,
                pAdjustMethod = "BH",
                organism      = 'hsa',
                verbose       = FALSE)
head(gse_DO, 20)
dotplot(gse_DO, showCategory=15, split=".sign") + facet_grid(.~.sign) + theme(axis.text.y = element_text(size=10))
```

```{r}
gseaplot(gse_DO, geneSetID = "DOID:10652", by = "runningScore", title = gse_DO$Description[10])
ggsave("GSEA_AD_22_DO_scoreplot.pdf", device = "pdf", width = 10, height = 6)
```


Module 39 
KEGG
```{r fig.height=8}
# Convert gene symbols to Entrez IDs
entrez_ids <- mapIds(org.Hs.eg.db, keys = names(ranked_genes_by_cluster_AD[[39]]), column = "ENTREZID", 
                     keytype = "SYMBOL")
names(ranked_genes_by_cluster_AD[[39]]) <- entrez_ids
ranked_genes_by_cluster_AD[[39]] <- ranked_genes_by_cluster_AD[[39]][!(is.na(names(ranked_genes_by_cluster_AD[[39]])))]

kk2 <- gseKEGG(geneList     = ranked_genes_by_cluster_AD[[39]],
               organism     = 'hsa',
               minGSSize    = 10, 
               pvalueCutoff = 0.05,
               verbose      = FALSE,
               eps = 0)
dotplot(kk2, showCategory=10, split=".sign") + facet_grid(.~.sign) + theme(axis.text.y = element_text(size=10))
ggsave("GSEA_AD_39_KEGG.pdf", device = "pdf", width = 8, height = 10)
```

```{r}
head(kk2, 20)
gseaplot(kk2, geneSetID = "hsa05010", by = "runningScore", title = kk2$Description[6])
ggsave("GSEA_AD_39_KEGG_scoreplot.pdf", device = "pdf", width = 10, height = 6)
```

DO
```{r fig.height=10}
gse_DO <- gseDO(geneList      = ranked_genes_by_cluster_AD[[39]],
                minGSSize     = 10,
                pvalueCutoff  = 0.05,
                pAdjustMethod = "BH",
                organism      = 'hsa',
                verbose       = FALSE,
                eps = 0)
head(gse_DO, 20)
dotplot(gse_DO, showCategory=16, split=".sign") + facet_grid(.~.sign) + theme(axis.text.y = element_text(size=10))
```

```{r}
gseaplot(gse_DO, geneSetID = "DOID:10652", by = "runningScore", title = gse_DO$Description[11])
ggsave("GSEA_AD_39_DO_scoreplot.pdf", device = "pdf", width = 10, height = 6)
```

Module 51
KEGG
```{r fig.height=8}
# Convert gene symbols to Entrez IDs
entrez_ids <- mapIds(org.Hs.eg.db, keys = names(ranked_genes_by_cluster_AD[[51]]), column = "ENTREZID", 
                     keytype = "SYMBOL")
names(ranked_genes_by_cluster_AD[[51]]) <- entrez_ids
ranked_genes_by_cluster_AD[[51]] <- ranked_genes_by_cluster_AD[[51]][!(is.na(names(ranked_genes_by_cluster_AD[[51]])))]

kk2 <- gseKEGG(geneList     = ranked_genes_by_cluster_AD[[51]],
               organism     = 'hsa',
               minGSSize    = 10, 
               pvalueCutoff = 0.05,
               verbose      = FALSE,
               eps = 0)
dotplot(kk2, showCategory=10, split=".sign") + facet_grid(.~.sign) + theme(axis.text.y = element_text(size=10))
ggsave("GSEA_AD_51_KEGG.pdf", device = "pdf", width = 8, height = 10)
```

```{r}
head(kk2, 20)
gseaplot(kk2, geneSetID = "hsa05010", by = "runningScore", title = kk2$Description[6])
ggsave("GSEA_AD_51_KEGG_scoreplot.pdf", device = "pdf", width = 10, height = 6)
```

DO
```{r fig.height=10}
gse_DO <- gseDO(geneList      = ranked_genes_by_cluster_AD[[51]],
                minGSSize     = 10,
                pvalueCutoff  = 0.05,
                pAdjustMethod = "BH",
                organism      = 'hsa',
                verbose       = FALSE)
head(gse_DO, 20)
dotplot(gse_DO, showCategory=15, split=".sign") + facet_grid(.~.sign) + theme(axis.text.y = element_text(size=10))
```

```{r}
gseaplot(gse_DO, geneSetID = "DOID:10652", by = "runningScore", title = gse_DO$Description[12])
ggsave("GSEA_AD_51_DO_scoreplot.pdf", device = "pdf", width = 10, height = 6)
```

Module 153

KEGG
```{r fig.height=8}
# Convert gene symbols to Entrez IDs
entrez_ids <- mapIds(org.Hs.eg.db, keys = names(ranked_genes_by_cluster_AD[[153]]), column = "ENTREZID", 
                     keytype = "SYMBOL")
names(ranked_genes_by_cluster_AD[[153]]) <- entrez_ids
ranked_genes_by_cluster_AD[[153]] <- ranked_genes_by_cluster_AD[[153]][!(is.na(names(ranked_genes_by_cluster_AD[[153]])))]

kk2 <- gseKEGG(geneList     = ranked_genes_by_cluster_AD[[153]],
               organism     = 'hsa',
               minGSSize    = 10, 
               pvalueCutoff = 0.05,
               verbose      = FALSE,
               eps = 0)
dotplot(kk2, showCategory=10, split=".sign") + facet_grid(.~.sign) + theme(axis.text.y = element_text(size=10))
ggsave("GSEA_AD_153_KEGG.pdf", device = "pdf", width = 8, height = 10)
```

```{r}
head(kk2, 20)
gseaplot(kk2, geneSetID = "hsa05010", by = "runningScore", title = kk2$Description[6])
ggsave("GSEA_AD_153_KEGG_scoreplot.pdf", device = "pdf", width = 10, height = 6)
```

DO
```{r fig.height=10}
gse_DO <- gseDO(geneList      = ranked_genes_by_cluster_AD[[153]],
                minGSSize     = 10,
                pvalueCutoff  = 0.05,
                pAdjustMethod = "BH",
                organism      = 'hsa',
                verbose       = FALSE)
head(gse_DO, 20)
dotplot(gse_DO, showCategory=15, split=".sign") + facet_grid(.~.sign) + theme(axis.text.y = element_text(size=10))
```

```{r}
gseaplot(gse_DO, geneSetID = "DOID:10652", by = "runningScore", title = gse_DO$Description[10])
ggsave("GSEA_AD_153_DO_scoreplot.pdf", device = "pdf", width = 10, height = 6)
```

## MDD 
### Load data
```{r}
modulesMDD <- read.table("kmed_modules_MDD.txt", header = F)
names(modulesMDD) <- c("gene", "module")
```

```{r}
counts_MDD <- read.table("Processed_counts_MDD_GS.txt", header = T)
counts_MDD <- column_to_rownames(counts_MDD, var = "GeneSymbol")
```

Prepare input data in right formats
```{r}
counts_MDD <- counts_MDD %>% filter(rownames(counts_MDD) %in% modulesMDD$gene)
```

```{r}
modulesMDD_vec <- modulesMDD$module
names(modulesMDD_vec) <- modulesMDD$gene
head(modulesMDD_vec)
```

Calculate eigengenes
```{r}
MEs0_MDD <- moduleEigengenes(t(counts_MDD), modulesMDD_vec)$eigengenes 
```

Select eigengenes of interest
```{r}
eigengene_MDD_24 <- MEs0_MDD[,"ME24"]
eigengene_MDD_36 <- MEs0_MDD[,"ME36"]
eigengene_MDD_110 <- MEs0_MDD[,"ME110"]
eigengene_MDD_115 <- MEs0_MDD[,"ME115"]

names(eigengene_MDD_24) <- rownames(MEs0_MDD)
names(eigengene_MDD_36) <- rownames(MEs0_MDD)
names(eigengene_MDD_110) <- rownames(MEs0_MDD)
names(eigengene_MDD_115) <- rownames(MEs0_MDD)
```

Calculate correlation
```{r}
ranked_genes_by_cluster_MDD <- list()
for (eigengene in c(24, 36, 110, 115)){
   # Calculate correlation of each gene with the eigengene
  correlations <- cor(t(counts_MDD), get(paste0("eigengene_MDD_", eigengene)), use = "pairwise.complete.obs")
  ranked_genes <- as.vector(correlations)
  names(ranked_genes) <- rownames(correlations)
  # Rank genes by correlation
  ranked_genes <- sort(ranked_genes, decreasing = TRUE)
  # Store ranked genes in the list
  ranked_genes_by_cluster_MDD[[eigengene]] <- ranked_genes
}
head(ranked_genes_by_cluster_MDD[[24]])
```

Module 24  
KEGG
```{r fig.height=8}
# Convert gene symbols to Entrez IDs
entrez_ids <- mapIds(org.Hs.eg.db, keys = names(ranked_genes_by_cluster_MDD[[24]]), column = "ENTREZID", 
                     keytype = "SYMBOL")
names(ranked_genes_by_cluster_MDD[[24]]) <- entrez_ids
ranked_genes_by_cluster_MDD[[24]] <- ranked_genes_by_cluster_MDD[[24]][!(is.na(names(ranked_genes_by_cluster_MDD[[24]])))]

kk2 <- gseKEGG(geneList     = ranked_genes_by_cluster_MDD[[24]],
               organism     = 'hsa',
               minGSSize    = 10, 
               pvalueCutoff = 0.05,
               verbose      = FALSE)
dotplot(kk2, showCategory=10, split=".sign") + facet_grid(.~.sign) + theme(axis.text.y = element_text(size=10))
ggsave("GSEA_MDD_24_KEGG.pdf", device = "pdf", width = 8, height = 10)
```

```{r}
head(kk2, 30)
```

```{r fig.height=10}
library(DOSE)
gse_DO <- gseDO(geneList      = ranked_genes_by_cluster_MDD[[24]],
                minGSSize     = 10,
                pvalueCutoff  = 0.05,
                pAdjustMethod = "BH",
                organism      = 'hsa',
                verbose       = FALSE)
head(gse_DO, 30)
dotplot(gse_DO, showCategory=15, split=".sign") + facet_grid(.~.sign) + theme(axis.text.y = element_text(size=10))
```

Module 36  

KEGG
```{r fig.height=8}
# Convert gene symbols to Entrez IDs
entrez_ids <- mapIds(org.Hs.eg.db, keys = names(ranked_genes_by_cluster_MDD[[36]]), column = "ENTREZID", 
                     keytype = "SYMBOL")
names(ranked_genes_by_cluster_MDD[[36]]) <- entrez_ids
ranked_genes_by_cluster_MDD[[36]] <- ranked_genes_by_cluster_MDD[[36]][!(is.na(names(ranked_genes_by_cluster_MDD[[36]])))]

kk2 <- gseKEGG(geneList     = ranked_genes_by_cluster_MDD[[36]],
               organism     = 'hsa',
               minGSSize    = 10, 
               pvalueCutoff = 0.05,
               verbose      = FALSE)
dotplot(kk2, showCategory=10, split=".sign") + facet_grid(.~.sign) + theme(axis.text.y = element_text(size=10))
ggsave("GSEA_MDD_36_KEGG.pdf", device = "pdf", width = 8, height = 10)
```

```{r}
head(kk2, 30)
```

DO
```{r fig.height=10}
gse_DO <- gseDO(geneList      = ranked_genes_by_cluster_MDD[[36]],
                minGSSize     = 10,
                pvalueCutoff  = 0.05,
                pAdjustMethod = "BH",
                organism      = 'hsa',
                verbose       = FALSE)
head(gse_DO, 30)
dotplot(gse_DO, showCategory=16, split=".sign") + facet_grid(.~.sign) + theme(axis.text.y = element_text(size=10))
```

```{r}
gseaplot(gse_DO, geneSetID = "DOID:0060037", by = "runningScore", title = gse_DO$Description[4])
ggsave("GSEA_MDD_36_DO_scoreplot.pdf", device = "pdf", width = 10, height = 6)
```

Module 110  
KEGG  
```{r fig.height=8}
# Convert gene symbols to Entrez IDs
entrez_ids <- mapIds(org.Hs.eg.db, keys = names(ranked_genes_by_cluster_MDD[[110]]), column = "ENTREZID", 
                     keytype = "SYMBOL")
names(ranked_genes_by_cluster_MDD[[110]]) <- entrez_ids
ranked_genes_by_cluster_MDD[[110]] <- ranked_genes_by_cluster_MDD[[110]][!(is.na(names(ranked_genes_by_cluster_MDD[[110]])))]

kk2 <- gseKEGG(geneList     = ranked_genes_by_cluster_MDD[[110]],
               organism     = 'hsa',
               minGSSize    = 10, 
               pvalueCutoff = 0.05,
               verbose      = FALSE)
dotplot(kk2, showCategory=10, split=".sign") + facet_grid(.~.sign) + theme(axis.text.y = element_text(size=10))
ggsave("GSEA_MDD_110_KEGG.pdf", device = "pdf", width = 8, height = 10)
```

```{r}
head(kk2, 20)
```

DO
```{r fig.height=10}
gse_DO <- gseDO(geneList      = ranked_genes_by_cluster_MDD[[110]],
                minGSSize     = 10,
                pvalueCutoff  = 0.05,
                pAdjustMethod = "BH",
                organism      = 'hsa',
                verbose       = FALSE)
head(gse_DO, 30)
dotplot(gse_DO, showCategory=16, split=".sign") + facet_grid(.~.sign) + theme(axis.text.y = element_text(size=10))
```

```{r}
gseaplot(gse_DO, geneSetID = "DOID:0060037", by = "runningScore", title = gse_DO$Description[22])
ggsave("GSEA_MDD_110_DO_scoreplot.pdf", device = "pdf", width = 10, height = 6)
```

Module 115  
KEGG 
```{r fig.height=8}
# Convert gene symbols to Entrez IDs
entrez_ids <- mapIds(org.Hs.eg.db, keys = names(ranked_genes_by_cluster_MDD[[115]]), column = "ENTREZID", 
                     keytype = "SYMBOL")
names(ranked_genes_by_cluster_MDD[[115]]) <- entrez_ids
ranked_genes_by_cluster_MDD[[115]] <- ranked_genes_by_cluster_MDD[[115]][!(is.na(names(ranked_genes_by_cluster_MDD[[115]])))]

kk2 <- gseKEGG(geneList     = ranked_genes_by_cluster_MDD[[115]],
               organism     = 'hsa',
               minGSSize    = 10, 
               pvalueCutoff = 0.05,
               verbose      = FALSE)
dotplot(kk2, showCategory=10, split=".sign") + facet_grid(.~.sign) + theme(axis.text.y = element_text(size=10))
ggsave("GSEA_MDD_115_KEGG.pdf", device = "pdf", width = 8, height = 10)
```

```{r}
head(kk2, 30)
```

DO
```{r fig.height=10}
gse_DO <- gseDO(geneList      = ranked_genes_by_cluster_MDD[[115]],
                minGSSize     = 10,
                pvalueCutoff  = 0.05,
                pAdjustMethod = "BH",
                organism      = 'hsa',
                verbose       = FALSE)
head(gse_DO, 30)
dotplot(gse_DO, showCategory=15, split=".sign") + facet_grid(.~.sign) + theme(axis.text.y = element_text(size=10))
```

```{r}
gseaplot(gse_DO, geneSetID = "DOID:0060037", by = "runningScore", title = gse_DO$Description[8])
ggsave("GSEA_MDD_115_DO_scoreplot.pdf", device = "pdf", width = 10, height = 6)
```

