---
title: "Differential node targeting"
author: "Hanne Puype"
date: "2023-04-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

# Load files

```{r}
library(tidyverse)
library(reshape2)
tableRegulons_AD <- read.csv("tableRegulons_AD.csv")
tableRegulons_MDD <- read.csv2("tableRegulons_MDD.csv")
tableRegulons_AD <- tableRegulons_AD %>% dplyr::select(-X)
colnames(tableRegulons_AD) <- gsub("_.*", "", colnames(tableRegulons_AD))
colnames(tableRegulons_MDD) <- gsub("_.*", "", colnames(tableRegulons_MDD))
tableRegulons_AD <- column_to_rownames(tableRegulons_AD, "gene")
tableRegulons_MDD <- column_to_rownames(tableRegulons_MDD, "gene")
```

# Select common nodes

```{r}
TF_AD <- colnames(tableRegulons_AD)
length(TF_AD)
TF_MDD <- colnames(tableRegulons_MDD)
length(TF_MDD)
TFs <- intersect(TF_AD, TF_MDD)
length(TFs)
```

```{r}
TG_AD <- rownames(tableRegulons_AD)
length(TG_AD)
TG_MDD <- rownames(tableRegulons_MDD)
length(TG_MDD)
TGs <- intersect(TG_AD, TG_MDD)
length(TGs)
```


```{r}
tableRegulons_AD_f <- tableRegulons_AD %>% dplyr::select(all_of(TFs))
tableRegulons_MDD_f <- tableRegulons_MDD %>% dplyr::select(all_of(TFs))
tableRegulons_AD_f <- tableRegulons_AD_f %>% dplyr::filter(rownames(tableRegulons_AD_f) %in% TGs)
dim(tableRegulons_AD_f)
tableRegulons_MDD_f <- tableRegulons_MDD_f %>% dplyr::filter(rownames(tableRegulons_MDD_f) %in% TGs)
dim(tableRegulons_MDD_f)
```

# Format for comparison

edge  AD value  MDD value

## Edge lists
```{r}
edgelist_AD <- melt(rownames_to_column(tableRegulons_AD_f, "gene"))
edgelist_MDD <- melt(rownames_to_column(tableRegulons_MDD_f, "gene"))
edgelist_AD <- edgelist_AD %>% unite("concat", variable:gene, sep = ":", remove = F)
edgelist_MDD <- edgelist_MDD %>% unite("concat", variable:gene, sep = ":", remove = F)
edgelist <- left_join(edgelist_AD, edgelist_MDD, by= "concat")
colnames(edgelist)[names(edgelist) == "value.x"] <- "AD"
colnames(edgelist)[names(edgelist) == "value.y"] <- "MDD"
edgelist2 <- edgelist
colnames(edgelist2)[names(edgelist2) == "gene.x"] <- "gene"
colnames(edgelist2)[names(edgelist2) == "variable.x"] <- "TF"
edgelist2 <- edgelist2 %>% dplyr::filter(AD == 1 | MDD == 1)
edgelist2 <- edgelist2 %>% dplyr::select(TF, gene, AD, MDD)
edgelist <- edgelist %>% dplyr::select(concat, AD, MDD)
```

# Comparison
```{r}
head(edgelist)
edgelist <- edgelist %>% mutate(Dif = case_when(AD == MDD ~ F,
                                                    T ~ T))
edgelist %>% group_by(Dif) %>% summarise(n())
edgelist_dif <- edgelist %>% dplyr::filter(Dif == T)
head(edgelist_dif)
```

# Back to genes and TFs

```{r}
edgelist_dif[, c("TF", "gene")] <- edgelist_dif$concat %>% str_split_fixed(":", 2)
edgelist_dif <- edgelist_dif %>% dplyr::select(TF, gene, AD, MDD)
#edgelist_dif %>% group_by(gene) %>% summarise(n()) # 4288 unique genes
edgelist_dif %>% group_by(TF) %>% summarise(n()) # 168 unique TFs
```

# More than 20% of edges differential

```{r}
diff_genes <- c()
for (genes in unique(edgelist_dif$gene)) {
  edgelist3 <- edgelist2 %>% dplyr::filter(gene == genes)
  a <- dim(edgelist3)
  a <- a[1]
  edgelist_diff <- edgelist_dif %>% dplyr::filter(gene == genes)
  b <- dim(edgelist_diff)
  b <- b[1]
  perc <- b/a
  if (perc > 0.20) {
    if (a > 5 & b > 1) { # more than 5 edges in the networks and more than 1 diff
      diff_genes <- append(diff_genes, genes)
    }
  }
}
# number of total edges of gene
# number of total different edges
# percentage of b/a
# save in vector which genes have percentage > 0.2
```

```{r}
diff_genes <- unique(diff_genes)
length(diff_genes)
write.table(diff_genes, file = "Differential_edge_analysis/all_diff_genes.txt", 
            sep = "\t", row.names = F, quote = F, col.names = T)
```

Let's see for the TFs

```{r}
diff_TFs <- c()
for (TFs in unique(edgelist_dif$TF)) {
  edgelist3 <- edgelist2 %>% dplyr::filter(TF == TFs)
  a <- dim(edgelist3)
  a <- a[1]
  edgelist_diff <- edgelist_dif %>% dplyr::filter(TF == TFs)
  b <- dim(edgelist_diff)
  b <- b[1]
  perc <- b/a
  if (perc > 0.20) {
    if (a > 5) { # more than 5 edges in the networks
      diff_TFs <- append(diff_TFs, TFs)
    }
  }
}
diff_TFs <- unique(diff_TFs)
length(diff_TFs)
write.table(diff_TFs, file = "Differential_edge_analysis/all_diff_TFs.txt", sep = "\t", row.names = F, quote = F, col.names = T)
```

# Active in AD vs MDD
```{r}
diff_AD_genes <- c()
diff_MDD_genes <- c()
diff_genes_neither <- c()
edgelist_dif <- edgelist_dif %>% dplyr::filter(gene %in% diff_genes)
for (genes in unique(edgelist_dif$gene)) {
  edgelist_diff <- edgelist_dif %>% dplyr::filter(gene == genes)
  a <- dim(edgelist_diff)[1]
  edgelist_diff <- edgelist_diff %>% dplyr::filter(AD == 1)
  b <- dim(edgelist_diff)[1]
  perc <- b/a
  if (perc > 0.9) {
    diff_AD_genes <- append(diff_AD_genes, genes)
  }
  else if (perc < 0.1) {
    diff_MDD_genes <- append(diff_MDD_genes, genes)
  }
  else {
    diff_genes_neither <- append(diff_genes_neither, genes)
  }
}
# number of differential edges for gene
# number of edges active in AD 
# b/a
# if > 0.9 --> AD-specific
# if < 0.1 --> MDD-specific
# else none 
```

```{r}
length(diff_AD_genes)
length(diff_MDD_genes)
length(diff_genes_neither)
```

TFs 
```{r}
diff_AD_TFs <- c()
diff_MDD_TFs <- c()
diff_TFs_neither <- c()
for (TFs in unique(edgelist_dif$TF)) {
  edgelist_diff <- edgelist_dif %>% dplyr::filter(TF == TFs)
  a <- dim(edgelist_diff)[1]
  edgelist_diff <- edgelist_diff %>% dplyr::filter(AD == 1)
  b <- dim(edgelist_diff)[1]
  perc <- b/a
  if (perc > 0.9) {
    diff_AD_TFs <- append(diff_AD_TFs, TFs)
  }
  else if (perc < 0.1) {
    diff_MDD_TFs <- append(diff_MDD_TFs, TFs)
  }
  else {
    diff_TFs_neither <- append(diff_TFs_neither, TFs)
  }
}
length(diff_AD_TFs)
length(diff_MDD_TFs)
length(diff_TFs_neither)
```

For both target genes and TFs, there are more AD biased ones  
This could also be the case because there are more edges in the AD network  

```{r}
diff_AD_TFs #IKZF1, IRF8, NEUROD2, RUNX3, SPI1, TAL1
diff_MDD_TFs 
diff_TFs_neither # FOS, FOSB, JUN, JUNB, NEUROD2, PAX6, RUNX1
```

# Save files

```{r}
write.table(diff_AD_genes, "diff_AD_genes_90.txt", sep = "\t", quote = F, row.names = F, col.names = T)
write.table(diff_MDD_genes, "diff_MDD_genes_90.txt", sep = "\t", quote = F, row.names = F, col.names = T)
write.table(diff_genes_neither, "diff_genes_noADMDD_90.txt", sep = "\t", quote = F, row.names = F, col.names = T)
write.table(diff_AD_TFs, "diff_AD_TFs_90.txt", sep = "\t", quote = F, row.names = F, col.names = T)
write.table(diff_MDD_TFs, "diff_MDD_TFs_90.txt", sep = "\t", quote = F, row.names = F, col.names = T)
write.table(diff_TFs_neither, "diff_TFs_noADMDD_90.txt", sep = "\t", quote = F, row.names = F, col.names = T)
```

# Functional enrichment analysis

## All differential edges genes
```{r}
library(enrichR)
dbs <- c("GO_Biological_Process_2021", "GO_Molecular_Function_2021", "KEGG_2021_Human", 
         "Reactome_2016", "WikiPathway_2021_Human")
enriched1 <- enrichr(diff_genes, dbs)
enriched1 <- lapply(enriched1, function(x) x <- x %>% dplyr::filter(Adjusted.P.value <= 0.05))
plotEnrich(enriched1[[1]], showTerms = 20, numChar = 80, y = "Count", orderBy = "P.value", 
           title = "GO Biological Process all genes with differential edges")
plotEnrich(enriched1[[2]], showTerms = 20, numChar = 80, y = "Count", orderBy = "P.value", 
           title = "GO Molecular Function all genes with differential edges")
plotEnrich(enriched1[[3]], showTerms = 20, numChar = 80, y = "Count", orderBy = "P.value", 
           title = "KEGG all genes with differential edges")
plotEnrich(enriched1[[4]], showTerms = 20, numChar = 80, y = "Count", orderBy = "P.value", 
           title = "Reactome all genes with differential edges")
plotEnrich(enriched1[[5]], showTerms = 20, numChar = 80, y = "Count", orderBy = "P.value", 
           title = "Wikipathways all genes with differential edges")
lapply(enriched1, function(x) write.table(data.frame(x), 'FunctionalEnrich_diff_edges/All_diff.txt',
                                           append = T, sep='\t', quote = F))
```

## AD-biased genes

```{r fig.width=10}
enriched_AD <- enrichr(diff_AD_genes, dbs)
enriched_AD <- lapply(enriched_AD, function(x) x <- x %>% dplyr::filter(Adjusted.P.value <= 0.05))
lapply(enriched_AD, function(x) write.table(data.frame(x), 'FunctionalEnrich_diff_edges/AD_biased.txt',
                                           append = T, sep='\t', quote = F))
enriched_AD[["GO_Biological_Process_2021"]][["Term"]] <- gsub("\\(.*","",
                                                              enriched_AD[["GO_Biological_Process_2021"]][["Term"]])
enriched_AD[["GO_Molecular_Function_2021"]][["Term"]] <- gsub("\\(.*","",
                                                              enriched_AD[["GO_Molecular_Function_2021"]][["Term"]])
enriched_AD[[4]][["Term"]] <- gsub(" R-HSA-.*", "", enriched_AD[[4]][["Term"]])
enriched_AD[[5]][["Term"]] <- gsub(" WP.*", "", enriched_AD[[5]][["Term"]])

enriched_AD[[1]][["P.value"]] <- enriched_AD[[1]][["Adjusted.P.value"]]
enriched_AD[[2]][["P.value"]] <- enriched_AD[[2]][["Adjusted.P.value"]]
enriched_AD[[3]][["P.value"]] <- enriched_AD[[3]][["Adjusted.P.value"]]
enriched_AD[[4]][["P.value"]] <- enriched_AD[[4]][["Adjusted.P.value"]]
enriched_AD[[5]][["P.value"]] <- enriched_AD[[5]][["Adjusted.P.value"]]

plotEnrich(enriched_AD[[1]], showTerms = 20, numChar = 80, y = "Count", orderBy = "Adjusted.P.value", 
           title = "GO BP \n AD biased genes") + 
  theme(text=element_text(size = 20)) +
  scale_fill_gradient(low = '#FFAABB', high = '#99DDFF') 
ggsave("GO_BP_difftarget_AD.pdf", height = 6, width = 11, device = "pdf")
plotEnrich(enriched_AD[[2]], showTerms = 20, numChar = 80, y = "Count", orderBy = "P.value", 
           title = "GO Molecular Function AD biased \n genes with differential edges") + 
  theme(text=element_text(size=20)) +
  scale_fill_gradient(low = '#FFAABB', high = '#99DDFF') 
ggsave("GO_MF_difftarget_AD.pdf", height = 6, width = 14, device = "pdf")
plotEnrich(enriched_AD[[3]], showTerms = 20, numChar = 80, y = "Count", orderBy = "P.value", 
           title = "KEGG AD biased genes with \n differential edges") + 
  theme(text=element_text(size=20)) +
  scale_fill_gradient(low = '#FFAABB', high = '#99DDFF') 
ggsave("KEGG_difftarget_AD.pdf", height = 6, width = 14, device = "pdf")
plotEnrich(enriched_AD[[4]], showTerms = 20, numChar = 80, y = "Count", orderBy = "P.value", 
           title = "Reactome AD biased genes with \n differential edges") + 
  theme(text=element_text(size=20)) +
  scale_fill_gradient(low = '#FFAABB', high = '#99DDFF') 
ggsave("Reactome_difftarget_AD.pdf", height = 6, width = 14, device = "pdf")
plotEnrich(enriched_AD[[5]], showTerms = 20, numChar = 80, y = "Count", orderBy = "P.value", 
           title = "WikiPathways AD biased genes \n with differential edges")+ 
  theme(text=element_text(size = 20)) +
  scale_fill_gradient(low = '#FFAABB', high = '#99DDFF') 
ggsave("WP_difftarget_AD.pdf", height = 6, width = 14, device = "pdf")
```

## MDD-biased genes

```{r}
enriched_MDD <- enrichr(diff_MDD_genes, dbs)
enriched_MDD <- lapply(enriched_MDD, function(x) x <- x %>% dplyr::filter(Adjusted.P.value <= 0.05))
lapply(enriched_MDD, function(x) write.table(data.frame(x), 'FunctionalEnrich_diff_edges/MDD_biased.txt',
                                           append = T, sep='\t', quote = F))
enriched_MDD[["GO_Biological_Process_2021"]][["Term"]] <- gsub("\\(.*","",
                                                              enriched_MDD[["GO_Biological_Process_2021"]][["Term"]])
enriched_MDD[["GO_Molecular_Function_2021"]][["Term"]] <- gsub("\\(.*","",
                                                              enriched_MDD[["GO_Molecular_Function_2021"]][["Term"]])
enriched_MDD[[4]][["Term"]] <- gsub(" R-HSA-.*", "", enriched_MDD[[4]][["Term"]])
enriched_MDD[[5]][["Term"]] <- gsub(" WP.*", "", enriched_MDD[[5]][["Term"]])

enriched_AD[[1]][["P.value"]] <- enriched_AD[[1]][["Adjusted.P.value"]]
enriched_AD[[2]][["P.value"]] <- enriched_AD[[2]][["Adjusted.P.value"]]
enriched_AD[[3]][["P.value"]] <- enriched_AD[[3]][["Adjusted.P.value"]]
enriched_AD[[4]][["P.value"]] <- enriched_AD[[4]][["Adjusted.P.value"]]
enriched_AD[[5]][["P.value"]] <- enriched_AD[[5]][["Adjusted.P.value"]]

plotEnrich(enriched_MDD[[1]], showTerms = 20, numChar = 80, y = "Count", orderBy = "P.value", 
           title = "GO BP \n MDD biased genes") + 
  theme(text=element_text(size = 20)) +
  scale_fill_gradient(low = '#FFAABB', high = '#99DDFF') 
ggsave("GO_BP_difftarget_MDD.pdf", height = 5, width = 10, device = "pdf")
plotEnrich(enriched_MDD[[2]], showTerms = 20, numChar = 80, y = "Count", orderBy = "P.value", 
           title = "GO Molecular Function MDD biased \n genes with differential edges") + 
  theme(text=element_text(size = 20)) +
  scale_fill_gradient(low = '#FFAABB', high = '#99DDFF') 
ggsave("GO_MF_difftarget_MDD.pdf", height = 6, width = 14, device = "pdf")
plotEnrich(enriched_MDD[[3]], showTerms = 20, numChar = 80, y = "Count", orderBy = "P.value", 
           title = "KEGG MDD biased genes with differential edges") + 
  theme(text=element_text(size = 20)) +
  scale_fill_gradient(low = '#FFAABB', high = '#99DDFF') 
ggsave("KEGG_difftarget_MDD.pdf", height = 6, width = 14, device = "pdf")
plotEnrich(enriched_MDD[[4]], showTerms = 20, numChar = 80, y = "Count", orderBy = "P.value", 
           title = "Reactome MDD biased genes with \n differential edges") + 
  theme(text=element_text(size = 20)) +
  scale_fill_gradient(low = '#FFAABB', high = '#99DDFF') 
ggsave("Reactome_difftarget_MDD.pdf", height = 6, width = 14, device = "pdf")
plotEnrich(enriched_MDD[[5]], showTerms = 20, numChar = 80, y = "Count", orderBy = "P.value", 
           title = "Wiki Pathways MDD biased genes with \n differential edges") + 
  theme(text=element_text(size = 20)) +
  scale_fill_gradient(low = '#FFAABB', high = '#99DDFF') 
ggsave("WP_difftarget_MDD.pdf", height = 6, width = 14, device = "pdf")
```

