---
title: "Pathways modules figure"
author: "Hanne Puype"
date: "2024-06-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

## Load packages
```{r}
library(tidyverse)
library(pheatmap)
library(RColorBrewer)
```

## Compare modules' pathways  
I need to get the enrichment terms with their p-values for the modules 
heatmap: pathways, modules, annotation of where the term came from and the colors are the adjusted p-values
white when it's not enriched  
Prioritized AD: modules 22, 39, 51 and 153; MDD: modules 24, 36, 110 and 115

steered by similar TFs: AD 141, MDD 37, 63 and 81
```{r}
files <- list.files(path = "FunctionalEnrich_ens", pattern = "*.txt", full.names = TRUE)
files <- files[c(48, 61, 71, 89, 103, 169, 174, 229, 242, 243, 272, 292)]
files
```

Now I need to retrieve the pathways and their p-values for the modules 
```{r}
pathways <- c("regulation of interleukin",
              "regulation of tumor necrosis factor production", "IL-3 signaling pathway", 
              "anti-inflammatory signaling pathway", "interferon gamma signaling", 
              "IL-18 signaling pathway", "antigen bound to MHC molecule", 
              "cell chemotaxis", "Th1 and Th2 cell differentiation", "TCR signaling", "T-cell receptor",
              "regulation of B cell proliferation", "neutrophil degranulation", 
              "TYROBP causal network in microglia","Microglia Pathogen Phagocytosis Pathway", 
              "regulation of microglial cell migration", "phagocytosis, engulfment", 
              "negative regulation of innate immune response",
              "negative regulation of immune response", "negative regulation of adaptive immune response",
              "Toll-like receptor signaling pathway", "response to unfolded protein", 
              "regulation of cell cycle", "regulation of cell apoptotic process", "TNF",
              "apopto", "amyloid beta binding", "Complement activation", "senescence")
```

Get the terms from the files for each module 
```{r}
df <- data.frame("Term" = character(0), "Overlap" = character(), "P.value" = numeric(), 
                 "Adjusted.P.value" = numeric(), "Old.P.value" = numeric(), "Old.Adjusted.P.value" = numeric(),
                 "Odds.Ratio" = numeric(), "Combined.Score" = numeric(), "Genes" = character(), "Module" = character())
for (thisfile in files) {
  data <- read.delim(thisfile, row.names = NULL)
  data <- data %>% dplyr::select(-row.names)
  data$Module <- basename(thisfile)
  data <- data %>% dplyr::filter(grepl(paste(pathways, collapse="|"), Term, ignore.case = T)) 
  df <- rbind(df, data)
}
```

Tweaking for plotting 
```{r}
df <- df %>% mutate(Module = gsub("FunctionalEnrichment_module", "AD_", df$Module))
df <- df %>% mutate(Module = gsub("FunctionalEnrichmentMDD_module", "MDD_", df$Module))
df <- df %>% mutate(Module = gsub(".txt", "", df$Module))
```

```{r}
df <- df %>% select(Term, Adjusted.P.value, Module)
df <- df %>% add_column(DB_pathway = case_when(
  grepl(" \\(GO", df$Term) ~ "GO",
  grepl(" WP", df$Term) ~ "WikiPathways",
  grepl(" Homo", df$Term) ~ "Reactome",
  T ~ "KEGG"))

df$Term <- gsub(" \\(GO.*", "", df$Term)
df$Term <- gsub(" WP.*", "", df$Term)
df$Term <- gsub(" Homo.*", "", df$Term)
```

Filter the redundant or unique ones to one module 
```{r}
df <- df %>% filter(!(Term %in% c("negative regulation of epithelial cell apoptotic process",
                                  "negative regulation of endothelial cell apoptotic process",
                                  "negative regulation of extrinsic apoptotic signaling pathway via death domain receptors",
                                  "regulation of extrinsic apoptotic signaling pathway via death domain receptors",
                                  "regulation of endothelial cell apoptotic process",
                                  "positive regulation of apoptotic signaling pathway",
                                  "positive regulation of cysteine-type endopeptidase activity involved in apoptotic process",
                                  "positive regulation of interleukin-23 production",
                                  "regulation of interleukin-23 production",
                                  "positive regulation of interleukin-13 production",
                                  "positive regulation of leukocyte apoptotic process",
                                  "regulation of interleukin-13 production",
                                  "activation of cysteine-type endopeptidase activity involved in apoptotic process", 
                                  "TRIF-dependent toll-like receptor signaling pathway",
                                  "MyD88-independent toll-like receptor signaling pathway", 
                                  "regulation of interleukin-17 production",
                                  "regulation of cell cycle G1/S phase transition",
                                  "cellular response to unfolded protein",
                                  "negative regulation of cell cycle",
                                  "negative regulation of extrinsic apoptotic signaling pathway",
                                  "negative regulation of extrinsic apoptotic signaling pathway in absence of ligand",
                                  "regulation of extrinsic apoptotic signaling pathway in absence of ligand",
                                  "negative regulation of apoptotic process",
                                  "IL-18 signaling pathway",
                                  "Apoptosis Modulation and Signaling",
                                  "Apoptosis-related network due to altered Notch3 in ovarian cancer",
                                  "negative regulation of dendritic cell apoptotic process",
                                  "positive regulation of cysteine-type endopeptidase activity involved in apoptotic signaling pathway",
                                  "positive regulation of dendritic cell chemotaxis",
                                  "positive regulation of cell cycle G1/S phase transition",
                                  "positive regulation of interleukin-8 production",
                                  "positive regulation of interleukin-6 production",
                                  "negative regulation of interleukin-2 production",
                                  "negative regulation of interleukin-1 beta production",
                                  "positive regulation of interleukin-12 production",
                                  "positive regulation of tumor necrosis factor production",
                                  "negative regulation of tumor necrosis factor production",
                                  "Downstream TCR signaling",
                                  "positive regulation of toll-like receptor signaling pathway",
                                  "negative regulation of B cell proliferation",
                                  "positive regulation of interleukin-4 production",
                                  "Classical antibody-mediated complement activation",
                                  "negative regulation of interleukin-8 production",
                                  "regulation of toll-like receptor signaling pathway",
                                  "positive regulation of interleukin-1 beta production",
                                  "positive regulation of phagocytosis, engulfment",
                                  "positive regulation of interleukin-2 production",
                                  "Toll-like receptor signaling pathway",
                                  "Toll-like Receptor Signaling Pathway",
                                  "positive regulation of apoptotic cell clearance",
                                  "apoptotic process",
                                  "Modulators of TCR signaling and T cell activation",
                                  "regulation of dendritic cell chemotaxis",
                                  "negative regulation of interleukin-6 production",
                                  "dendritic cell chemotaxis",
                                  "negative regulation of interleukin-1 production",
                                  "positive regulation of interleukin-1 production",
                                  "regulation of complement activation",
                                  "negative regulation of toll-like receptor signaling pathway",
                                  "positive regulation of microglial cell migration",
                                  "Regulation of toll-like receptor signaling pathway",
                                  "MyD88-dependent toll-like receptor signaling pathway",
                                  "regulation of neutrophil degranulation",
                                  "complement activation, classical pathway",
                                  "regulation of phagocytosis, engulfment",
                                  "TNF-alpha signaling pathway",
                                  "Senescence-Associated Secretory Phenotype (SASP)",
                                  "Senescence and Autophagy in Cancer",
                                  "Oxidative Stress Induced Senescence")))
```

Some tweaking of the terms to make them more readable
```{r}
df <- df %>% mutate(Term = case_when(Term == 
                       "regulation of T cell activation via T cell receptor contact with antigen bound to MHC molecule on antigen presenting cell" ~ 
                       "regulation T cell activation via TCR contact with antigen bound to MHC molecule on APC", 
                       T ~ Term))
df <- df %>% mutate(Term = case_when(grepl("apopto", df$Term, ignore.case = T) ~ "Apoptosis", 
                                     T ~ Term))
df <- df %>% mutate(DB_pathway = case_when(Term == "Apoptosis" ~ "Mixed", 
                                           T ~ DB_pathway))
df <- df %>% mutate(Term = case_when(grepl("-10", df$Term, ignore.case = T) ~ "IL-10 signaling pathway",
                                     T ~ Term))
df <- df %>% mutate(DB_pathway = case_when(Term == "IL-10 signaling pathway" ~ "Mixed", 
                                           T ~ DB_pathway))
df <- df %>% mutate(Term = case_when(grepl("TNF", df$Term, ignore.case = T) ~ "TNF signaling pathway",
                                     T ~ Term))
df <- df %>% mutate(Term = case_when(grepl("tumor necrosis factor", df$Term, ignore.case = T) ~ "TNF signaling pathway",
                                     T ~ Term))
df <- df %>% mutate(DB_pathway = case_when(Term == "TNF signaling pathway" ~ "Mixed", 
                                           T ~ DB_pathway))
df <- df %>% mutate(Term = case_when(grepl("TCR", df$Term, ignore.case = T) ~ "TCR signaling",
                                     T ~ Term))
df <- df %>% mutate(DB_pathway = case_when(Term == "TCR signaling" ~ "Mixed", 
                                           T ~ DB_pathway))
df <- df %>% mutate(Term = case_when(grepl("Cellular Senescence", df$Term, ignore.case = T) ~ "Cellular senescence",
                                     T ~ Term))
df <- df %>% mutate(DB_pathway = case_when(Term == "Cellular senescence" ~ "Mixed", 
                                           T ~ DB_pathway))
df <- df %>% mutate(Term = case_when(Term == "negative regulation of innate immune response" ~ 
                                       "negative regulation of immune response",
                                     T ~ Term))
df <- df %>% distinct(Term, Module, .keep_all = T)
```

```{r}
df_wide <- df %>% select(-DB_pathway) %>% pivot_wider(names_from = Module, values_from = Adjusted.P.value)
df_wide <- column_to_rownames(df_wide, "Term")
```

```{r}
my_colour <- list(
    DB_pathway = c(GO = "#44BB99", WikiPathways = "#FFAABB",
                   KEGG = "#99DDFF", Reactome = "#EEDD88",
                   Mixed = "#EE8866"))
```

## Plot the heatmap
Only with eight prioritizes modules 
```{r}
df_wide_prior <- df_wide %>% select("AD_22", "AD_39", "AD_51", "AD_153", "MDD_24", "MDD_36", "MDD_110", "MDD_115")
df_wide_prior <- df_wide_prior[rowSums(is.na(df_wide_prior)) < 7,]
df_wide_prior <- df_wide_prior %>% filter(!(rownames(df_wide_prior) %in% 
                                              c("positive regulation of B cell proliferation",
                                                "regulation of cell cycle",
                                                "response to unfolded protein")))
```

```{r}
df_wide_prior_t <- t(df_wide_prior)
df_wide_prior_tt <- apply(df_wide_prior_t, 2, as.numeric)
rownames(df_wide_prior_tt) <- rownames(df_wide_prior_t)
df_wide_prior_tt <- df_wide_prior_tt[,order(colSums(is.na(df_wide_prior_tt)))]
colnames(df_wide_prior_tt) <- toTitleCase(colnames(df_wide_prior_tt))
```

```{r}
df_prior <- df %>% filter(Module %in% c("AD_22", "AD_39", "AD_51", "AD_153", "MDD_24", "MDD_36", "MDD_110", "MDD_115"))
df_prior$Term <- toTitleCase(df_prior$Term)
annot_prior <- df_prior %>% select(Term, DB_pathway) %>% distinct() %>% column_to_rownames("Term")
```

```{r}
df_wide_prior_tt <- as.data.frame(df_wide_prior_tt)
df_wide_prior_tt <- df_wide_prior_tt %>% select("TNF Signaling Pathway", "Apoptosis",
                                                "Regulation of Interleukin-1 Beta Production", 
                                                "Regulation of Interleukin-8 Production",
                                                "TCR Signaling", "Neutrophil Degranulation", 
                                                "IL-10 Signaling Pathway", "Interferon Gamma Signaling",
                                                "TYROBP Causal Network in Microglia",
                                                "Microglia Pathogen Phagocytosis Pathway",
                                                "Regulation of Interleukin-6 Production",
                                                "IL-3 Signaling Pathway", "Negative Regulation of Immune Response",
                                                "Cellular Senescence", everything())
``` 

```{r fig.height=12}
pheatmap(df_wide_prior_tt, annotation_col = annot_prior, cellheight = 20, cellwidth = 20,
         annotation_colors = my_colour, fontsize = 15, annotation_names_col = F,
         cluster_cols = F, cluster_rows = F, na_col = "grey95",
         color = viridis(100, option = "D", direction = 1, begin = 0.1, end = 0.95, alpha = 0.8),
         angle_col = 315, main = "Enriched pathways in prioritized modules")
```

```{r}
pdf("Heatmap_modules_prior.pdf", width = 14, height = 12)
pheatmap(df_wide_prior_tt, annotation_col = annot_prior, cellheight = 20, cellwidth = 21,
         annotation_colors = my_colour, fontsize = 15, annotation_names_col = F,
         cluster_cols = F, cluster_rows = F, na_col = "grey95",
         color = viridis(100, option = "D", direction = 1, begin = 0.1, end = 0.95, alpha = 0.8),
         angle_col = 315, main = "Enriched pathways in prioritized modules",
         labels_row = c("22 AD", "39 AD", "51 AD", "153 AD", "24 MDD", "36 MDD", "110 MDD", "115 MDD"))
dev.off()
```
