---
title: "Heatmap dotplot SCENIC"
author: "Hanne Puype"
date: "2023-07-24"
output:   
  html_document:
    toc: TRUE
    toc_float: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

# Read in data
```{r}
library(Seurat)
library(tidyverse)
library(SCENIC)
#install.packages("cellpypes")
library(cellpypes)
library(viridis)
library(scales)
library(SCopeLoomR)
library(BBmisc)
seuratObjAD <- readRDS("~/MyFiles/Master/single-cell/Seurat_AD.rds")
seuratObjMDD <- readRDS("~/MyFiles/Master/single-cell/Seurat_MDD.rds")
getwd()
```

This script is heavily based on https://github.com/aertslab/scenicplus/blob/main/notebooks/pbmc_multiome_tutorial.ipynb, section dotplot-heatmap  

# AD

```{r}
metaAD <- seuratObjAD@meta.data
metaAD <- metaAD %>% select(celltype)
```

## Get AUC scores, RSS and TF expression 

### TF expression
Pseudobulk per cell type
https://search.r-project.org/CRAN/refmans/cellpypes/html/pseudobulk.html
```{r}
pypeAD <- pype_from_seurat(seuratObjAD)
metaAD$pseudobulk_id <- pseudobulk_id(metaAD)

pseudoCountsAD <- pseudobulk(pypeAD$raw, metaAD$pseudobulk_id)
pseudoCountsAD <- as.data.frame(as.matrix(pseudoCountsAD))
```

### AUC scores
```{r}
AUC_AD <- read_csv2("regulonAUC_AD.csv")
AUC_AD$...1 <- gsub("_.*", "", AUC_AD$...1)
AUC_AD <- AUC_AD %>% column_to_rownames("...1")
colnames(AUC_AD) <- gsub("\\.", "-", colnames(AUC_AD))
head(colnames(AUC_AD))

pseudoAUC_AD <- pseudobulk(as.matrix(AUC_AD), metaAD$pseudobulk_id)
pseudoAUC_AD <- as.data.frame(as.matrix(pseudoAUC_AD))
``` 

```{r}
pseudoCountsAD <- pseudoCountsAD %>% filter(rownames(pseudoCountsAD) %in% rownames(pseudoAUC_AD))
(pseudoAUC_AD %>% filter(!rownames(pseudoAUC_AD) %in% rownames(pseudoCountsAD)))[,1:5]
```

### RSS scores of all cell types
```{r}
RSS_AD <- read.csv("RSS_AD.csv")
rownames(RSS_AD) <- gsub("_.*", "", rownames(RSS_AD))
pseudoCountsAD <- pseudoCountsAD %>% filter(rownames(pseudoCountsAD) %in% rownames(RSS_AD))
```

Highest AUC values
```{r}
ASC <- pseudoAUC_AD["ASC"]
ASC <- ASC %>% arrange(desc(ASC))
ASC <- ASC %>% slice(1:12)

MG <- pseudoAUC_AD["MG"]
MG <- MG %>% arrange(desc(MG))
MG <- MG %>% slice(1:12)

ODC <- pseudoAUC_AD["ODC"]
ODC <- ODC %>% arrange(desc(ODC))
ODC <- ODC %>% slice(1:12)

OPC <- pseudoAUC_AD["OPC"]
OPC <- OPC %>% arrange(desc(OPC))
OPC <- OPC %>% slice(1:12)

EX <- pseudoAUC_AD["EX"]
EX <- EX %>% arrange(desc(EX))
EX <- EX %>% slice(1:12)

INH <- pseudoAUC_AD["INH"]
INH <- INH %>% arrange(desc(INH))
INH <- INH %>% slice(1:12)

PER.END <- pseudoAUC_AD["PER.END"]
PER.END <- PER.END %>% arrange(desc(PER.END))
PER.END <- PER.END %>% slice(1:12)

TFs_selected_AUC <- c(rownames(ASC), rownames(EX), rownames(INH), rownames(MG), rownames(ODC), rownames(OPC), 
                  rownames(PER.END))
TFs_selected_AUC <- unique(TFs_selected_AUC)
TFs_selected_AUC
```

The expression of the TFs is dominated by ODC because these cell were most numerous  
I will divide by the number of counts per cell type to correct for this  
```{r}
sum_ASC <- sum(pseudoCountsAD$ASC)
sum_OPC <- sum(pseudoCountsAD$OPC)
sum_ODC <- sum(pseudoCountsAD$ODC)
sum_MG <- sum(pseudoCountsAD$MG)
sum_INH <- sum(pseudoCountsAD$INH)
sum_EX <- sum(pseudoCountsAD$EX)
sum_PEREND <- sum(pseudoCountsAD$PER.END)
pseudoCountsAD <- pseudoCountsAD %>% mutate(ASC1 = ASC/sum_ASC)
pseudoCountsAD <- pseudoCountsAD %>% mutate(OPC1 = OPC/sum_OPC)
pseudoCountsAD <- pseudoCountsAD %>% mutate(ODC1 = ODC/sum_ODC)
pseudoCountsAD <- pseudoCountsAD %>% mutate(MG1 = MG/sum_MG)
pseudoCountsAD <- pseudoCountsAD %>% mutate(EX1 = EX/sum_EX)
pseudoCountsAD <- pseudoCountsAD %>% mutate(INH1 = INH/sum_INH)
pseudoCountsAD <- pseudoCountsAD %>% mutate(PER.END1 = PER.END/sum_PEREND)
pseudoCountsAD <- pseudoCountsAD %>% select(ASC1, EX1, INH1, MG1, OPC1, ODC1, PER.END1)
colnames(pseudoCountsAD) <- c("ASC", "EX", "INH", "MG", "OPC", "ODC", "PER.END")
```

Also normalize RSS scores as the scores are also highest for ODCs  
```{r}
head(RSS_AD)
sum_ASC <- sum(RSS_AD$ASC)
sum_OPC <- sum(RSS_AD$OPC)
sum_ODC <- sum(RSS_AD$ODC)
sum_MG <- sum(RSS_AD$MG)
sum_INH <- sum(RSS_AD$INH)
sum_EX <- sum(RSS_AD$EX)
sum_PEREND <- sum(RSS_AD$PER.END)
RSS_AD <- RSS_AD %>% mutate(ASC1 = ASC/sum_ASC)
RSS_AD <- RSS_AD %>% mutate(OPC1 = OPC/sum_OPC)
RSS_AD <- RSS_AD %>% mutate(ODC1 = ODC/sum_ODC)
RSS_AD <- RSS_AD %>% mutate(MG1 = MG/sum_MG)
RSS_AD <- RSS_AD %>% mutate(EX1 = EX/sum_EX)
RSS_AD <- RSS_AD %>% mutate(INH1 = INH/sum_INH)
RSS_AD <- RSS_AD %>% mutate(PER.END1 = PER.END/sum_PEREND)
RSS_AD <- RSS_AD %>% select(ASC1, EX1, INH1, MG1, OPC1, ODC1, PER.END1)
colnames(RSS_AD) <- c("ASC", "EX", "INH", "MG", "OPC", "ODC", "PER.END")
```

Normalize AUC scores
```{r}
sum_ASC <- sum(pseudoAUC_AD$ASC)
sum_OPC <- sum(pseudoAUC_AD$OPC)
sum_ODC <- sum(pseudoAUC_AD$ODC)
sum_MG <- sum(pseudoAUC_AD$MG)
sum_INH <- sum(pseudoAUC_AD$INH)
sum_EX <- sum(pseudoAUC_AD$EX)
sum_PEREND <- sum(pseudoAUC_AD$PER.END)
pseudoAUC_AD <- pseudoAUC_AD %>% mutate(ASC1 = ASC/sum_ASC)
pseudoAUC_AD <- pseudoAUC_AD %>% mutate(OPC1 = OPC/sum_OPC)
pseudoAUC_AD <- pseudoAUC_AD %>% mutate(ODC1 = ODC/sum_ODC)
pseudoAUC_AD <- pseudoAUC_AD %>% mutate(MG1 = MG/sum_MG)
pseudoAUC_AD <- pseudoAUC_AD %>% mutate(EX1 = EX/sum_EX)
pseudoAUC_AD <- pseudoAUC_AD %>% mutate(INH1 = INH/sum_INH)
pseudoAUC_AD <- pseudoAUC_AD %>% mutate(PER.END1 = PER.END/sum_PEREND)
pseudoAUC_AD <- pseudoAUC_AD %>% select(ASC1, EX1, INH1, MG1, OPC1, ODC1, PER.END1)
colnames(pseudoAUC_AD) <- c("ASC", "EX", "INH", "MG", "OPC", "ODC", "PER.END")
```

### Select TFs 
```{r}
pseudoCountsAD <- pseudoCountsAD %>% filter(rownames(pseudoCountsAD) %in% TFs_selected_AUC) 
RSS_AD <- RSS_AD %>% filter(rownames(RSS_AD) %in% TFs_selected_AUC)
pseudoAUC_AD <- pseudoAUC_AD %>% filter(rownames(pseudoAUC_AD) %in% TFs_selected_AUC)
```

format df for plotting

```{r}
RSS_AD <- RSS_AD %>% rownames_to_column("TF")
RSS_AD_long <- pivot_longer(RSS_AD, cols = !TF, names_to = "Celltype", values_to = "RSS")

pseudoCountsAD <- pseudoCountsAD %>% rownames_to_column("TF")
counts_AD_long <- pivot_longer(pseudoCountsAD, cols = !TF, names_to = "Celltype", values_to = "Counts")

pseudoAUC_AD <- pseudoAUC_AD %>% rownames_to_column("TF")
AUC_AD_long <- pivot_longer(pseudoAUC_AD, cols = !TF, names_to = "Celltype", values_to = "AUC")
```

```{r}
plotting_df <- merge(counts_AD_long, RSS_AD_long, by = c("TF", "Celltype"), all = T)
plotting_df <- merge(plotting_df, AUC_AD_long, by = c("TF", "Celltype"), all = T)
plotting_df$Counts <- rescale(plotting_df$Counts)
plotting_df$RSS <- rescale(plotting_df$RSS)
plotting_df$AUC <- rescale(plotting_df$AUC)
plotting_df$Celltype <- factor(plotting_df$Celltype, levels = rev(unique(plotting_df$Celltype)))
plotting_df$TF <- paste0(plotting_df$TF, "(+)")
```

### Heatmap dotplot with selected TFs, showing AUC values, TF expression and RSS score  
https://scenicplus.readthedocs.io/en/latest/_modules/scenicplus/plotting/dotplot.html  
```{r fig.height=6, fig.width=13}
ggplot(plotting_df, aes(x = TF, y = Celltype)) +
  geom_tile(mapping = aes(fill = AUC)) + 
  scale_fill_viridis(name = "Regulon activity", option = "A", direction = -1, begin = 0.1, end = 0.9) +
  new_scale_fill() +
  geom_point(mapping = aes(size = RSS, fill = Counts),
              shape = 21, colour = "black", stroke = 0.1) + 
  scale_size_continuous(range = c(0.3, 5), name = "RSS") + # change range size of points
  scale_fill_viridis(name = "TF expression", option = "E", begin = 0.1) +
  coord_fixed(ratio = 1.7) +
  scale_y_discrete(labels = c("Pericytes-endothelial cells", "OPCs", "Oligodendrocytes", 
                              "Microglia", "Inhibitory neurons", "Excitatory neurons",  
                              "Astrocytes")) +
  scale_x_discrete(limits = c(paste0(TFs_selected_AUC, "(+)"))) +
  theme_classic() +
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(), axis.line = element_blank(),
        axis.text.x = element_text(angle = 90, hjust = 1), text = element_text(size = 15), 
        legend.position = "bottom")
ggsave("Heatmap_dotplot_AD_top12AUC.pdf", device = "pdf", width = 13, height = 4)
```


# MDD

```{r}
metaMDD <- seuratObjMDD@meta.data
metaMDD <- metaMDD %>% select(Celltype) 
```

### RSS scores and TF expression
```{r}
RSS_MDD <- read.csv("RSS_MDD.csv")
rownames(RSS_MDD) <- gsub("_.*", "", rownames(RSS_MDD))

pypeMDD <- pype_from_seurat(seuratObjMDD)
metaMDD$pseudobulk_id <- pseudobulk_id(metaMDD)

pseudoCountsMDD <- pseudobulk(pypeMDD$raw, metaMDD$pseudobulk_id)
pseudoCountsMDD <- as.data.frame(as.matrix(pseudoCountsMDD))
pseudoCountsMDD <- pseudoCountsMDD %>% filter(rownames(pseudoCountsMDD) %in% rownames(RSS_MDD))

RSS_MDD <- RSS_MDD %>% select(-MIX)
pseudoCountsMDD <- pseudoCountsMDD %>% select(-MIX)
```

### AUC scores
```{r}
#AUC_MDD <- read_csv2("regulonAUC_MDD.csv")
loom <- open_loom(file.path = "~/MyFiles/Master/single-cell/SCENIC_output_MDD.loom")
regulonAUC <- get_regulons_AUC(loom)
close_loom(loom)
head(regulonAUC)
AUCdata <- getAUC(regulonAUC) %>% BBmisc::normalize("range")
AUC_MDD <- as.data.frame(AUCdata)
rownames(AUC_MDD) <- gsub("_.*", "", rownames(AUC_MDD))

metaMDD_AUC <- metaMDD
rownames(metaMDD_AUC) <- sub("/", ".", rownames(metaMDD))
metaMDD_AUC <- metaMDD_AUC %>% filter(rownames(metaMDD_AUC) %in% colnames(AUC_MDD))
pseudoAUC_MDD <- pseudobulk(as.matrix(AUC_MDD), metaMDD_AUC$pseudobulk_id)
pseudoAUC_MDD <- as.data.frame(as.matrix(pseudoAUC_MDD))
```

```{r}
pseudoCountsMDD <- pseudoCountsMDD %>% filter(rownames(pseudoCountsMDD) %in% rownames(pseudoAUC_MDD))
(pseudoAUC_MDD %>% filter(!rownames(pseudoAUC_MDD) %in% rownames(pseudoCountsMDD)))[,1:5]
```

Select TFs according to AUC scores
```{r}
ASC <- pseudoAUC_MDD["ASC"]
ASC <- ASC %>% arrange(desc(ASC))
ASC <- ASC %>% slice(1:10)

MICRO <- pseudoAUC_MDD["MICRO"]
MICRO <- MICRO %>% arrange(desc(MICRO))
MICRO <- MICRO %>% slice(1:10)

ODC <- pseudoAUC_MDD["ODC"]
ODC <- ODC %>% arrange(desc(ODC))
ODC <- ODC %>% slice(1:10)

OPC <- pseudoAUC_MDD["OPC"]
OPC <- OPC %>% arrange(desc(OPC))
OPC <- OPC %>% slice(1:10)

EX <- pseudoAUC_MDD["EX"]
EX <- EX %>% arrange(desc(EX))
EX <- EX %>% slice(1:10)

INH <- pseudoAUC_MDD["INH"]
INH <- INH %>% arrange(desc(INH))
INH <- INH %>% slice(1:10)

END <- pseudoAUC_MDD["END"]
END <- END %>% arrange(desc(END))
END <- END %>% slice(1:10)

TFs_selected_AUC <- c(rownames(ASC), rownames(END), rownames(EX), rownames(INH), 
                      rownames(MICRO), rownames(ODC), rownames(OPC))
TFs_selected_AUC <- unique(TFs_selected_AUC)
TFs_selected_AUC
```

I will divide by the number of counts per cell type to correct for the number of counts per cell type
```{r}
sum_ASC <- sum(pseudoCountsMDD$ASC)
sum_OPC <- sum(pseudoCountsMDD$OPC)
sum_ODC <- sum(pseudoCountsMDD$ODC)
sum_MICRO <- sum(pseudoCountsMDD$MICRO)
sum_EX <- sum(pseudoCountsMDD$EX)
sum_INH <- sum(pseudoCountsMDD$INH)
sum_END <- sum(pseudoCountsMDD$END)
pseudoCountsMDD <- pseudoCountsMDD %>% mutate(ASC1 = ASC/sum_ASC)
pseudoCountsMDD <- pseudoCountsMDD %>% mutate(OPC1 = OPC/sum_OPC)
pseudoCountsMDD <- pseudoCountsMDD %>% mutate(ODC1 = ODC/sum_ODC)
pseudoCountsMDD <- pseudoCountsMDD %>% mutate(MICRO1 = MICRO/sum_MICRO)
pseudoCountsMDD <- pseudoCountsMDD %>% mutate(EX1 = EX/sum_EX)
pseudoCountsMDD <- pseudoCountsMDD %>% mutate(INH1 = INH/sum_INH)
pseudoCountsMDD <- pseudoCountsMDD %>% mutate(END1 = END/sum_END)
pseudoCountsMDD <- pseudoCountsMDD %>% select(ASC1, END1, EX1, INH1, MICRO1, OPC1, ODC1)
colnames(pseudoCountsMDD) <- c("ASC", "END", "EX", "INH", "MICRO", "OPC", "ODC")
```


Normalize RSS scores
```{r}
head(RSS_MDD)
sum_ASC <- sum(RSS_MDD$ASC)
sum_OPC <- sum(RSS_MDD$OPC)
sum_ODC <- sum(RSS_MDD$ODC)
sum_MICRO <- sum(RSS_MDD$MICRO)
sum_EX <- sum(RSS_MDD$EX)
sum_INH <- sum(RSS_MDD$INH)
sum_END <- sum(RSS_MDD$END)
RSS_MDD <- RSS_MDD %>% mutate(ASC1 = ASC/sum_ASC)
RSS_MDD <- RSS_MDD %>% mutate(OPC1 = OPC/sum_OPC)
RSS_MDD <- RSS_MDD %>% mutate(ODC1 = ODC/sum_ODC)
RSS_MDD <- RSS_MDD %>% mutate(MICRO1 = MICRO/sum_MICRO)
RSS_MDD <- RSS_MDD %>% mutate(EX1 = EX/sum_EX)
RSS_MDD <- RSS_MDD %>% mutate(INH1 = INH/sum_INH)
RSS_MDD <- RSS_MDD %>% mutate(END1 = END/sum_END)
RSS_MDD <- RSS_MDD %>% select(ASC1, END1, EX1, INH1, MICRO1, OPC1, ODC1)
colnames(RSS_MDD) <- c("ASC", "END", "EX", "INH", "MICRO", "OPC", "ODC")
```

Normalize AUC scores
```{r}
sum_ASC <- sum(pseudoAUC_MDD$ASC)
sum_OPC <- sum(pseudoAUC_MDD$OPC)
sum_ODC <- sum(pseudoAUC_MDD$ODC)
sum_MG <- sum(pseudoAUC_MDD$MICRO)
sum_INH <- sum(pseudoAUC_MDD$INH)
sum_EX <- sum(pseudoAUC_MDD$EX)
sum_PEREND <- sum(pseudoAUC_MDD$END)
pseudoAUC_MDD <- pseudoAUC_MDD %>% mutate(ASC1 = ASC/sum_ASC)
pseudoAUC_MDD <- pseudoAUC_MDD %>% mutate(OPC1 = OPC/sum_OPC)
pseudoAUC_MDD <- pseudoAUC_MDD %>% mutate(ODC1 = ODC/sum_ODC)
pseudoAUC_MDD <- pseudoAUC_MDD %>% mutate(MG1 = MICRO/sum_MG)
pseudoAUC_MDD <- pseudoAUC_MDD %>% mutate(EX1 = EX/sum_EX)
pseudoAUC_MDD <- pseudoAUC_MDD %>% mutate(INH1 = INH/sum_INH)
pseudoAUC_MDD <- pseudoAUC_MDD %>% mutate(PER.END1 = END/sum_PEREND)
pseudoAUC_MDD <- pseudoAUC_MDD %>% select(ASC1, EX1, INH1, MG1, OPC1, ODC1, PER.END1)
colnames(pseudoAUC_MDD) <- c("ASC", "EX", "INH", "MICRO", "OPC", "ODC", "END")
```

Select TFs 
```{r}
pseudoCountsMDD <- pseudoCountsMDD %>% filter(rownames(pseudoCountsMDD) %in% TFs_selected_AUC) 
RSS_MDD <- RSS_MDD %>% filter(rownames(RSS_MDD) %in% TFs_selected_AUC)
pseudoAUC_MDD <- pseudoAUC_MDD %>% filter(rownames(pseudoAUC_MDD) %in% TFs_selected_AUC)
```

```{r}
RSS_MDD <- RSS_MDD %>% rownames_to_column("TF")
RSS_MDD_long <- pivot_longer(RSS_MDD, cols = !TF, names_to = "Celltype", values_to = "RSS")

pseudoCountsMDD <- pseudoCountsMDD %>% rownames_to_column("TF")
counts_MDD_long <- pivot_longer(pseudoCountsMDD, cols = !TF, names_to = "Celltype", values_to = "Counts")

pseudoAUC_MDD <- pseudoAUC_MDD %>% rownames_to_column("TF")
AUC_MDD_long <- pivot_longer(pseudoAUC_MDD, cols = !TF, names_to = "Celltype", values_to = "AUC")
```

format df for plotting  
```{r}
plotting_df <- merge(counts_MDD_long, RSS_MDD_long, by = c("TF", "Celltype"), all = T)
plotting_df <- merge(plotting_df, AUC_MDD_long, by = c("TF", "Celltype"), all = T)
plotting_df$Counts <- rescale(plotting_df$Counts)
plotting_df$RSS <- rescale(plotting_df$RSS)
plotting_df$AUC <- rescale(plotting_df$AUC)
plotting_df$Celltype <- factor(plotting_df$Celltype, levels = rev(unique(plotting_df$Celltype)))
plotting_df$TF <- paste0(plotting_df$TF, "(+)")
```


### Heatmap dotplot with selected TFs, showing AUC scores, TF expression and RSS score  
https://scenicplus.readthedocs.io/en/latest/_modules/scenicplus/plotting/dotplot.html  
```{r fig.height=6, fig.width=13}
ggplot(plotting_df, aes(x = TF, y = Celltype)) +
  geom_tile(mapping = aes(fill = AUC)) + 
  scale_fill_viridis(name = "Regulon activity", option = "A", direction = -1, begin = 0.2, end = 0.95) +
  new_scale_fill() +
  geom_point(mapping = aes(size = RSS, fill = Counts),
              shape = 21, colour = "black", stroke = 0.1) + 
  scale_size_continuous(range = c(0.3, 5), name = "RSS") + # change range size of points
  scale_fill_viridis(name = "TF expression", option = "E", begin = 0.1) +
  coord_fixed(ratio = 1.7) +
  scale_y_discrete(labels = c("OPCs", "Oligodendrocytes", "Microglia", "Inhibitory neurons", 
                              "Excitatory neurons", "Endothelial cells", "Astrocytes")) +
  scale_x_discrete(limits = c(paste0(TFs_selected_AUC, "(+)"))) +
  theme_classic() +
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(), axis.line = element_blank(),
        axis.text.x = element_text(angle = 90, hjust = 1), text = element_text(size = 15), 
        legend.position = "bottom")
ggsave("Heatmap_dotplot_MDD_top10AUC.pdf", device = "pdf", width = 13, height = 4)
```
