---
title: "Post processing of SCENIC analysis on MDD female dataset"
output: html_notebook
---

Load libraries: 

```{r message=FALSE}
library("SingleCellExperiment")
library("SCENIC")
library("BBmisc")
library("SCopeLoomR")
library("Seurat")
library("dplyr")
library("readr")
library("ggplot2")
library("openxlsx")
library("circlize")
library("RColorBrewer")
library("magick")
library("ComplexHeatmap")
library("VennDiagram")
library("dendextend")
```

Create Directories: 
```{r}
out_dir = "~/Desktop/rebuttal_Hanne/"
setwd(out_dir)
dir.create(paste0(out_dir, "Figures/"))
dir.create(paste0(out_dir, "Tables/"))
```

Set directory for output: 
```{r}
fig_dir <- paste0(out_dir, "Figures/")
table_dir <- paste0(out_dir, "Tables/")
```

Import metadata: 
```{r}
metaData = read.table("Tables/MDD_M_F_metadata.csv", header = T, sep = ",", row.names = 1)
head(metaData)
```


Read SCENIC output from loom file:
```{r}
#for some reason paths only work when absolute
loom <- open_loom(file = "~/Desktop/rebuttal_Hanne/SCENIC_output_MDD_female.loom")
ps_cells <- get_cell_ids(loom)
regulonAUC <- get_regulons_AUC(loom)
thresholds <- get_regulon_thresholds(loom)
regulons_incidMat <- get_regulons(loom)
regulons <- regulonsToGeneLists(regulons_incidMat)
close_loom(loom)
```


```{r}
# Subset metadata to cells -> here metaData has differently names Microglia, but the cell types are already part of the cell prefixes so I will use those for the Microglia

row.names(metaData) <- ifelse(
  startsWith(row.names(metaData), "Micro"), 
  gsub("Micro/Macro", "Micro.Macro", row.names(metaData)), 
  row.names(metaData)
)

metaData <- metaData[ps_cells,]
dim(metaData)
```



First checkpoint for evaluating the outcome of SCENIC: did it find regulons?
```{r}
dim(regulonAUC)
length(thresholds)
length(regulons)
```

```{r}
head(regulonAUC)
head(thresholds)
```

Save regulon matrix
```{r}
toExport<-as.data.frame(t(regulons_incidMat), stringsAsFactors=F)
toExport$gene<-rownames(toExport)
toExport<-toExport[,c('gene',setdiff(colnames(toExport),"gene"))]
tmpList<-list(toExport)
names(tmpList)<-"regulonMatrix"
write.xlsx(tmpList,paste0(table_dir, "tableRegulons_MDD.xlsx"))
```


FIX THRESHOLDS: names and values switched in output. pySCENIC/loom bug?

```{r}
thresholds_values <- names(thresholds)
thresholds_names <- thresholds
thresholds <- thresholds_values
names(thresholds) <- thresholds_names
length(thresholds)
length(regulons)
head(thresholds)
```

Normally the thresholds and regulons should have the same dimensions: I will filter the regulons not in common out

```{r}
sum(names(regulons) %in% names(thresholds))
which(names(thresholds) %in% names(regulons)) 
thresholds <- thresholds[which(names(thresholds) %in% names(regulons))]
length(thresholds)
length(regulons)
head(thresholds)
```

Get AUC values in matrix + scale to range 0-1 for nicer heatmaps

```{r}
AUCdata <- getAUC(regulonAUC) %>% BBmisc::normalize("range") #by default, operates on rows (regulons)
dim(AUCdata)
AUCdata[1:5,1:5]
```

# Define regulons per cell

```{r}
regulonsCells <- setNames(lapply(names(thresholds), 
                                 function(x) {
                                   trh <- thresholds[x]
                                   names(which(getAUC(regulonAUC)[x,]>trh))
                                 }),names(thresholds))

length(regulonsCells)


regulonActivity <- reshape2::melt(regulonsCells)
head(regulonActivity)
```


# Create binary activity (regulon On or Off based on thresholds)

```{r}
binaryRegulonActivity <- t(table(regulonActivity[,1], regulonActivity[,2])) #table can also build matrices
class(binaryRegulonActivity) <- "matrix"
dim(binaryRegulonActivity)
binaryRegulonActivity[1:5,1:5]
```


# Only non-duplicated regulons

```{r}
if (sum(duplicated(rownames(binaryRegulonActivity))) != 0){
  
  binaryRegulonActivity_nonDupl <- binaryRegulonActivity[which(rownames(binaryRegulonActivity) %in% onlyNonDuplicatedExtended(rownames(binaryRegulonActivity))),]
  dim(binaryRegulonActivity_nonDupl)
  
} else {
  print("No duplicated regulons... Results from RegulonActivity_nonDupl and binaryRegulonActivity will be the same")
  binaryRegulonActivity_nonDupl = binaryRegulonActivity
}

```


# Create set of CRCs

```{r}
regulonSelection <- list()
regulonSelection[["CRCs"]] <- c("IKZF1", "IRF8", "NFATC2", "MAF", "RUNX1", "TAL1", "SPI1", "MAFF", "STAT3", "MYC", "FOS", "JUNB", "CEBPD", "PLSCR1")
regulonSelection[["CRCs"]] <- paste0(regulonSelection[["CRCs"]], "_(+)-motif")
regulonSelection[["CRCs"]] <- names(regulons)[names(regulons) %in% regulonSelection[["CRCs"]]]
regulonSelection[["CRCs"]]
```


# Get minimum cells

I lowered the amount of minamal cells where a regulon is active in this case, it would be 143 cells. Which is probably too high. 

```{r}
minCells <- ncol(binaryRegulonActivity) * .001 # 1% cells
```

### All regulons

```{r}
regulonSelection[["All_regulons_with_duplicated"]] <- rownames(binaryRegulonActivity)
length(rownames(binaryRegulonActivity))
```

### Active in > 1% cells

The binary activity is one ("on") in more than 1% of the cells for this regulon. 
FROM the supplementary in the SCENIC paper : "For example, in Fig. 1b, we have plotted only regulons active in at least 1% of the cells and correlated with other regulons in the matrix (absolute correlation > 0.30). However, the downstream analyses include all the regulons.

```{r}
regMinCells <- names(which(rowSums(binaryRegulonActivity_nonDupl) > minCells))
regulonSelection[["Regs_active_morethan_1perc_cells"]] <- regMinCells
length(regMinCells)
```

### Correlation between regulons
```{r}
reguCor <- cor(t(binaryRegulonActivity_nonDupl[regMinCells,]))
diag(reguCor) <- 0
nrow(reguCor)
```


### Regulons correlated with other regulons and active in > 1% cells

```{r}
corrRegs <- names(which(rowSums(abs(reguCor) > 0.3) > 0)) 
regulonSelection[["Regs_active_and_corr_0.3"]]  <- corrRegs
length(corrRegs)
```

```{r}
#Extra: add corr>0.5
corrRegs_05 <- names(which(rowSums(abs(reguCor) > 0.5) > 0)) 
regulonSelection[["Regs_active_and_corr_0.5"]]  <- corrRegs_05
length(corrRegs_05)
```



### Regulons NOT correlated with other regulons and/or active in < 1% cells

```{r}
missingRegs <- rownames(binaryRegulonActivity_nonDupl)[which(!rownames(binaryRegulonActivity_nonDupl) %in% corrRegs)]
regulonSelection[["Regs_inactive_or_noncorr"]]  <- missingRegs
length(missingRegs)
```

## Aggregated heatmaps

# Heatmaps

```{r}
v <- metaData$Celltype
names(v) <- row.names(metaData)
idents <- sort(v)
ordered <- names(idents)

nrClust <- length(levels(as.factor(v)))
nrClust
```


```{r}
colors <- c("#ED8141")
colors <- colorRampPalette(colors)(nrClust)
names(colors) <- unique(metaData$Celltype)
colors <- list(idents = colors)

tables <- table(v)
names(tables) <- NULL
```


### Setting gaps

```{r}
gaps <- tables[1]
for (k in 1:nrClust) { #add values together
  value <- tables[k]
  gaps <- append(gaps, value+gaps[k-1])
}


for (i in seq_len(length(regulonSelection))) {
  selRegs <- names(regulonSelection)[i]
  if(length(regulonSelection[[selRegs]])>1) {
    
    AUCMat <- AUCdata[regulonSelection[[selRegs]], ,drop=FALSE]
    AUCMat <- AUCMat[,ordered]
    
    avg_exp_df <- data.frame(row.names = regulonSelection[[selRegs]]) #Create empty dataframe
    
    for (j in levels(as.factor(v))) {
      AUCMatTmp <- AUCMat[,names(as.factor(v)[which(v==j)])] #Subset for cell population
      AUCMatTmp <- as.matrix(AUCMatTmp)
      
      if (dim(AUCMatTmp)[2] == 1) {
        average_expressions <- AUCMatTmp
      } else {
        average_expressions <- apply(AUCMatTmp, 1, function(x){ #apply across row: average activity in regulon for that cell pop
          sum(x)/length(colnames(AUCMatTmp))
        })
      }
      
      average_expressions <- as.numeric(average_expressions)
      avg_exp_df[,j] <- average_expressions
    }
    
    names(avg_exp_df) <- levels(as.factor(v))
    AUCMatT <- avg_exp_df
    
    p <- pheatmap(as.matrix(AUCMatT), cluster_cols = F, cluster_rows = F, scale = "none", 
                  main = selRegs, treeheight_row = 0, 
                  cellwidth = 10, cellheight = 10, 
                  show_colnames = T, border_color = "black",
                  fontsize_row = 4, fontsize_col = 4, angle_col = "315")
    
    file_name <- paste0(fig_dir, "Heatmap_AUCRegulons_Averaged_", selRegs ,"_MDD_F.pdf")
    
    p.height <- 1 + 0.15 * length(rownames(AUCMatT))
    
    pdf(file = file_name, 
        width = 10, 
        height = p.height, 
    ) 
    
    draw(p, merge_legend = T)
    
    dev.off()
  }
}

```


## Heatmaps with ComplexHeatmap

### Create annotations


```{r}
cols_celltype <- brewer.pal(length(levels(as.factor(metaData$Celltype))), "Set1")
names(cols_celltype) <- levels(as.factor(metaData$Celltype))

col_ann_types = columnAnnotation( Celltype = metaData[colnames(AUCdata), 'Celltype'], show_legend = T, annotation_legend_param = list(Celltype = list(title = "Celltype")),
                            show_annotation_name = T, annotation_label = "Celltype", 
                            col = list(Celltype = cols_celltype))
```


```{r}
for(i in seq_len(length(regulonSelection))){ 
  
  selRegs <- names(regulonSelection)[i]
  
  if (selRegs == "CRCs"){
    next
  }
  
  if(length(regulonSelection[[selRegs]])>1) {
    
    AUCMat <- AUCdata[regulonSelection[[selRegs]],,drop=FALSE]

    col_fun = colorRamp2(c(0, 0.5, 1), c("steelblue3", "#FFFFBF", "#D73027")) 
    col_fun(seq(-50, 50))
    
    hm <- Heatmap(AUCMat, show_column_names = FALSE, name="AUCell score", 
                    clustering_distance_rows = "euclidean", 
                    clustering_distance_columns = "euclidean", 
                    clustering_method_rows = "ward.D", 
                    clustering_method_columns = "ward.D", 
                    column_dend_height = unit(1.5, "cm"), row_dend_width = unit(1.5, "cm"), 
                    width = unit(15, "cm"), height = unit((0.07*dim(AUCMat)[1]), "cm"), 
                    row_names_gp = gpar(fontsize = 2), column_names_gp = gpar(fontsize = 1),
                    col=col_fun, top_annotation = col_ann_types
    )
    
    message("Creating heatmap for", selRegs)
    
    pdf(paste0(fig_dir, "RegulonActivity_Heatmap_",selRegs,"_ComplexHeatmap_MDD_F.pdf"), width=12.45, height=17.55)
    draw(hm, merge_legend = TRUE)  
    dev.off()
    
    
  }
}
```

