---
title: "Post processing of SCENIC analysis on AD dataset - old runs made by Joke"
output: html_notebook
---

Load libraries: 

```{r message=FALSE}
library("SingleCellExperiment")
library("SCENIC")
library("BBmisc")
library("SCopeLoomR")
library("Seurat")
library("dplyr")
library("readr")
library("ggplot2")
library("openxlsx")
library("circlize")
library("RColorBrewer")
library("magick")
library("ComplexHeatmap")
library("VennDiagram")
library("dendextend")
```

Create Directories: 
```{r}
out_dir = "~/Desktop/rebuttal_Hanne/"
setwd(out_dir)
dir.create(paste0(out_dir, "Figures/"))
dir.create(paste0(out_dir, "Tables/"))
```

Set directory for output: 
```{r}
fig_dir <- paste0(out_dir, "Figures/")
table_dir <- paste0(out_dir, "Tables/")
```

Import metadata: 
```{r}
metaData = read.table("Tables/snRNA_metadta.csv", header = T, sep = ",", row.names = 1)
head(metaData)
```


Read SCENIC output from loom file:
```{r}
#for some reason paths only work when absolute
loom <- open_loom(file = "~/Desktop/rebuttal_Hanne/SCENIC_output_AD_large.loom")
ps_cells <- get_cell_ids(loom)
regulonAUC <- get_regulons_AUC(loom)
thresholds <- get_regulon_thresholds(loom)
regulons_incidMat <- get_regulons(loom)
regulons <- regulonsToGeneLists(regulons_incidMat)
close_loom(loom)
```

```{r}
# Create edge list

edges <- data.frame(unlist(regulons))
edges$TFs <- unlist(lapply(strsplit(row.names(edges), "_"), "[", 1))
write.csv( edges, paste0(table_dir, "Edges_AD_large.csv"), quote = F)
```

```{r}
# Create edge list restricted for DEGs

degs <- readr::read_csv("Tables/diff_exp_celltypes_ADvsC.csv")
head(degs)
```



```{r}
edges_degs<-edges[edges$unlist.regulons. %in% degs[startsWith(degs$celltype, "MG"), ]$genes ,] # for microglia
dim(edges_degs)
```

```{r}
# Rank TFs by degree

degree_tfs <- table(edges_degs$TFs)
degree_tfs <- sort(degree_tfs, decreasing = T)
head(degree_tfs)
```

```{r}
# Make rank plot

df_plot <- data.frame(degree_tfs)
df_plot$Rank <- 1:nrow(df_plot)
colnames(df_plot)[1:2] <- c("TF", "Degree")
df_plot
```

```{r}
ggplot(df_plot, aes(x=Rank, y=Degree)) +
  geom_point() +
  # geom_text(aes(label=ifelse(Count<=5,as.character(TF),'')),hjust=-0.2,vjust=0) +
  ggrepel::geom_label_repel(aes(label=ifelse(Rank<=10,as.character(TF),'')),
                  box.padding   = 0.35, 
                  point.padding = 0.5,
                  nudge_x = 2,
                  segment.color = 'grey50') +
  theme_bw() +
  ggtitle("Degree of TFs regulating DEGs in AD vs control")

ggsave(paste0(fig_dir, "AD_GRN_fromDEGs_MG_degreeRankPlot.png"))
ggsave(paste0(fig_dir, "AD_GRN_fromDEGs_MG_degreeRankPlot.svg"))
```

```{r}
# Repeat for all DEGs
edges_degs<-edges[edges$unlist.regulons. %in% degs[startsWith(degs$celltype, "MG"), ]$genes ,] 
dim(edges_degs)
```

```{r}
# Rank TFs by degree

degree_tfs <- table(edges_degs$TFs)
degree_tfs <- sort(degree_tfs, decreasing = T)
head(degree_tfs)
```

```{r}
# Make rank plot

df_plot <- data.frame(degree_tfs)
df_plot$Rank <- 1:nrow(df_plot)
colnames(df_plot)[1:2] <- c("TF", "Degree")
df_plot
```

```{r}
# Export table for CRCs

crc_tfs <- c("IKZF1", "IRF8", "NFATC2", "MAF", "RUNX1", "TAL1", "SPI1", "MAFF", "STAT3", "MYC", "FOS", "JUNB", "CEBPD", "PLSCR1")

write.csv(df_plot[df_plot$TF %in% crc_tfs,], paste0(table_dir, "AD_GRN_fromDEGs_ALL_degreeRanks_CRCs.csv"), row.names = F, quote = F)
```


```{r}
ggplot(df_plot, aes(x=Rank, y=Degree)) +
  geom_point() +
  # geom_text(aes(label=ifelse(Count<=5,as.character(TF),'')),hjust=-0.2,vjust=0) +
  ggrepel::geom_label_repel(aes(label=ifelse(Rank<=20,as.character(TF),'')),
                  box.padding   = 0.35, 
                  point.padding = 0.5,
                  nudge_x = 2,
                  segment.color = 'grey50', 
                  max.overlaps = Inf) +
  theme_bw() +
  ggtitle("Degree of TFs regulating DEGs in AD vs control - all cells")

ggsave(paste0(fig_dir, "AD_GRN_fromDEGs_ALL_degreeRankPlot.png"))
ggsave(paste0(fig_dir, "AD_GRN_fromDEGs_ALL_degreeRankPlot.svg"))
```


```{r}
# - converted to . in SCENIC analysis, convert metadata as well

row.names(metaData) <- gsub("-", "\\.", row.names(metaData))
head(metaData)

dim(metaData)
metaData <- metaData[ps_cells, ]
dim(metaData)
```


First checkpoint for evaluating the outcome of SCENIC: did it find regulons?
```{r}
dim(regulonAUC)
length(thresholds)
length(regulons)
```

```{r}
head(regulonAUC)
head(thresholds)
```

Save regulon matrix
```{r}
toExport<-as.data.frame(t(regulons_incidMat), stringsAsFactors=F)
toExport$gene<-rownames(toExport)
toExport<-toExport[,c('gene',setdiff(colnames(toExport),"gene"))]
tmpList<-list(toExport)
names(tmpList)<-"regulonMatrix"
write.xlsx(tmpList,paste0(table_dir, "tableRegulons_AD_large.xlsx"))
```


FIX THRESHOLDS: names and values switched in output. pySCENIC/loom bug?

```{r}
thresholds_values <- names(thresholds)
thresholds_names <- thresholds
thresholds <- thresholds_values
names(thresholds) <- thresholds_names
length(thresholds)
length(regulons)
head(thresholds)
```

Normally the thresholds and regulons should have the same dimensions: I will filter the regulons not in common out

```{r}
sum(names(regulons) %in% names(thresholds))
which(names(thresholds) %in% names(regulons)) 
thresholds <- thresholds[which(names(thresholds) %in% names(regulons))]
length(thresholds)
length(regulons)
head(thresholds)
```

Get AUC values in matrix + scale to range 0-1 for nicer heatmaps

```{r}
AUCdata <- getAUC(regulonAUC) %>% BBmisc::normalize("range") #by default, operates on rows (regulons)
dim(AUCdata)
AUCdata[1:5,1:5]
```

# Define regulons per cell

```{r}
regulonsCells <- setNames(lapply(names(thresholds), 
                                 function(x) {
                                   trh <- thresholds[x]
                                   names(which(getAUC(regulonAUC)[x,]>trh))
                                 }),names(thresholds))

length(regulonsCells)


regulonActivity <- reshape2::melt(regulonsCells)
head(regulonActivity)
```


# Create binary activity (regulon On or Off based on thresholds)

```{r}
binaryRegulonActivity <- t(table(regulonActivity[,1], regulonActivity[,2])) #table can also build matrices
class(binaryRegulonActivity) <- "matrix"
dim(binaryRegulonActivity)
binaryRegulonActivity[1:5,1:5]
```


# Only non-duplicated regulons

```{r}
if (sum(duplicated(rownames(binaryRegulonActivity))) != 0){
  
  binaryRegulonActivity_nonDupl <- binaryRegulonActivity[which(rownames(binaryRegulonActivity) %in% onlyNonDuplicatedExtended(rownames(binaryRegulonActivity))),]
  dim(binaryRegulonActivity_nonDupl)
  
} else {
  print("No duplicated regulons... Results from RegulonActivity_nonDupl and binaryRegulonActivity will be the same")
  binaryRegulonActivity_nonDupl = binaryRegulonActivity
}

```

# Get minimum cells

I lowered the amount of minamal cells where a regulon is active in this case, it would be 143 cells. Which is probably too high. 

```{r}
minCells <- ncol(binaryRegulonActivity) * .001 # 1% cells
```

### All regulons

```{r}
regulonSelection <- list()
regulonSelection[["All_regulons_with_duplicated"]] <- rownames(binaryRegulonActivity)
length(rownames(binaryRegulonActivity))
```

### Active in > 1% cells

The binary activity is one ("on") in more than 1% of the cells for this regulon. 
FROM the supplementary in the SCENIC paper : "For example, in Fig. 1b, we have plotted only regulons active in at least 1% of the cells and correlated with other regulons in the matrix (absolute correlation > 0.30). However, the downstream analyses include all the regulons.

```{r}
regMinCells <- names(which(rowSums(binaryRegulonActivity_nonDupl) > minCells))
regulonSelection[["Regs_active_morethan_1perc_cells"]] <- regMinCells
length(regMinCells)
```

### Correlation between regulons
```{r}
reguCor <- cor(t(binaryRegulonActivity_nonDupl[regMinCells,]))
diag(reguCor) <- 0
nrow(reguCor)
```


### Regulons correlated with other regulons and active in > 1% cells

```{r}
corrRegs <- names(which(rowSums(abs(reguCor) > 0.3) > 0)) 
regulonSelection[["Regs_active_and_corr_0.3"]]  <- corrRegs
length(corrRegs)
```

```{r}
#Extra: add corr>0.5
corrRegs_05 <- names(which(rowSums(abs(reguCor) > 0.5) > 0)) 
regulonSelection[["Regs_active_and_corr_0.5"]]  <- corrRegs_05
length(corrRegs_05)
```



### Regulons NOT correlated with other regulons and/or active in < 1% cells

```{r}
missingRegs <- rownames(binaryRegulonActivity_nonDupl)[which(!rownames(binaryRegulonActivity_nonDupl) %in% corrRegs)]
regulonSelection[["Regs_inactive_or_noncorr"]]  <- missingRegs
length(missingRegs)
```


```{r}
# Save so you can also run plots on HPC

saveRDS(AUCdata,"Tables/AUCdata_AD_large.rds")
saveRDS(regulonSelection, "Tables/RegSelections_AD_large.rds")
```


## Prepare plots

### UMAP

```{r}
p1<-ggplot(metaData, aes(UMAP_1, UMAP_2, color=celltype)) +
 geom_point(size=0.05) +
  theme_bw()

metaData<- cbind(metaData, t(AUCdata[regulonSelection[['Regs_active_and_corr_0.5']],row.names(metaData)]))

p2 <- ggplot(metaData, aes(UMAP_1, UMAP_2, color=`IRF8_(+)-motif`)) +
 geom_point(size=0.05) +
 scale_color_gradient2(low="blue", mid = "white", high="red", midpoint=0.5, space="Lab") +
  theme_bw()

p1 + p2
```

### Heatmaps

#### Aggregated heatmaps: cell types as groups

```{r}
v <- metaData$celltype
names(v) <- row.names(metaData)
idents <- sort(v)
ordered <- names(idents)

nrClust <- length(levels(as.factor(v)))
nrClust
```


```{r}
colors <- c("#ED8141")
colors <- colorRampPalette(colors)(nrClust)
names(colors) <- unique(metaData$celltype)
colors <- list(idents = colors)

tables <- table(v)
names(tables) <- NULL
```


### Setting gaps

```{r}
gaps <- tables[1]
for (k in 1:nrClust) { #add values together
  value <- tables[k]
  gaps <- append(gaps, value+gaps[k-1])
}


for (i in seq_len(length(regulonSelection))) {
  selRegs <- names(regulonSelection)[i]
  if(length(regulonSelection[[selRegs]])>1) {
    
    AUCMat <- AUCdata[regulonSelection[[selRegs]], ,drop=FALSE]
    AUCMat <- AUCMat[,ordered]
    
    avg_exp_df <- data.frame(row.names = regulonSelection[[selRegs]]) #Create empty dataframe
    
    for (j in levels(as.factor(v))) {
      AUCMatTmp <- AUCMat[,names(as.factor(v)[which(v==j)])] #Subset for cell population
      AUCMatTmp <- as.matrix(AUCMatTmp)
      
      if (dim(AUCMatTmp)[2] == 1) {
        average_expressions <- AUCMatTmp
      } else {
        average_expressions <- apply(AUCMatTmp, 1, function(x){ #apply across row: average activity in regulon for that cell pop
          sum(x)/length(colnames(AUCMatTmp))
        })
      }
      
      average_expressions <- as.numeric(average_expressions)
      avg_exp_df[,j] <- average_expressions
    }
    
    names(avg_exp_df) <- levels(as.factor(v))
    AUCMatT <- avg_exp_df
    
    p <- pheatmap(as.matrix(AUCMatT), cluster_cols = F, cluster_rows = F, scale = "none", 
                  main = selRegs, treeheight_row = 0, 
                  cellwidth = 10, cellheight = 10, 
                  show_colnames = T, border_color = "black",
                  fontsize_row = 4, fontsize_col = 4, angle_col = "315")
    
    file_name <- paste0(fig_dir, "Heatmap_AUCRegulons_Averaged_", selRegs ,"_ADlarge.pdf")
    
    p.height <- 1 + 0.15 * length(rownames(AUCMatT))
    
    pdf(file = file_name, 
        width = 10, 
        height = p.height, 
    ) 
    
    draw(p, merge_legend = T)
    
    dev.off()
  }
}

```

#### Aggregated heatmaps: diagnosis as group

```{r}
v <- metaData$Diagnosis
names(v) <- row.names(metaData)
idents <- sort(v)
ordered <- names(idents)

nrClust <- length(levels(as.factor(v)))
nrClust
```


```{r}
colors <- c("#ED8141")
colors <- colorRampPalette(colors)(nrClust)
names(colors) <- unique(metaData$Diagnosis)
colors <- list(idents = colors)

tables <- table(v)
names(tables) <- NULL
```


### Setting gaps

```{r}
gaps <- tables[1]
for (k in 1:nrClust) { #add values together
  value <- tables[k]
  gaps <- append(gaps, value+gaps[k-1])
}


for (i in seq_len(length(regulonSelection))) {
  selRegs <- names(regulonSelection)[i]
  if(length(regulonSelection[[selRegs]])>1) {
    
    AUCMat <- AUCdata[regulonSelection[[selRegs]], ,drop=FALSE]
    AUCMat <- AUCMat[,ordered]
    
    avg_exp_df <- data.frame(row.names = regulonSelection[[selRegs]]) #Create empty dataframe
    
    for (j in levels(as.factor(v))) {
      AUCMatTmp <- AUCMat[,names(as.factor(v)[which(v==j)])] #Subset for cell population
      AUCMatTmp <- as.matrix(AUCMatTmp)
      
      if (dim(AUCMatTmp)[2] == 1) {
        average_expressions <- AUCMatTmp
      } else {
        average_expressions <- apply(AUCMatTmp, 1, function(x){ #apply across row: average activity in regulon for that cell pop
          sum(x)/length(colnames(AUCMatTmp))
        })
      }
      
      average_expressions <- as.numeric(average_expressions)
      avg_exp_df[,j] <- average_expressions
    }
    
    names(avg_exp_df) <- levels(as.factor(v))
    AUCMatT <- avg_exp_df
    
    p <- pheatmap(as.matrix(AUCMatT), cluster_cols = F, cluster_rows = F, scale = "none", 
                  main = selRegs, treeheight_row = 0, 
                  cellwidth = 10, cellheight = 10, 
                  show_colnames = T, border_color = "black",
                  fontsize_row = 4, fontsize_col = 4, angle_col = "315")
    
    file_name <- paste0(fig_dir, "Heatmap_AUCRegulons_Averaged_Diagnosis_", selRegs ,"_ADlarge.pdf")
    
    p.height <- 1 + 0.15 * length(rownames(AUCMatT))
    
    pdf(file = file_name, 
        width = 10, 
        height = p.height, 
    ) 
    
    draw(p, merge_legend = T)
    
    dev.off()
  }
}

```

#### Complex heatmap

Create annotations for ComplexHeatmap
```{r}
cols_celltype <- brewer.pal(length(levels(as.factor(metaData$celltype))), "Set1")
names(cols_celltype) <- levels(as.factor(metaData$celltype))

col_ann_types = columnAnnotation( Celltype = metaData[colnames(AUCdata), 'celltype'], show_legend = T, annotation_legend_param = list(Celltype = list(title = "Celltype")),
                            show_annotation_name = T, annotation_label = "Celltype", 
                            col = list(Celltype = cols_celltype))
```


```{r}
cols_diagnosis <- brewer.pal(length(levels(as.factor(metaData$Diagnosis))), "Set2")[c(2,1)]
names(cols_diagnosis) <- levels(as.factor(metaData$Diagnosis))

col_ann_diag = columnAnnotation( Diagnosis = metaData[colnames(AUCdata), 'Diagnosis'], show_legend = T, annotation_legend_param = list(Diagnosis = list(title = "Diagnosis")),
                            show_annotation_name = T, annotation_label = "Diagnosis", 
                            col = list(Diagnosis = cols_diagnosis))

```


```{r}
for(i in seq_len(length(regulonSelection))){ 
  
  message("Print heatmap ",i)
  selRegs <- names(regulonSelection)[i]
  
  if(length(regulonSelection[[selRegs]])>1) {
    
    AUCMat <- AUCdata[regulonSelection[[selRegs]],,drop=FALSE]

    col_fun = colorRamp2(c(0, 0.5, 1), c("steelblue3", "#FFFFBF", "#D73027")) 
    col_fun(seq(-50, 50))
    
    hm <- Heatmap(AUCMat, show_column_names = FALSE, name="AUCell score", 
                    clustering_distance_rows = "euclidean", 
                    clustering_distance_columns = "euclidean", 
                    clustering_method_rows = "ward.D", 
                    clustering_method_columns = "ward.D", 
                    column_dend_height = unit(1.5, "cm"), row_dend_width = unit(1.5, "cm"), 
                    width = unit(15, "cm"), height = unit((0.07*dim(AUCMat)[1]), "cm"), 
                    row_names_gp = gpar(fontsize = 2), column_names_gp = gpar(fontsize = 1),
                    col=col_fun, top_annotation = c(col_ann_types, col_ann_diag)
    )
    
    pdf(paste0(fig_dir, "RegulonActivity_Heatmap_",selRegs,"_ComplexHeatmap_ADlarge.pdf"), width=12.45, height=17.55)
    draw(hm, merge_legend = TRUE)  
    dev.off()
    
    
  }
}
```


## Test TFs of interest

```{r}
AUCdata <- readRDS("Tables/AUCdata_AD_large.rds")
dim(AUCdata)
```


```{r}
#test

AUCdata <- assay(regulonAUC)[1,]
for (i in 2:nrow(regulonAUC)) {
  AUCdata <- rbind(AUCdata, assay(regulonAUC)[i,])
}

row.names(AUCdata) <- row.names(regulonAUC)
colnames(AUCdata) <- gsub("\\.", "-", colnames(AUCdata))
dim(AUCdata)
```



```{r}
metaData <- metaData[colnames(AUCdata),]
metaData <- cbind(metaData, t(AUCdata))
metaData
```

```{r}
ggplot(metaData, aes(x=Diagnosis, y=`IRF8_(+)-motif`, fill=Diagnosis)) +
  geom_violin() + 
  theme_bw()
```
Create Seurat object for for convenience in plotting

```{r}
library(Seurat)
regdata <- t(metaData[,- (1:9)])
row.names(regdata) <- substr(row.names(regdata), 1, nchar(row.names(regdata)) - 10 )

so <- CreateSeuratObject(regdata, assay = "Regulons", meta.data = metaData[,1:5])
so
```



```{r}
RidgePlot(so, features = c("IKZF1", "IRF8", "NFATC2", "MAF", "RUNX1", "TAL1", "SPI1", "MAFF", "STAT3", "MYC", "FOS", "JUNB", "CEBPD", "PLSCR1"), ncol = 3, group.by = "Diagnosis")
```

```{r}
library(presto)
```

```{r}
### Focus tests on MICORGLIA and save results
wilc<- wilcoxauc(so, group_by = "Diagnosis", seurat_assay = "Regulons", assay = "counts")
head(wilc)
```


```{r}
# Subset to microglia

so@meta.data<- metaData[,1:9]

so <- SetIdent(so, value="celltype")

so_sub <- subset(so, idents="MG")
so_sub
```

```{r}
wilc_mg<- wilcoxauc(so_sub, group_by = "Diagnosis", seurat_assay = "Regulons", assay = "counts")
head(wilc_mg)
```

```{r}
wilc_mg <- wilc_mg[wilc_mg$group == "AD" & wilc_mg$padj < 0.05,]
write.csv(wilc_mg[order(wilc_mg$logFC, decreasing = T),], paste0(table_dir, "TestWilc_MG_ADDlarge.csv"), quote = F)
```


```{r}
tfs_oi <- c("IKZF1", "IRF8", "NFATC2", "MAF", "RUNX1", "TAL1", "SPI1", "MAFF", "STAT3", "MYC", "FOS", "JUNB", "CEBPD", "PLSCR1")

wilc[wilc$feature %in% tfs_oi, ]
```

```{r}
wilc[order(wilc$logFC, decreasing = T),]
```

```{r}
VlnPlot(so, features = tfs_oi, split.by = "Diagnosis")
```

```{r}
tfs_oi <- c(paste0(tfs_oi, "_(+)-motif"))
selec_tfs <- tfs_oi[tfs_oi %in% colnames(metaData)]

df_plot <- metaData[, c(selec_tfs, "Diagnosis") ]
head(df_plot)
```

```{r}
df_plot$Cell <- row.names(df_plot)
df_plot <- melt(df_plot)
colnames(df_plot)[3:4] <- c("Regulon", "Activity")
head(df_plot)
```

```{r}
ggplot(df_plot, aes(x=Regulon, y=Activity, fill=Diagnosis)) +
  geom_boxplot(outlier.size = 0.5) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))

ggsave(paste0(fig_dir, "Boxplot_ADlarge_TFsOI_byCondition.png"))
ggsave(paste0(fig_dir, "Boxplot_ADlarge_TFsOI_byCondition.svg"))
ggsave(paste0(fig_dir, "Boxplot_ADlarge_TFsOI_byCondition.pdf"))
```

```{r}
for (ct in levels(as.factor(metaData$celltype))){
  p<- ggplot(df_plot[df_plot$Cell %in% row.names(metaData[metaData$celltype == ct,]), ], aes(x=Regulon, y=Activity, fill=Diagnosis)) +
    geom_boxplot(outlier.size = 0.5) +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
  
  ggsave(paste0(fig_dir, "Boxplot_ADlarge_TFsOI_", ct,"_byCondition.png"), p)
  ggsave(paste0(fig_dir, "Boxplot_ADlarge_TFsOI_", ct,"_byCondition.svg"), p)
  ggsave(paste0(fig_dir, "Boxplot_ADlarge_TFsOI_", ct,"_byCondition.pdf"), p)
}
```

