---
title: "Preprocessing2"
author: "Hanne Puype"
date: "8 november 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

```{r}
library(tidyverse)
library(edgeR)
library(biomaRt)
```

# Datasets
## Alzheimer's disease
GSE174367
```{r}
counts_AD <- read.table("data_HTSC_CTX.txt", header = T, sep = '\t', stringsAsFactors = FALSE, check.names = FALSE)
counts_AD <- rownames_to_column(counts_AD)
colnames(counts_AD)[names(counts_AD) == "rowname"] <- "GeneID"
#head(counts_AD)
dim(counts_AD)
```

```{r}
anno_AD <- read.csv("targets.CTX.csv")
#head(anno_AD)
#tail(anno_AD)
dim(anno_AD)
colnames(anno_AD)[names(anno_AD) == "SampleID"] <- "Sample"
```

Keep metadata: Sample, Neuropath.DX.1, Sex, Tangle.Stage, Plaque.Stage, Diagnosis

How many AD and controls?
```{r}
anno_AD %>%
  group_by(Diagnosis) %>%
  summarise(n())
anno_AD %>%
  group_by(Neuropath.Dx.1) %>%
  summarise(n())
anno_AD %>% 
  group_by(Sex) %>%
  summarise(n())
anno_AD %>% 
  group_by(Tangle.Stage) %>%
  summarise(n())
anno_AD %>% 
  group_by(Plaque.Stage) %>%
  summarise(n())
```

## Depression
GSE101521
```{r}
counts_MDD1 <- read.table("GSE101521_totalRNA_counts.csv.gz", header = T, sep = ',', stringsAsFactors = FALSE, 
                           check.names = FALSE)
colnames(counts_MDD1)[names(counts_MDD1) == ""] <- "GeneID"
counts_MDD1 <- as_tibble(counts_MDD1)
#head(counts_MDD1)
dim(counts_MDD1)
```

```{r}
anno_MDD1 <- read.table("SraRunTable_GSE101521.txt", header = T, sep = ',', stringsAsFactors = FALSE, check.names = 
                          FALSE)
anno_MDD1 <- as_tibble(anno_MDD1)
anno_MDD1 <- anno_MDD1 %>% 
  add_column(Sample = c("V1", "V2", "V3", "V4", "V5", "V6", "V7", "V8", "V9", "V10", "V11", "V12", "V13", "V14", "V15",
                        "V16", "V17", "V18", "V19", "V20", "V21", "V22", "V23", "V24", "V25", "V26", "V27", "V28",
                        "V29", "V30", "V31", "V32", "V33", "V34", "V35", "V36", "V37", "V38", "V39", "V40", "V41",
                        "V42", "V43", "V44", "V45", "V46", "V47", "V48", "V49", "V50", "V51", "V52", "V53", "V54",
                        "V55", "V56", "V57", "V58", "V59"))
#head(anno_MDD1)
```

Metadata: keep Diagnosis, sex, Sample

```{r}
anno_MDD1 %>%
  group_by(sex) %>%
  summarise(n())
anno_MDD1 %>%
  group_by(Diagnosis) %>%
  summarise(n())
```

GSE80655
```{r}
counts_MDD2 <- read.table("GSE80655_GeneExpressionData_Updated_3-26-2018.txt.gz", header = T, sep = '\t', 
                               stringsAsFactors = FALSE, check.names = FALSE)
dim(counts_MDD2)
#tail(counts_MDD2)
colnames(counts_MDD2)[names(counts_MDD2) == "gene_id"] <- "GeneID"
counts_MDD2 <- counts_MDD2 %>% dplyr::select(GeneID, SL32230, SL32231, SL32242, SL32243, SL32244, SL32245, SL32246, 
                                             SL32247, SL32248, SL32249, SL5886, SL5997, SL5998, SL5999, SL6000, SL6128, 
                                             SL6129, SL6130, SL6340, SL6341, SL6412, SL6413, SL6609, SL6612, SL6643, 
                                             SL6645, SL6666, SL6668, SL6669, SL7141, SL7152, SL7154, SL7155, SL7381, 
                                             SL7383, SL7384, SL7393, SL7403, SL7404, SL7614, SL7625, SL7632, SL7698, 
                                             SL7801, SL7804, SL7806, SL7807)
counts_MDD2 <- counts_MDD2 %>% arrange(GeneID)
#head(counts_MDD2)
```

```{r}
anno_MDD2 <- read.table("SraRunTable_GSE80655.txt", header = T, sep = ',', stringsAsFactors = FALSE, check.names = 
                          FALSE)
colnames(anno_MDD2)[names(anno_MDD2) == "clinical_diagnosis"] <- "Diagnosis"
#head(anno_MDD2)
anno_MDD2 <- anno_MDD2 %>% dplyr::slice(-2, -4,-6, -8, -10, -12, -14, -16, -18, -20, -29, -31, -58, -60)
dim(anno_MDD2)
anno_MDD2 <- anno_MDD2 %>% add_column(Sample = c("SL32230", "SL32231", "SL32242", "SL32243", "SL32244", "SL32245", 
                                                 "SL32246", "SL32247", "SL32248", "SL32249", "SL5886", "SL5997", 
                                                 "SL5998", "SL5999", "SL6000", "SL6128", "SL6129", "SL6130", "SL6340", 
                                                 "SL6341", "SL6412", "SL6413", "SL6609", "SL6612", "SL6643", "SL6645", 
                                                 "SL6666", "SL6668", "SL6669", "SL7141", "SL7152", "SL7154", "SL7155", 
                                                 "SL7381", "SL7383", "SL7384", "SL7393", "SL7403", "SL7404", "SL7614", 
                                                 "SL7625", "SL7632", "SL7698", "SL7801", "SL7804", "SL7806", "SL7807"))
#tail(anno_MDD2)
```

Keep metadata: Diagnosis, ETHNICITY, gender, Sample

```{r}
anno_MDD2 %>%
  group_by(Diagnosis) %>%
  summarise(n())
anno_MDD2 %>%
  group_by(gender) %>%
  summarise(n())
anno_MDD2 %>%
  group_by(ETHNICITY) %>%
  summarise(n())
```

Keeping ethnicity won't really give extra information because there are only 3 non-white people.

## Merging datasets and batch correction
ALzheimer 
```{r}
#head(counts_AD)
counts_AD$GeneID <- gsub('\\..+$', '', counts_AD$GeneID)
counts_AD[duplicated(counts_AD)]
```

Depression
```{r}
#head(counts_MDD1)
#head(counts_MDD2)
counts_MDD <- inner_join(counts_MDD1, counts_MDD2, by = "GeneID")
#head(counts_MDD)
dim(counts_MDD)
```

Annotation datasets
```{r}
anno_AD <- anno_AD %>% dplyr::select(Sample, Diagnosis, Sex, everything())
anno_AD <- anno_AD %>% mutate(Sex = case_when(Sex == "M" ~ "male", T ~ "female"))
#head(anno_AD)
anno_MDD1 <- anno_MDD1 %>% dplyr::select(Sample, Diagnosis, sex, everything())
colnames(anno_MDD1)[names(anno_MDD1) == "sex"] <- "Sex"
#head(anno_MDD1)
anno_MDD2 <- anno_MDD2 %>% dplyr::select(Sample, Diagnosis, gender, everything())
colnames(anno_MDD2)[names(anno_MDD2) == "gender"] <- "Sex"
#head(anno_MDD2)
anno_MDD2 %>%
  group_by(`SRA Study`) %>%
  summarise(n())
```

```{r}
anno_MDD1 <- anno_MDD1 %>% mutate(Diagnosis = case_when(Diagnosis == "non-psychiatric controls (CON)" ~ "Control", T ~ 
                                                          Diagnosis))
anno_MDD1 <- anno_MDD1 %>% dplyr::select(Sample, Diagnosis, Sex, BioProject)
anno_MDD2 <- anno_MDD2 %>% dplyr::select(Sample, Diagnosis, Sex, BioProject)
anno_MDD <- rbind(anno_MDD1, anno_MDD2)
#head(anno_MDD)
anno_MDD <- anno_MDD %>% mutate(Diagnosis = case_when(Diagnosis == "DSM-IV major depressive disorder suicides (MDD-S)" ~ "Major depression suicide", 
                                                      Diagnosis == "DSM-IV major depressive disorder non-suicides (MDD)" ~ "Major depression non-suicide", 
                                                      T ~ Diagnosis))
anno_MDD %>%
  group_by(Diagnosis) %>%
  summarise(n())

anno_AD <- anno_AD %>% dplyr::select(Sample, Diagnosis, Sex, Neuropath.Dx.1, Tangle.Stage, Plaque.Stage)
#head(anno_AD)
anno_AD %>%
  group_by(Neuropath.Dx.1) %>%
  summarise(n())
```


# Preprocessing
## Normalization and logarithmic transformation with EdgeR
```{r}
preprocess <- function(C, A) {
groupExp = factor(A$Diagnosis[A$Sample %in% names(C)])
CC <- as.matrix(C[,-1])
rownames(CC) <- C$GeneID
y <- DGEList(counts = CC, group = groupExp)
y_ori <- y
keep <- rowSums(cpm(y) > 1) >= 5 #more than 1 count in at least 5 samples
y <- y[keep, , keep.lib.sizes=FALSE]
y <- calcNormFactors(y) #TMM normalization for lib sizes

name_plot = deparse(substitute(C)) #from variable to variable name as a string
name_1 = paste0('mds_',name_plot, '.png')
png(filename = name_1)
plotMDS(y, col = c("blue", "red", "green")[groupExp]) 
dev.off()

logcpm <- cpm(y, log = T, prior.count = 1)
logcpmS <- as_tibble(rownames_to_column(as.data.frame(logcpm), "GeneID"))

name_2 = paste0('boxplot_',name_plot, '.png')
png(filename = name_2)
boxplot(logcpm, main = "Filtered_Normalized", ylab = "log-cpm", xaxt = "n")
axis(side = 1, labels = FALSE)
text(x = 1:length(colnames(logcpm)), y = par("usr")[3]-0.5, labels = colnames(logcpm), srt = 45, adj = 1, xpd = NA, cex 
     = 0.8)
dev.off() 
name_3 = paste0('mds_afterlog_',name_plot, '.png')
png(filename = name_3)
plotMDS(logcpm, col = c("blue", "red", "green")[groupExp]) 
dev.off()

return(logcpmS)
}
```


```{r}
procounts_AD_N <- preprocess(counts_AD, anno_AD)
procounts_MDD_N <- preprocess(counts_MDD, anno_MDD)
#head(procounts_AD_N)
#head(procounts_MDD_N)
```

```{r}
dim(procounts_AD_N) #no genes gone
dim(procounts_MDD_N) #33114 genes gone
```

## Batch correction after normalization
```{r}
#head(procounts_MDD_N)
GeneID <- procounts_MDD_N$GeneID
procounts_MDD_b <- procounts_MDD_N %>% dplyr::select(-GeneID)
procounts_MDD_b <- procounts_MDD_b %>% mutate_if(is.integer, as.numeric)
procounts_MDD_b <- as.matrix(procounts_MDD_b)
procounts_MDD_N <- removeBatchEffect(procounts_MDD_b, anno_MDD$BioProject, covariates = NULL)
procounts_MDD_N <- as_tibble(procounts_MDD_N)
procounts_MDD_N <- procounts_MDD_N %>% add_column(GeneID)
procounts_MDD_N <- procounts_MDD_N %>% dplyr::select(GeneID, everything())
#head(procounts_MDD_N)
```

PCAplot MDD
```{r}
pca <- prcomp(procounts_MDD_N[, -1])
anno_MDD %>%
  mutate(PC1 = pca$rotation[,1], PC2 = pca$rotation[,2]) %>%
  ggplot(aes(x = PC1, y = PC2, colour = Diagnosis)) +
        geom_point() + 
  geom_text(label = colnames(procounts_MDD_N[, -1]), check_overlap = F)
```

```{r}
png(filename = 'mds_afterlog_procounts_MDD.png')
plotMDS(procounts_MDD_N[, -1], col = c("blue", "red", "green")[factor(anno_MDD$Diagnosis)]) 
dev.off()
```

## Removal of outliers
Transpose datasets
```{r}
data_AD_t <- procounts_AD_N %>% 
    pivot_longer(!c('GeneID'), names_to = 'Samples', values_to = 'logcpm') %>% #names_to: column with all column names (samples) and values_to: all counts from these samples
    pivot_wider(names_from = c('GeneID'), values_from = 'logcpm') #transposed: samples in rows, genes in columns
mat_AD <- as.matrix(data_AD_t[,-1])
rownames(mat_AD) <- data_AD_t$Samples

data_MDD_t <- procounts_MDD_N %>% 
    pivot_longer(!c('GeneID'), names_to = 'Samples', values_to = 'logcpm') %>% 
    pivot_wider(names_from = c('GeneID'), values_from = 'logcpm')
mat_MDD <- as.matrix(data_MDD_t[,-1])
rownames(mat_MDD) <- data_MDD_t$Samples
```

```{r}
hclust_AD <- hclust(dist(mat_AD[, -1]), method = "average")
plot(hclust_AD, main = "Cluster dendogram AD", xlab = "", sub = "", cex = 0.7)
```

I am not sure whether there are outliers. The six/seven samples at the left could be. Sample18, 96, 27, 87, 41, 9

```{r}
hclust_MDD <- hclust(dist(mat_MDD[, -1]), method = "average")
plot(hclust_MDD, main = "Cluster dendogram MDD", xlab = "", sub = "", cex = 0.7)
```

V10 is a clear outlier.

```{r}
procounts_MDD_N <- procounts_MDD_N %>% dplyr::select(-V10)
anno_MDD <- anno_MDD %>% dplyr::slice(-10)
#head(anno_MDD, 10)
```

I will do PCA of the AD dataset to see whether I will delete these samples. 
```{r}
pca <- prcomp(procounts_AD_N[, -1])
anno_AD %>%
  mutate(PC1 = pca$rotation[,1], PC2 = pca$rotation[,2]) %>%
  ggplot(aes(x = PC1, y = PC2, colour = Diagnosis)) +
        geom_point() + 
  geom_text(label = colnames(procounts_AD_N[, -1]), check_overlap = F)
```

Here as well samples 9, 18, 27, 41, 87 and 96 form a distinct cluster in the PCA plot. 
This is the case as well in the MDS plot. 

```{r}
procounts_AD_N <- procounts_AD_N %>% dplyr::select(-c("Sample-9", "Sample-18", "Sample-27", "Sample-41", "Sample-87", 
                                                      "Sample-96"))
#head(procounts_AD_N)
anno_AD <- anno_AD %>% dplyr::slice(-c(11, 21, 35, 82, 85, 92))
```

Save annotation files 
```{r}
write.table(anno_MDD, file = "annotation_MDD.txt", sep = "\t", col.names = T, row.names = F, quote = F)
write.table(anno_AD, file = "annotation_AD.txt", sep = "\t", col.names = T, row.names = F, quote = F)
```

## Selection of highly variable genes 
```{r}
#remove the geneId column 
procounts_AD_E <- as.data.frame(procounts_AD_N, row.names = GeneID)
procounts_MDD_E <- as.data.frame(procounts_MDD_N, row.names = GeneID)

# apply() takes Data frame or matrix as an input and gives output in vector, list or array.
# var, calcualte rowwise variance
vars_AD <- apply(procounts_AD_E[, -1], 1, var) #1 is rows, 2 columns
vars_MDD <- apply(procounts_MDD_E[, -1], 1, var)

# create histogram
hist_AD <- hist(vars_AD, 200, plot = F)
hist_MDD <- hist(vars_MDD, 500, plot = F)
```

Plot histogram and draw lines
```{r}
png(filename = "Histogram_highlyvariable_AD.png")
plot(hist_AD,
     col = 'gold',
     xlim = c(0,2),
     main = "Histogram of vars (500 breaks)",
     ylim = c(0,3500))+
  abline(v = c(0.20, 1, 0.45, 0.5, 1.75), col = c("red", "orange", "darkblue", "green", "pink"), lty = 2, lwd = 3)
dev.off()
```

0.1 seems good. 

```{r}
png(filename = "Histogram_highlyvariable_MDD.png")
plot(hist_MDD,
     col = 'gold',
     xlim = c(0,3),
     main = "Histogram of vars (500 breaks)",
     ylim = c(0,5000))+
  abline(v = c(0.20, 1, 0.45, 0.5, 1.75), col = c("red", "orange", "darkblue", "green", "pink"), lty = 2, lwd = 3)
dev.off()
```

0.2 seems good.

```{r}
names(vars_AD) <- procounts_AD_E$GeneID
length(names(vars_AD[vars_AD >= 0.1])) #13970 genes -> 4896 gone 
length(names(vars_AD[vars_AD >= 0.2])) #8596 
length(names(vars_AD[vars_AD >= 0.5])) #1714
names(vars_MDD) <- procounts_MDD_N$GeneID
length(names(vars_MDD[vars_MDD >= 0.2])) #11012 -> 11458 gone
length(names(vars_MDD[vars_MDD >= 0.1])) #15911
```

```{r}
print("Number of features remaing after variance filter:")
print("AD dataset:")
names_AD <- names(vars_AD[vars_AD >= 0.1])
procounts_AD <- procounts_AD_N %>%
    dplyr::filter(GeneID %in% names_AD)
dim(procounts_AD)
print("MDD dataset:")
names_MDD <- names(vars_MDD[vars_MDD >= 0.1])
procounts_MDD <- procounts_MDD_N %>%
    dplyr::filter(GeneID %in% names_MDD)
dim(procounts_MDD)
```

## From GeneID to GeneSymbol
```{r}
#BiocManager::install("EnsDb.Hsapiens.v79")
library(EnsDb.Hsapiens.v79)
ensembl.genes <- procounts_AD$GeneID
geneIDs1 <- ensembldb::select(EnsDb.Hsapiens.v79, keys = ensembl.genes, keytype = "GENEID", columns = 
                                c("SYMBOL","GENEID"))
head(geneIDs1)
dim(geneIDs1)
geneIDs1 <- geneIDs1 %>% mutate(SYMBOL = case_when(SYMBOL == "" ~ GENEID, T ~ SYMBOL))
#tail(geneIDs1)

colnames(geneIDs1)[names(geneIDs1) == "GENEID"] <- "GeneID"
procounts_AD <- left_join(geneIDs1, procounts_AD, by = "GeneID")
procounts_AD <- procounts_AD %>% dplyr::select(GeneID, SYMBOL, everything())
colnames(procounts_AD)[names(procounts_AD) == "SYMBOL"] <- "GeneSymbol"
#head(procounts_AD)
dim(procounts_AD)
```

Protein-coding genes only
```{r}
mart <- useMart(biomart = "ENSEMBL_MART_ENSEMBL", dataset = "hsapiens_gene_ensembl", host = 'www.ensembl.org')
genes_ENSID <- biomaRt::getBM(attributes = c("external_gene_name", "ensembl_gene_id", "chromosome_name", 
                                             "transcript_biotype"), filters = c("transcript_biotype", 
                                                                                "chromosome_name"), values = 
                                list("protein_coding", c(1:22)), mart = mart)
procounts_AD <- procounts_AD %>%
  dplyr::filter(GeneID %in% genes_ENSID$ensembl_gene_id)
dim(procounts_AD) #4760 genes gone 
```

```{r}
procounts_AD$GeneSymbol[duplicated(procounts_AD$GeneSymbol)]
procounts_AD %>% dplyr::filter(GeneSymbol == "CBS")
procounts_AD <- procounts_AD %>% mutate(GeneSymbol = case_when(GeneID == "ENSG00000274276" ~ "CBSL", T ~ GeneSymbol))
procounts_AD %>% dplyr::filter(GeneSymbol == "ZNF709")
procounts_AD <- procounts_AD %>% mutate(GeneSymbol = case_when(GeneID == "ENSG00000196826" ~ "ENSG00000196826", T ~ GeneSymbol))
procounts_AD %>% dplyr::filter(GeneSymbol == "FAM47E-STBD1")
procounts_AD <- procounts_AD %>% mutate(GeneSymbol = case_when(GeneID == "ENSG00000118804" ~ "STBD1", T ~ GeneSymbol))
procounts_AD %>% dplyr::filter(GeneSymbol == "SMIM11")
procounts_AD <- procounts_AD %>% mutate(GeneSymbol = case_when(GeneID == "ENSG00000205670" ~ "SMIM11A", GeneID == "ENSG00000273590" ~ "SMIM11B", T ~ GeneSymbol))
procounts_AD %>% dplyr::filter(GeneSymbol == "TAF9")
procounts_AD <- procounts_AD %>% mutate(GeneSymbol = case_when(GeneID == "ENSG00000085231" ~ "AK6", T ~ GeneSymbol))
procounts_AD %>% dplyr::filter(GeneSymbol == "U2AF1")
procounts_AD <- procounts_AD %>% mutate(GeneSymbol = case_when(GeneID == "ENSG00000275895" ~ "U2AF1L5", T ~ GeneSymbol))
procounts_AD %>% dplyr::filter(GeneSymbol == "ICOSLG")
procounts_AD <- procounts_AD %>% mutate(GeneSymbol = case_when(GeneID == "ENSG00000277117" ~ "ENSG00000277117", T ~ GeneSymbol))
procounts_AD %>% dplyr::filter(GeneSymbol == "SYT15")
procounts_AD <- procounts_AD %>% mutate(GeneSymbol = case_when(GeneID == "ENSG00000277758" ~ "ENSG00000277758", T ~ GeneSymbol))
```

```{r}
ensembl.genes <- procounts_MDD$GeneID
geneIDs1 <- ensembldb::select(EnsDb.Hsapiens.v79, keys = ensembl.genes, keytype = "GENEID", columns = 
                                c("SYMBOL","GENEID"))
#head(geneIDs1)
dim(geneIDs1)
geneIDs1 <- geneIDs1 %>% mutate(SYMBOL = case_when(SYMBOL == "" ~ GENEID, T ~ SYMBOL))

colnames(geneIDs1)[names(geneIDs1) == "GENEID"] <- "GeneID"
procounts_MDD <- left_join(geneIDs1, procounts_MDD, by = "GeneID")
procounts_MDD <- procounts_MDD %>% dplyr::select(GeneID, SYMBOL, everything())
colnames(procounts_MDD)[names(procounts_MDD) == "SYMBOL"] <- "GeneSymbol"
#head(procounts_MDD)
dim(procounts_MDD)
```

```{r}
procounts_MDD <- procounts_MDD %>%
  dplyr::filter(GeneID %in% genes_ENSID$ensembl_gene_id)
dim(procounts_MDD) #7251 genes gone 
```

## Add regulators again
```{r}
regulators <- read.table("dbTF.csv", header = T)
#head(regulators)
dim(regulators)
#regulators[duplicated(regulators)]
```

Alzheimer
Add GeneSymbol
```{r}
ensembl.genes <- procounts_AD_N$GeneID
geneIDs1 <- ensembldb::select(EnsDb.Hsapiens.v79, keys = ensembl.genes, keytype = "GENEID", columns = 
                                c("SYMBOL","GENEID"))
#head(geneIDs1)
geneIDs1 <- geneIDs1 %>% mutate(SYMBOL = case_when(SYMBOL == "" ~ GENEID, T ~ SYMBOL))
colnames(geneIDs1)[names(geneIDs1) == "GENEID"] <- "GeneID"
procounts_AD_N <- left_join(geneIDs1, procounts_AD_N, by = "GeneID")
procounts_AD_N <- procounts_AD_N %>% dplyr::select(GeneID, SYMBOL, everything())
colnames(procounts_AD_N)[names(procounts_AD_N) == "SYMBOL"] <- "GeneSymbol"
#head(procounts_AD_N)
dim(procounts_AD_N)
#procounts_AD_N$GeneSymbol[duplicated(procounts_AD_N$GeneSymbol)]
```

```{r}
#head(procounts_AD)
# find TF of regulators in procounts_AD_N
# add these to procounts_AD, but no duplicates
regulators_AD <- procounts_AD_N %>% dplyr::filter(GeneSymbol %in% regulators$Gene_symbol)
#head(regulators_AD)
dim(regulators_AD) #1045
regulators_AD$GeneSymbol[duplicated(regulators_AD$GeneSymbol)]
regulators_AD %>% dplyr::filter(GeneSymbol == "ZNF709")
regulators_AD <- regulators_AD[!(regulators_AD$GeneID == "ENSG00000196826"),]
regulators_AD %>% dplyr::filter(GeneSymbol == "ZNF286A")
regulators_AD <- regulators_AD[!(regulators_AD$GeneID == "ENSG00000255104"),]
regulators_AD %>% dplyr::filter(GeneSymbol == "ZNF670")
"ZNF670-ZNF695" %in% regulators
regulators_AD <- regulators_AD[!(regulators_AD$GeneID == "ENSG00000135747"),]
regulators_AD %>% dplyr::filter(GeneSymbol == "SP110")
regulators_AD <- regulators_AD[!(regulators_AD$GeneID == "ENSG00000280755"),]
dim(procounts_AD) #9210
procounts_AD_ <- full_join(procounts_AD, regulators_AD)
procounts_AD_$GeneSymbol[duplicated(procounts_AD_$GeneSymbol)]
#head(procounts_AD_)
dim(procounts_AD_) #9727 -> 517 extra
```

Depression
```{r}
ensembl.genes <- procounts_MDD_N$GeneID
geneIDs1 <- ensembldb::select(EnsDb.Hsapiens.v79, keys = ensembl.genes, keytype = "GENEID", columns = 
                                c("SYMBOL","GENEID"))
geneIDs1 <- geneIDs1 %>% mutate(SYMBOL = case_when(SYMBOL == "" ~ GENEID, T ~ SYMBOL))

colnames(geneIDs1)[names(geneIDs1) == "GENEID"] <- "GeneID"
procounts_MDD_N <- left_join(geneIDs1, procounts_MDD_N, by = "GeneID")
procounts_MDD_N <- procounts_MDD_N %>% dplyr::select(GeneID, SYMBOL, everything())
colnames(procounts_MDD_N)[names(procounts_MDD_N) == "SYMBOL"] <- "GeneSymbol"
#head(procounts_MDD_N)
```

```{r}
regulators_MDD <- procounts_MDD_N %>% dplyr::filter(GeneSymbol %in% regulators$Gene_symbol)
#head(regulators_MDD)
dim(regulators_MDD) #1101
dim(procounts_MDD) #8660
procounts_MDD_ <- full_join(procounts_MDD, regulators_MDD)
#head(procounts_MDD_)
dim(procounts_MDD_) #9217 -> 557 extra
```


## Scaling
```{r}
Genes_AD <- procounts_AD_ %>% dplyr::select(GeneID, GeneSymbol)
Genes_MDD <- procounts_MDD_ %>% dplyr::select(GeneID, GeneSymbol)
procounts_AD_S <- t(scale(t(procounts_AD_[, c(-1, -2)])))
procounts_AD_S <- cbind(procounts_AD_S, Genes_AD)
procounts_AD_S <- procounts_AD_S %>% dplyr::select(GeneID, GeneSymbol, everything())

procounts_AD_S <- as.data.frame(procounts_AD_S)
procounts_MDD_S <- t(scale(t(procounts_MDD_[, c(-1, -2)])))
procounts_MDD_S <- cbind(procounts_MDD_S, Genes_MDD)
procounts_MDD_S <- procounts_MDD_S %>% dplyr::select(GeneID, GeneSymbol, everything())
procounts_MDD_S <- as.data.frame(procounts_MDD_S)
```

Files of processed counts 
```{r}
#procounts_AD_S <- procounts_AD_S %>% dplyr::select(-GeneID)
#procounts_MDD_S <- procounts_MDD_S %>% dplyr::select(-GeneID)
write.table(procounts_AD_S, file = "Processed_counts_AD.csv", sep = '\t', row.names = F, col.names = T)
write.table(procounts_MDD_S, file = "Processed_counts_MDD.csv", sep = '\t', row.names = F, col.names = T)
```

Write files for Lemon-tree
```{r}
#head(procounts_AD_S)
write.table(procounts_AD_S, file = "Processed_counts_AD.txt", sep = '\t', row.names = F, col.names = T)
write.table(procounts_MDD_S, file = "Processed_counts_MDD.txt", sep = '\t', row.names = F, col.names = T)
```

Regulators
```{r}
regulators_AD <- regulators_AD$GeneSymbol
regulators_MDD <- regulators_MDD$GeneSymbol
regulators_AD <- as.vector(regulators_AD)
regulators_MDD <- as.vector(regulators_MDD)
head(regulators_AD)
write.table(regulators_AD, file = "RegulatorsAD.txt", sep = ',', col.names = F, row.names = F)
write.table(regulators_MDD, file = "RegulatorsMDD.txt", sep = ',', col.names = F, row.names = F)
```

Gene names
```{r}
procounts_AD_G <- procounts_AD_S$GeneSymbol
class(procounts_AD_G)
procounts_MDD_G <- procounts_MDD_S$GeneSymbol
head(procounts_MDD_G)
write.table(procounts_AD_G, file = "GeneNamesAD.txt", sep = ',', col.names = F, row.names = F)
write.table(procounts_MDD_G, file = "GeneNamesMDD.txt", sep = ',', col.names = F, row.names = F)
```

#ModuleViewer
AD
```{r}
modulesAD <- read.table("Lemon-Tree/tight_clustersAD.txt")
dim(modulesAD) #8674
#head(procounts_AD_)
dim(procounts_AD_) #9724
procounts_AD_G <- procounts_AD_ %>% dplyr::select(GeneID, GeneSymbol)
procounts_AD_T <- procounts_AD_G %>% dplyr::filter(GeneSymbol %in% modulesAD$V1)
dim(procounts_AD_T) #8246
colnames(modulesAD)[names(modulesAD) == "V1"] <- "GeneSymbol"
modulesAD <- full_join(modulesAD, procounts_AD_T, by = "GeneSymbol")
modulesAD <- modulesAD %>% dplyr::select(GeneSymbol, GeneID, V2)
head(modulesAD)
modulesAD <- modulesAD %>% dplyr::select(GeneID, V2)
write.table(modulesAD, file = "ModulesAD.txt", sep = '\t', col.names = F, row.names = F)
```
MDD
```{r}
modulesMDD <- read.table("Lemon-Tree/tight_clustersMDD.txt")
dim(modulesMDD) #8674
#head(procounts_MDD_)
dim(procounts_MDD_) #9724
procounts_MDD_G <- procounts_MDD_ %>% dplyr::select(GeneID, GeneSymbol)
procounts_MDD_T <- procounts_MDD_G %>% dplyr::filter(GeneSymbol %in% modulesMDD$V1)
dim(procounts_MDD_T) #8246
colnames(modulesMDD)[names(modulesMDD) == "V1"] <- "GeneSymbol"
modulesMDD <- full_join(modulesMDD, procounts_MDD_T, by = "GeneSymbol")
modulesMDD <- modulesMDD %>% dplyr::select(GeneSymbol, GeneID, V2)
head(modulesMDD)
modulesMDD <- modulesMDD %>% dplyr::select(GeneID, V2)
write.table(modulesMDD, file = "ModulesMDD.txt", sep = '\t', col.names = F, row.names = F)
```

With GeneID
```{r}
write.table(procounts_AD_S, file = "Processed_counts_AD.txt", sep = '\t', row.names = F, col.names = T, quote = F)
write.table(procounts_MDD_S, file = "Processed_counts_MDD.txt", sep = '\t', row.names = F, col.names = T, quote = F)
```

Name mapping for ModuleViewer
```{r}
procounts_AD_G <- procounts_AD_S %>% dplyr::select(GeneID, GeneSymbol)
head(procounts_AD_G)
procounts_MDD_G <- procounts_MDD_S %>% dplyr::select(GeneID, GeneSymbol)
write.table(procounts_AD_G, file = "NameMappingAD.txt", sep = '\t', col.names = F, row.names = F, quote = F)
write.table(procounts_MDD_G, file = "NameMappingMDD.txt", sep = '\t', col.names = F, row.names = F, quote = F)
```

