---
title: "RcisTarget"
author: "Hanne Puype"
date: "2023-12-21"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```


The function cisTarget() allows to perform the motif-enrichment analysis on a gene list. The main input parameters are the gene list and the motif databases, which should be chosen depending on the organism and the search space around the TSS of the genes.\
<https://resources.aertslab.org/cistarget/databases/homo_sapiens/hg38/refseq_r80/mc_v10_clust/gene_based/>\
<https://resources.aertslab.org/cistarget/motif2tf/>  

## Load packages
```{r}
library(tidyverse)
library(RcisTarget)
library(arrow)
library(data.table)
library(DT)
library(reshape2)
library(visNetwork)
```

## Load modules
```{r}
modulesAD <- read.table("kmed_modules_AD.txt", header = F, sep = "\t")
modulesMDD <- read.table("kmed_modules_MDD.txt", header = F, sep = "\t")
colnames(modulesAD) <- c("Gene", "Module")
colnames(modulesMDD) <- c("Gene", "Module")
modulesAD <- as_tibble(modulesAD)
modulesMDD <- as_tibble(modulesMDD)
B <- modulesAD %>% group_by(Module) %>% mutate(rid = row_number()) %>% spread(Module, Gene)
for (i in 2:length(names(B))){
  names(B)[i] <- paste0('module', names(B)[i])
}
B_MDD <- modulesMDD %>% group_by(Module) %>% mutate(rid = row_number()) %>% spread(Module, Gene)
for (i in 2:length(names(B_MDD))){
  names(B_MDD)[i] <- paste0('module', names(B_MDD)[i])
}
```

## RcisTarget
Currently the function importAnnotations doesn't work  
```{r}
# select motif database
# data(motifAnnotations_hgnc) # for previous version 
motifAnnotations_hgnc_v10 <- read.table("motifs-v10nr_clust-nr.hgnc-m0.001-o0.0.txt", 
                                        header = T, sep = "\t")
motifAnnotations_hgnc_v10 <- motifAnnotations_hgnc_v10 %>% rename("motif" = "motif_id")
motifAnnotations_hgnc_v10 <- motifAnnotations_hgnc_v10 %>% rename("TF" = "gene_name")
motifAnnotations_hgnc_v10 <- as.data.table(motifAnnotations_hgnc_v10)
#motifAnnotations_hgnc_v10 <- importAnnotations("RcisTarget/motifs-v10nr_clust-nr.hgnc-m0.001-o0.0.txt")
motifRankings <- importRankings("hg38_10kbp_up_10kbp_down_full_tx_v10_clust.genes_vs_motifs.rankings.feather",
                                indexCol = "motifs")
```

Regulators 
```{r}
reg_AD <- read.table("reg_clust_ens_AD.txt", header = F, sep = "\t")
reg_MDD <- read.table("reg_clust_ens_MDD.txt", header = F, sep = "\t")
names(reg_AD) <- c("TF", "Module")
names(reg_MDD) <- c("TF", "Module")
```

### Loop over all modules
Retrieve TFs which were found by RcisTarget (motif) and inferred 
Put in mvf file for ModuleViewer
AD
```{r}
modules_loop <- paste0("module", c(1:155))
i <- 1
df <- data.frame("Module" = numeric(0), "enrichedGenes" = character(0), "Tf" = character(0))

for (module in modules_loop) {
  modIds <- B[module]
  modIds <- as_tibble(modIds)
  names(modIds) <- "GeneSymbol"
  geneList <- modIds$GeneSymbol
  geneList <- geneList[!is.na(geneList)]
  geneLists <- list(module_genes = geneList)
  
  # motif enrichment analysis
  motifEnrichmentTable_wGenes <- cisTarget(geneLists, motifRankings,
                                 motifAnnot = motifAnnotations_hgnc_v10)
  
  # TFs annotated to enriched motifs
  motifEnrichmentTable_wGenes <- motifEnrichmentTable_wGenes %>% 
    left_join(motifAnnotations_hgnc_v10, by = c("motif" = "motif"))
  motifEnrichmentTable_wGenes <- motifEnrichmentTable_wGenes %>% select(geneSet, motif, TF, NES, AUC, nEnrGenes, enrichedGenes)
  motifEnrichmentTable_wGenes <- motifEnrichmentTable_wGenes %>% distinct(TF, .keep_all = T)
  write.table(motifEnrichmentTable_wGenes, paste0("RcisTarget/MotifEnrich_AD/motifEnrichmentTable",module,".txt"), sep = "\t", row.names = F, quote = F)
  
  # Which enriched motifs overlap with inferred regulators
  reg_AD_module <- reg_AD %>% filter(Module == i)
  motifEnrichmentTable_wGenes_inf <- motifEnrichmentTable_wGenes[motifEnrichmentTable_wGenes$TF %in% reg_AD_module$TF,]
  # for ModuleViewer we need: module   genes|again|split  term  
  motifEnrichmentTable_wGenes_inf <- motifEnrichmentTable_wGenes_inf %>% add_column(Module = i) %>% dplyr::select(Module, enrichedGenes, TF)
  motifEnrichmentTable_wGenes_inf$enrichedGenes <- gsub(";", "|", motifEnrichmentTable_wGenes_inf$enrichedGenes)
  df <- rbind(df, motifEnrichmentTable_wGenes_inf)
  i <- i + 1
}
dim(df) # 103 
write.table(df, "Motif_enrich_RcisTarget_inferred_AD.mvf", quote = F, sep = "\t", row.names = F, col.names = F)
```

```{r}
dim(modulesAD[!(modulesAD$Gene %in% colnames(motifRankings@rankings)),]) # 179 genes of modules not in motifRankings
```

MDD
```{r}
modules_loop <- paste0("module", c(1:156))
i <- 1
df <- data.frame("Module" = numeric(0), "enrichedGenes" = character(0), "Tf" = character(0))

for (module in modules_loop) {
  modIds <- B_MDD[module]
  modIds <- as_tibble(modIds)
  names(modIds) <- "GeneSymbol"
  geneList <- modIds$GeneSymbol
  geneList <- geneList[!is.na(geneList)]
  geneLists <- list(module_genes = geneList)
  
  # motif enrichment analysis
  motifEnrichmentTable_wGenes <- cisTarget(geneLists, motifRankings,
                                 motifAnnot = motifAnnotations_hgnc_v10)
  
  # TFs annotated to enriched motifs
  motifEnrichmentTable_wGenes <- motifEnrichmentTable_wGenes %>% 
    left_join(motifAnnotations_hgnc_v10, by = c("motif" = "motif"))
  motifEnrichmentTable_wGenes <- motifEnrichmentTable_wGenes %>% select(geneSet, motif, TF, NES, AUC, nEnrGenes, enrichedGenes)
  motifEnrichmentTable_wGenes <- motifEnrichmentTable_wGenes %>% distinct(TF, .keep_all = T)
  write.table(motifEnrichmentTable_wGenes, paste0("RcisTarget/MotifEnrich_MDD/motifEnrichmentTable",module,".txt"), sep = "\t", row.names = F, quote = F)
  
  # Which enriched motifs overlap with inferred regulators
  reg_MDD_module <- reg_MDD %>% filter(Module == i)
  motifEnrichmentTable_wGenes_inf <- motifEnrichmentTable_wGenes[motifEnrichmentTable_wGenes$TF %in% reg_MDD_module$TF,]
  # for ModuleViewer we need: module   genes|again|split  term  
  motifEnrichmentTable_wGenes_inf <- motifEnrichmentTable_wGenes_inf %>% add_column(Module = i) %>% dplyr::select(Module, enrichedGenes, TF)
  motifEnrichmentTable_wGenes_inf$enrichedGenes <- gsub(";", "|", motifEnrichmentTable_wGenes_inf$enrichedGenes)
  df <- rbind(df, motifEnrichmentTable_wGenes_inf)
  i <- i + 1
}
dim(df) # 193 
write.table(df, "Motif_enrich_RcisTarget_inferred_MDD.mvf", quote = F, sep = "\t", row.names = F, col.names = F)
```

```{r}
dim(modulesMDD[!(modulesMDD$Gene %in% colnames(motifRankings@rankings)),]) # 156 genes of modules not in motifRankings
```

The final output of RcisTarget is a data.table containing the information about the motif enrichment and its annotation organized in the following fields:\
- geneSet: Name of the gene set\
- motif: ID of the motif\
- NES: Normalized enrichment score of the motif in the gene-set\
- AUC: Area Under the Curve (used to calculate the NES)\
- TFinDB: Indicates whether the highlighted TFs are included within the high confidence annotation (two asterisks) or low confidence annotation (one asterisk).\
- TF_highConf: Transcription factors annotated to the motif according to 'motifAnnot_highConfCat'.\
- TF_lowConf: Transcription factors annotated to the motif according to 'motifAnnot_lowConfCat'.\
- enrichedGenes: Genes that are highly ranked for the given motif.\
- nErnGenes: Number of genes highly ranked\
- rankAtMax: Ranking at the maximum enrichment, used to determine the number of enriched genes.

